---
title: "Effects of crop rotation on common waterhemp population dynamics: a prospective analysis"
author: "Huong Nguyen"
date: "10/19/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction 
Seedbank depletion can be promoted by inducing seed fatal germination so that seeds germinate, but the seedlings fail to reach the soil surface and thus die [@davisInfluenceSeedDepth2007], and by fostering granivore activities so that viable, dormant seeds are consumed or damaged [@bagavathiannanPostdispersalLossImportant2013; @kurstjensPreciseTillageSystems2007]. 

*why first column only*

### Starting density  

Emergence: use emergence density from 2018, 2019 and 2020 (Number_begin by cohort)

## Seedbank

### Tillage-induced seedbank composition  

### Recruiment: emergence as proportion of seedbank density 

## Seedling

### Growth and survival  

## Mature ??
### Fecundity estimation from aboveground mass at maturity 

### Mature density: reported in Waterhemp-fecundity 
, with the assumptions that 1) the sex ratio parity [@costeaBiologyInvasiveAlien2005] applies to all stages except for mature plants [@nguyenImpactCroppingSystem2022], 2) a waterhemp plant's sex is determined at seed formation [@montgomerySexspecificMarkersWaterhemp2019; @montgomeryMaleSpecificChromosomal2021], and 3) pollens are affluently available to all the female flowers during the reproductive period. Data collection for each of the demographic details were provided below. 
### Female density: 

```{r}
ft1 <- head(iris) %>%
  flextable() %>%
set_header_labels(
  values = list(
    Sepal.Length = "S_1",
    Sepal.Width = "S_2",
    Petal.Length = "S_3",
    Petal.Width = "S_4"
  )
)

my_list <- setNames(as.list(paste0("\U03A3_", 1:4)), names(iris[1:4]))

ft1 %>%
  set_header_labels(
  values = my_list )
```

```{r, child= "~/Documents/Dissertation/AMATA-demography/5-Report/emergence-pattern-timing-Julian.Rmd"}
# diagnosis plots still look concerning
```