---
output:
    bookdown::word_document2:
      toc: false
      fig_caption: yes
      keep_md: true
#    reference_docx: style_template.docx
bibliography: WH-pop-dynamics.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(emmeans)
library(here)
library(nlme)
library(flextable)
```


```{r seedbank-19, include=FALSE}
## Read (female) seed column data (from /AMATA-demography/4-Analysis/seedbank-fates-2014-through-2019.Rmd at chunk seed-column-19) 
 seed_column <- readRDS("../2-Data/Raw/seed-column-2019-list.RData")

#seed_column_df <- readRDS("../2-Data/Raw/seed-column-2019-df.RData")

seed_column_round <- readRDS("../2-Data/Raw/seed-column-2019-list-poisson.RData")
## To track possible errors, the seed columns are multiplied by each matrix step-by-step
```

The seed column before post-harvest tillage was transitioned from post-harvest tillage through spring tillage with the general form of $A_{h+1} = A_h N_h$ by @caswellMatrixPopulationModels2001.  

```{r post-harvest-tillage, include=FALSE}
## Raw data from 'FLOW-literature-based-matrices'
mean_post_harvest_tillage <- readRDS("../2-Data/Clean/mean-post-harvest-tillage.RData")

mean_post_harvest_tillage_pop <- purrr::map2(mean_post_harvest_tillage, seed_column, `%*%`)
```

```{r overwinter-seed-survival, include=FALSE}
## Scenario 1 - Sosnoskie et al., 2013
mean_overwinter_scenario1 <- readRDS("../2-Data/Clean/mean-winter-seed-survival-Sosnoskie.RData")

mean_over_winter_pop_scenario1 <- purrr::map2(mean_overwinter_scenario1, mean_post_harvest_tillage_pop, `%*%`)
```

```{r spring-tillage, include=FALSE}
## Load spring tillage matrices
mean_spring_tillage <-  readRDS("../2-Data/Clean/mean-pre-planting-tillage.RData")

## Apply spring tillage on overwinter-ed seed columns under the condition reported by Sosnoskie et al. 2013
mean_after_spring_tillage_pop_scenario1 <- purrr::map2(mean_spring_tillage, mean_over_winter_pop_scenario1, `%*%`)

## Keep the non-zeros in the 8x1 columns only, making a set of 2x1 columns
mean_after_spring_tillage_pop_scenario1_2by1 <- mean_after_spring_tillage_pop_scenario1   %>%
  map(~ head(.,2))

```


```{r top-stratum-after-spring-tillage}

### Write the top stratum of the newly shuffled seed columns (from fall tillage through spring tillage) into a dummy matrix
# https://community.rstudio.com/t/extract-matrix-rows-from-list/19357/2
#https://stackoverflow.com/questions/29511215/convert-row-names-into-first-column

mean_after_spring_tillage_pop_scenario1_top_stratum_df <- mean_after_spring_tillage_pop_scenario1  %>%
  map(~.x[1, ])  %>% 
  unlist(use.names = TRUE) %>%
  as.data.frame() %>%
  rownames_to_column("matrix_id") %>% 
  rename(top_stratum_density = ".") %>%
  mutate(top_stratum_female_density = top_stratum_density/2) #assume 1:1 male:female


#mean_after_spring_tillage_pop_scenario1_dummy <-  lapply(split(mean_after_spring_tillage_pop_scenario1_df, mean_after_spring_tillage_pop_scenario1_df$matrix_id),
#       function(x)(matrix(x$top_stratum_density, nrow = 6)))

mean_after_spring_tillage_pop_scenario1_bottom_stratum_df <- mean_after_spring_tillage_pop_scenario1  %>%
  map(~.x[2, ])  %>% 
  unlist(use.names = TRUE) %>%
  as.data.frame() %>%
  rownames_to_column("matrix_id") %>% 
  rename(bottom_stratum_density = ".") %>%
  mutate(bottom_stratum_female_density = bottom_stratum_density/2)

```

```{r emergence-20-density-matrix, include=FALSE}
## 2020 Emergence density, non-integers, from AMATA-demography/female-emergence chunk emerge-20-emm-df

#New, in dataframe format: 
female_emerge_20_mean_df <- readRDS("../2-Data/Raw/six-cohort-female-emerge-df-2020-raw.RData")

# Old: emerge_20 <- readRDS("../2-Data/Raw/six-cohort-female-emerge-list-2020-raw.RData")

## Turn the emerge_20_mean_df to wide format 
female_emerge_20_mean_wide <- female_emerge_20_mean_df %>%
  pivot_wider(names_from = Cohort, values_from = response)
```


**include in the matrix assembly section**
In a soil seedbank of 5 cm deep that was undisturbed mechanically in the first burial year and unexposed to herbicides throughout the experiment, 
5% of the waterhemp seedlings emerged a year after seed burial [@buhlerEmergencePersistenceSeed2001]. Annually, 23.5% +/- 16.6% sd of waterhemp seeds that were not treated with herbicides and undisturbed mechanically emerged from the top 1 cm soil layer [@schutteCommonWaterhempAmaranthus2014]. Mesotrione applied at 75 g ha$^-1$ rate was 76% and 96% efficacious against *A retroflexus* L grown in corn that were susceptible and resistant to atrazine, respectively [@suttonActivityMesotrioneResistant2002]. The Thiencarbazone-methyl + isoxaflutole mixture was 93.5% efficacious and mesotrione was 70.75% efficacious against *A. palmeri* grown in corn [@janakWeedControlCorn2016]. Waterhemp control in soybean treated with pre-emergence, post-emergence, or sequential pre- and post-emergence herbicides [Jhala, Hay, ...] was studied but no information on seedling emergence as a proportion of the seedbank is available. 


The resistance profile of waterhemp at our experiment site was undetermined, but the raw estimation of seedling emergence proportion with respect to the top 0 - 2 cm soil seedbank density seems unrealistically low (Table \@ref(tab:general-emerge-assign). We combined the findings on other *Amaranthus* species from @suttonActivityMesotrioneResistant2002; @janakWeedControlCorn2016 for herbicide efficacy and from @buhlerEmergencePersistenceSeed2001 and @schutteCommonWaterhempAmaranthus2014 for herbicide-unexposed germinants to assign emergence rates based on the crop-specific weed management (Table \@ref(tab:general-emerge-assign)). 

The same emergence successful rate was assumed the same in corn and soybean for the lack of available data in the soybean crop environment. The emergence successful rates in the oat/red clover intercrop, oat/alfalfa intercrop, and alfalfa sole crop were set at 50% of all germinated seeds to reflect a weaker potency of allelochemicals versus pre-emergence herbicides. With a uniform germination rate at 20% in all the crop identity crossed with corn weed management, the remaining seedbank density in the 0 - 2 cm soil stratum is calculated using the following equation: 

$$
N_{B_{e,11}} = (1 - G)*N_{B_{st,11}} = (1 - (E + D)) * N_{B_{st,11}}
$$
where,  
$N_{B_{e,11}}$ is the remaining seedbank density after seed germination
$N_{B_{st,11}}$ is the seed density in the top 0 - 2 cm soil stratum upon completion of pre-planting tillage
$G$ is the proportion of germinated seeds 
$E$ is the proportion of germinated seeds that successfully emerge as seedlings 
$D$ is the proportion of germinated seeds that were killed by weed control measures

$1 - G = 0.8$ is filled in the [1,1] position of the seedling recruitment matrix ($B_e[1,1]$). The cohort-specific emergence rates ($e_1$ through $e_6$) were adjusted from the raw data ($\frac{p_k}{N_{B_{st,11}}}$ with $k = \{1,...,6\}$) to reflect 5% emergence success rate (equivalent to $E = 0.01 * N_{B_{st,11}}$) in crop environments that received pre-emergence herbicides (C2, C3 and C4 under conventional weed management and all the S2, S3, and S4 [Table 1, @nguyenWeedCommunityComposition2022]), 100% emergence success rate (equivalent to $E = 1 * N_{B_{st,11}}$) in the crop environments that received post-emergence herbicides (C2, C3 and C4 under low herbicide weed management, [Table 1, @nguyenWeedCommunityComposition2022]) and 50% emergence success rate (equivalent to $E = 0.1 * N_{B_{st,11}}$).   




```{r emerge-prop-tab, echo=FALSE}
#Reality check of emergence proportion
# merge whole-column seed density with top stratum density
female_emerge_prop_20 <- female_emerge_20_mean_wide %>% 
  left_join(mean_after_spring_tillage_pop_scenario1_top_stratum_df, by = "matrix_id") %>%
  left_join(mean_after_spring_tillage_pop_scenario1_bottom_stratum_df, by = "matrix_id") %>%
  mutate(whole_seedbank_female_density = top_stratum_female_density + bottom_stratum_female_density) %>%
  mutate(cohort1_mean_prop_wrt_top = `1`/top_stratum_female_density,
         cohort2_mean_prop_wrt_top = `2`/top_stratum_female_density,
         cohort3_mean_prop_wrt_top = `3`/top_stratum_female_density,
         cohort4_mean_prop_wrt_top = `4`/top_stratum_female_density,
         cohort5_mean_prop_wrt_top = `5`/top_stratum_female_density,
         cohort6_mean_prop_wrt_top = `6`/top_stratum_female_density,
         cohort1_mean_prop_wrt_whole = `1`/whole_seedbank_female_density,
         cohort2_mean_prop_wrt_whole = `2`/whole_seedbank_female_density,
         cohort3_mean_prop_wrt_whole = `3`/whole_seedbank_female_density,
         cohort4_mean_prop_wrt_whole = `4`/whole_seedbank_female_density,
         cohort5_mean_prop_wrt_whole = `5`/whole_seedbank_female_density,
         cohort6_mean_prop_wrt_whole = `6`/whole_seedbank_female_density) %>%
  mutate(assumed_total_germination_rate = 0.2,
         assumed_total_successful_emergence_rate = ifelse(Crop_ID %in% c("O4", "O3", "A4"), 0.5, "NA")) 

female_emerge_prop_20$assumed_total_successful_emergence_rate[c(3, 5, 7, 13:18)] <- 0.05
female_emerge_prop_20$assumed_total_successful_emergence_rate[c(4,6,8)] <- 1

female_emerge_prop_20$assumed_total_successful_emergence_rate <- as.numeric( female_emerge_prop_20$assumed_total_successful_emergence_rate)

# View(female_emerge_prop_20)
# table of arithmetic means

```


```{r general-emerge-assign, echo=FALSE, warning=FALSE, message=FALSE}

female_emerge_prop_20_total <- female_emerge_prop_20 %>%
  select(Crop_ID, Corn_weed_management, `1`:`6`,
         top_stratum_female_density, 
         assumed_total_germination_rate,
         assumed_total_successful_emergence_rate) %>%
  mutate(total_emerge_from_top_estimated = `1` + `2` + `3` + `4` + `5`+ `6`,
         total_emerge_from_top_estimated_prop = total_emerge_from_top_estimated/top_stratum_female_density,
         total_emerge_from_top_assigned =
           assumed_total_germination_rate*assumed_total_successful_emergence_rate) %>%
  mutate(cohort_1_share = total_emerge_from_top_assigned * `1`/total_emerge_from_top_estimated,
         cohort_2_share = total_emerge_from_top_assigned * `2`/total_emerge_from_top_estimated,
         cohort_3_share = total_emerge_from_top_assigned * `3`/total_emerge_from_top_estimated,
         cohort_4_share = total_emerge_from_top_assigned * `4`/total_emerge_from_top_estimated,
         cohort_5_share = total_emerge_from_top_assigned * `5`/total_emerge_from_top_estimated,
         cohort_6_share = total_emerge_from_top_assigned * `6`/total_emerge_from_top_estimated) 

# add crop-specific weed management or herbicide


female_emerge_prop_20_total %>%
  mutate(phase_order = ifelse(str_detect(Crop_ID, "C"), 1,
                              ifelse(str_detect(Crop_ID, "S"), 2,
                                     ifelse(str_detect(Crop_ID, "O"), 3, 4)))) %>%
  mutate(Rot = substr(Crop_ID,2,2))%>%
  arrange(Rot, phase_order, Corn_weed_management) %>%

  select(Crop_ID, Corn_weed_management, 
         total_emerge_from_top_estimated_prop,
         total_emerge_from_top_assigned, cohort_1_share:cohort_6_share) %>%
  mutate_at(vars(total_emerge_from_top_estimated_prop, cohort_1_share:cohort_6_share), funs(round(., 5))) %>%
  flextable() %>%
    set_caption("Estimated seedling emergence proportion with respect to the top 0 - 2 cm soil stratum using stratified soil seedbank densities sampled in 2019 and  seedling emergence densities sampled in 2020; and adjusted seedling emergence proportions for use in the modelling excercises. The seedbank densities in 2019 were vertically redistributed with two passes of tillage: post-harvest (fall 2019) and pre-planting (spring 2020). Between the two passes of tillage, overwinter seed survial rates were calculated using the equations provided in Figures 1 and 3 of Sosnoskie et al., 2013.") %>%
  set_header_labels(values = list(Crop_ID = "Crop ID",
                Corn_weed_management = "Corn weed management",
               total_emerge_from_top_estimated_prop = "0 - 2 cm soil stratum",
               total_emerge_from_top_assigned = "0 - 2 cm soil stratum",
               cohort_1_share = "cohort 1",
               cohort_2_share = "cohort 2",
               cohort_3_share = "cohort 2",
               cohort_4_share = "cohort 4",
               cohort_5_share = "cohort 5",
               cohort_6_share = "cohort 6")) %>%
  add_header_row(top = TRUE, values = c(" ", " ", 
                                        "Estimated total emergence proportion from",
                                        "Assigned total emergence proportion from",
                                        "Adjusted total emergence proportion of"), 
                 colwidths = c(1, 1,  1, 1, 6)) %>%
   autofit() 

  
```



```{r recruit-list, include=FALSE}
# save emergence matrices in a list 

female_emerge_prop_cohort <- female_emerge_prop_20_total %>%
  select(Crop_ID, Corn_weed_management, cohort_1_share:cohort_6_share) %>%
  mutate(matrix_id = paste(Crop_ID, Corn_weed_management, sep = "_")) %>%
  pivot_longer(!c(Crop_ID, Corn_weed_management, matrix_id), names_to='cohort',
                    values_to='emerge')

female_emerge_prop_cohort_list <- lapply(split(female_emerge_prop_cohort, 
                                                  female_emerge_prop_cohort$matrix_id),
       function(x) rbind(cbind(matrix(c(0.8, 0, 0, 1), byrow = TRUE, ncol = 2), matrix(0, nrow = 2, ncol = 6)),
                         cbind(matrix(x$emerge, nrow = 6), matrix(0, ncol = 7, nrow = 6))))
                         


# saveRDS(female_emerge_prop_cohort_list, file="../2-Data/Clean/mean-emergence-prop-adjusted.RData")
```


