---
output:
    bookdown::word_document2:
      toc: false
      fig_caption: yes
    reference_docx: style_template.docx
---

Matrices in this document are listed chronologically, from natural seed rain (at AMATA senescence) to new seed production. 

```{r, include=FALSE, message=FALSE}
library(emmeans)
library(here)
library(readr)
library(data.table) #for rbindlist
library(tidyverse)
```

```{r fresh-seed-viability}
# Table 2 in Buhler and Hartzler, 2001, at <http://www.jstor.com/stable/4046508> 

fresh_seed_viability_BH <- expand.grid(Crop_ID = c("C2","S2",
                                                "C3","S3","O3",
                                                "C4", "S4", "O4", "A4"),
                                    Corn_weed_management = c("conv","low"),
                                    Cohort = c(1:6),
                                  Viability = 0.83)

fresh_seed_viability_BH_matrix <- fresh_seed_viability_BH %>%
  mutate(matrix_id = paste(Crop_ID,Corn_weed_management, sep = "_")) %>%
  arrange(matrix_id)

fresh_seed_viability_BH_matrix_list <- lapply(split(fresh_seed_viability_BH_matrix, fresh_seed_viability_BH_matrix$matrix_id),
       function(x)(diag(x$Viability, 6)))  #seed fecundity matrix and rearrange if needed

# Table 1 in Korres et al., 2018, at <https://doi.org/10.1017/wsc.2018.27>
fresh_seed_viability_Korres <- expand.grid(Crop_ID = c("C2","S2",
                                                "C3","S3","O3",
                                                "C4", "S4", "O4", "A4"),
                                    Corn_weed_management = c("conv","low"),
                                    Cohort = c(1:6),
                                  Viability = 0.935) # mean(c(98, 97.3, 92, 88.2, 91.9)) = 93.48


```

SeedChaser is a Java application that simulates how seeds are moved vertically by a tillage method. The Seed Chaser tillage matrices relevant to this exercise are field cultivator, moldboard, and chisel. The Seed Chaser matrices describe seed movement in 1 cm increment from 0 to 20 cm depth while our soil samples were taken down to 20 cm depth. The raw data was provided by Dr. Kurt Spokas.

The overall mechanism of resizing the matrices is the law of total probability. Each column in the original and the resized matrices sums to 1. To resize a 20x20 matrix to a 2x2 matrix using the raw Seed Chaser data, four summations were performed:  
- Seeds that stay at the top 2 cm stratum, $X_{1,1}$, is the summation of the top left 3x3 section in the original post-tillage matrix.
- Seeds that move from the top 2 cm stratum to the bottom 18 cm stratum, $X_{1,2}$, is the summation of the bottom left 18x3 section in the original post-tillage matrix.
- Seeds that move from the bottom 18 cm stratum to the top 2 cm stratum, $X_{2,1}$, is the summation of the top right 3x18 section in the original post-tillage matrix.
- Seeds that stay at the bottom 18 cm stratum, $X_{2,2}$, is the summation of the bottom right 18x18 section in the original post-tillage matrix.


The mean proportion of each $P_{ij}$, the probability that a seed moves from stratum $i$ to stratum$j$ was pooled over all the replications for each tillage method: $P_{i,j} = \frac{X_{i,j}}{n_{.j}}$, where $X_{ij}$ is the corresponding number of seeds moving from stratum $i$ to stratum $j$ and $n_{.j}$ is the total number of seed at stratum $j$ (top or bottom).

The variance of P[i,j] is calculated with $Var(\frac{X_{ij}}{n_{.j}}) = \frac{p_{ij}(1-p_{ij})}{n_{.j}}$  
```{r post-harvest-tillage-matrix-new}
notill_mean <- diag(1,2)
notill_var <- diag(0,2)

moldboard_mean <- matrix(c(0.017857143,	0.072961329,
0.982142857,	0.927038671), byrow = TRUE, nrow = 2)

moldboard_var <- matrix(c(5.5269e-05,	-1.43463e-05,
-5.5269e-05,	1.43463e-05), byrow = TRUE, nrow = 2)

chisel_mean <- matrix(c(0.589108911,	0.103123159,
0.410891089,	0.896876841), byrow = TRUE, nrow = 2)

chisel_var <- matrix(c(0.000599157,	-5.45013e-05,
-0.000599157,	5.45013e-05), byrow = TRUE, nrow = 2)


phX1 <- c(rep(chisel_mean[,1],6), rep(notill_mean[,1],8), rep(moldboard_mean[,1],4)) 
phX2 <- c(rep(chisel_mean[,2],6), rep(notill_mean[,2],8), rep(moldboard_mean[,2],4)) 
matrix_id <- c(rep("C2_low", 2), rep("C2_conv",2),
                                    rep("C3_low", 2), rep("C3_conv",2),
                                    rep("C4_low", 2), rep("C4_conv",2),
                                    rep("S2_low", 2), rep("S2_conv",2),
                                    rep("S3_low", 2), rep("S3_conv",2),
                                    rep("S4_low", 2), rep("S4_conv",2),
                                    rep("O4_low", 2), rep("O4_conv",2),
                                    rep("A4_low", 2), rep("A4_conv",2),
                                    rep("O3_low", 2), rep("O3_conv",2))

mean_post_harvest_tillage_df <- data_frame(phX1, phX2, matrix_id)

mean_post_harvest_tillage_list <- lapply(split(mean_post_harvest_tillage_df , mean_post_harvest_tillage_df$matrix_id),
       function(x)(rbind(cbind(matrix(c(x$phX1, x$phX2), byrow = FALSE, nrow=2),
                   matrix(rep(0,12), nrow = 2)),
                   matrix(rep(0,48), nrow = 6))))

# saveRDS(mean_post_harvest_tillage_list , file="../2-Data/Clean/mean-post-harvest-tillage.RData")

phV1 <- c(rep(chisel_var[,1],6), rep(notill_var[,1],8), rep(moldboard_var[,1],4)) 
phV2 <- c(rep(chisel_var[,2],6), rep(notill_var[,2],8), rep(moldboard_var[,2],4)) 

var_post_harvest_tillage_df <- data_frame(phV1, phV2, matrix_id)

var_post_harvest_tillage_list <- lapply(split(var_post_harvest_tillage_df, 
                                              var_post_harvest_tillage_df$matrix_id),
       function(x) rbind(cbind(matrix(c(x$phV1, x$phV2), byrow = FALSE, nrow=2),
                               matrix(rep(0,12), nrow = 2)),
                         matrix(rep(0,48), nrow = 6)))

# saveRDS(var_post_harvest_tillage_list, file="../2-Data/Clean/var-post-harvest-tillage.RData")
```

Before the raw data was available, the following steps were performed
The overall mechanism of resizing the matrices is the law of total probability. Each column in the original and the resized matrices sums to 1. For example: 

$$ 
T_{2,1} =  P[top | bottom] = P[top | bottom 1] P[depth1 | bottom] + P[top | bottom 2] P[bottom 2 | bottom] + ... + P[top | bottom 20] P[bottom 20 | bottom]
$$
where,  
P[top | bottom] is the probability that a seed move from the bottom stratum to the top,  
P[top | depth i] is the probability that a seed move from depth i to the top stratum, and  
P[depth i | bottom] is the probability that a seed at depth i remain in the bottom stratum.  

Each of the original 20 x 20 matrices was resized to 2 x 2 using the following steps: 
+ Round 1: Calculation of the probability of a seed in the top 0 - 2 cm stratum remains in the top ($T_{1,1}$) and the probability that that it moves to the bottom 2 - 20 cm stratum ($T_{1,2}$). 

- Distribute 4800 seeds evenly from 0 to 2 cm by filling 1600 to the first three elements of a 18 x 1 vector.
- Fill the remaining sixteen elements with 0.
- Run one pass of a tillage method to redistribute the seed column.
- Divide each element of the redistributed seed column by 4800 to make a top-down vector.
- Sum the first three elements of the top-down vector to have $T_{1,1}$ and the last sixteen elements to have $T_{1,2}$.

+ Round 2: Calculation of the probability of a seed in the bottom 2 - 20 cm stratum remains in the bottom stratum ($T_{2,2}$) and the probability that it moves to the top 0 - 2 stratum ($T_{2,1}$). 

- Distribute 4800 seeds evenly from 3 to 18 cm by filling 3000 to the last sixteen elements of a 18 x 1 vector.
- Fill the remaining three elements with 0.
- Run one pass of a tillage method to redistribute the seed column.
- Divide each element of the redistributed seed column by 4800 to make a bottom-up vector.
- Sum the first three elements of the bottom-up vector to have $T_{2,1}$ and the last sixteen elements to have $T_{2,2}$.

```{r post-harvest-tillage-matrix-old, include=FALSE}
## remove later?

## Old matrices, when raw data wasn't available
# Post-harvest tillage, imported from folder `Tillage-resized`
chisel <- read_csv("../2-Data/Raw/Tillage-resized/chisel_2by2.csv", col_names = FALSE, show_col_types = FALSE)
names(chisel) <- NULL

moldboard <- read_csv("../2-Data/Raw/Tillage-resized/moldboard_2by2.csv", col_names = FALSE, show_col_types = FALSE)
#names(moldboard) <- NULL

no_till <- read_csv("../2-Data/Raw/Tillage-resized/notill_2by2.csv", col_names = FALSE, show_col_types = FALSE)
#names(no_till) <- NULL
# Make a list of post-harvest tillage practices
## C2, C3, C4 are followed by chisel
## S2, S3, S4; and O4 are followed by no till
## O3 and A4 are followed by moldboard

post_harvest_tillage <- rbindlist(list(chisel, chisel, chisel, chisel, chisel, chisel,
                                       no_till, no_till, no_till, no_till,
                                       no_till, no_till, no_till, no_till,
                                       moldboard, moldboard, moldboard, moldboard),
                              use.names = TRUE, fill = FALSE)

post_harvest_tillage$matrix_id <- c(rep("C2_low", 2), rep("C2_conv",2),
                                    rep("C3_low", 2), rep("C3_conv",2),
                                    rep("C4_low", 2), rep("C4_conv",2),
                                    rep("S2_low", 2), rep("S2_conv",2),
                                    rep("S3_low", 2), rep("S3_conv",2),
                                    rep("S4_low", 2), rep("S4_conv",2),
                                    rep("O4_low", 2), rep("O4_conv",2),
                                    rep("A4_low", 2), rep("A4_conv",2),
                                    rep("O3_low", 2), rep("O3_conv",2))

post_harvest_tillage_list <- lapply(split(post_harvest_tillage, post_harvest_tillage$matrix_id),
       function(x)(rbind(cbind(matrix(c(x$X1, x$X2), byrow = FALSE, nrow=2),
                   matrix(rep(0,12), nrow = 2)),
                   matrix(rep(0,48), nrow = 6))))

```



### Off-season survival 

Multiple scenarios of off-season seed survival at the top and bottom stratum were used. Our exhaustive search in the current literature has returned few crop environments: corn, soybean, and rye cover crop. 

To account for the difference in field conditions provided by cover crops between no till with cover crop (from O4 to A4 phases) and no till without cover crop (from S2 to C2, S3 to O4, and S4 to O4 phases), educated guesses were made of overwinter seed viability were examined. 

Scenario 1: Using AMAPA's seedbank behavior in place of AMATA's.  

@sosnoskieGlyphosateResistanceDoes2013 tested *Amaranthus Palmeri* seed viability as a function of 

@sosnoskieGlyphosateResistanceDoes2013 did not find significant effect of rye cover crop's presence on overwinter viability of Palmer amaranth's seeds, but other studies have recognized the effects of cover crops on herbivore activities [@gallandtEffectCovercroppingSystems2005; @heggenstallerSeasonalPatternsPostdispersal2006]. 

$o_1 = y = 98.9 e^{-0.068x} \%$ and $o_2 = y = 99.5 e^{-0.05x} \%$ from Figures 1 and 3 in [@sosnoskieGlyphosateResistanceDoes2013] were used for the top and bottom survival rate between all crop phases.


The x values are filled with the gap in months between active growing of the preceding and succeeding crop phases, such as 6 months from corn harvest to soybean planting and 9 month from oat harvest to alfalfa active growing, etc.  

Since the raw data in Sosnoskie et al., 2013 is not available, the variance of Y at each X was visually assessed from figures 1 and 3 @sosnoskieGlyphosateResistanceDoes2013.

In a more sophisticated, but unproductive effort: 
Functions $y = ae^{-bx}$ are first linearized as $Y^* = ln(Y) = ln(a) -bx$ following the linearization formulae on page 29/125 <https://pdixon.stat.iastate.edu/stat511/notes/part%205.pdf>, and

$Var(Y^*) = Var[ln(a) - bx] = Var[ln(a)] + Var[bx] = \frac{1}{a^2}Var(a) + x^2Var(b)$ because $x$ are random numbers representing the number of months seeds are buried


Applying the delta method to find Var(Y) from Var(ln(Y)), $Var(Y) = \frac{Var[ln(Y)]}{Y^2}$^, with $Var(Y)$ being visually estimated from the error bars in figures 1 and 3 @sosnoskieGlyphosateResistanceDoes2013. Does Y need to be normally distributed for  $Var(Y) = \frac{Var[ln(Y)]}{Y^2}$ to hold?

Var of transformed function of X here <https://stats.stackexchange.com/questions/57715/expected-value-and-variance-of-loga>
<https://stats.stackexchange.com/questions/418313/variance-of-x-and-variance-of-logx-how-to-relate-them>

Scenario 2: Using AMATA undisturbed seedbank behavior

relevant for the no till

 
*Each scenario of overwinter or summer seed viability was examined by holding all the values in the empirical matrices constant.*




```{r off-season-viability-matrix-Sosnoskie, echo=FALSE}
# After seed rain and post-harvest vertical redistribution
#################### Scenario 1: Sosnoskie et al. 2013 -  No distinguishable cover crop effect
# Assign 6 months of summer to corn and soybean, 5.5 months of summer to oat and 4.5 month of summer to alfalfa per the observation at Marsden that oat and alfalfa delayed waterhemp emergence.
# By eye-balling, SE(top survival ) = 5% and SE(bottom survival) = 7%, so var_survival(top) = 0.025 and var_survival(bottom) = 0.049
 
overwinter_Sosnoskie_13 <- read_csv("../2-Data/Raw/seed-survival-Sosnoskie-2013.csv", show_col_types = FALSE)

overwinter_Sosnoskie_13_cohort <- overwinter_Sosnoskie_13 %>%
  mutate(Seed_overwinter = Coefficient_a*exp(Coefficient_b*winter_month),
         Seed_overwinter_prop = Seed_overwinter/100,
         var_overwinter_survival = ifelse(Stratum == "Top", 0.025, 0.049),
         matrix_id = paste(from_Phase,Corn_weed_management, sep = "_")) %>%
  arrange(matrix_id, desc(Stratum)) #Matrices are identified by the crop species that presented before the month of December in a particular plot.

## Save overwinter survival mean

overwinter_mean_Sosnoskie_13_cohort_list <- lapply(split(overwinter_Sosnoskie_13_cohort, overwinter_Sosnoskie_13_cohort$matrix_id),
       function(x) rbind(cbind(diag(x$Seed_overwinter_prop, 2), 
                         matrix(rep(0,12), nrow = 2)),
                   matrix(rep(0,48), nrow = 6)))
                         

# saveRDS(overwinter_mean_Sosnoskie_13_cohort_list, file="../2-Data/Clean/mean-winter-seed-survival-scenario1.RData")

## Save overwinter survival variance 

var_overwinter_Sosnoskie_13_cohort_list <- lapply(split(overwinter_Sosnoskie_13_cohort, overwinter_Sosnoskie_13_cohort$matrix_id),
       function(x) rbind(cbind(diag(x$var_overwinter_survival,2), 
                         matrix(rep(0,12), nrow = 2)),
                   matrix(rep(0,48), nrow = 6)))

# saveRDS(var_overwinter_Sosnoskie_13_cohort_list , file="../2-Data/Clean/var-winter-seed-survival-scenario1.RData")
```

```{r off-season-viability-matrix-Korres, echo=FALSE}
#################### Scenario 2: Korres et al., 2018 - No crop environment effect

overwinter_Korres_18 <-  expand.grid(Crop_ID = c("C2","S2",
                                                "C3","S3","O3",
                                                "C4", "S4", "O4", "A4"),
                                    Corn_weed_management = c("conv","low"),
                                    Stratum = c("Top", "Bottom")) 

overwinter_Korres_18_deduced <- overwinter_Korres_18 %>%
  mutate(Seed_overwinter_prop = ifelse(Stratum == "Top", 0.44, 0.31)) %>%
  mutate(from_Phase = Crop_ID,
         matrix_id = paste(from_Phase,Corn_weed_management, sep = "_")) %>%
  arrange(matrix_id, Stratum)


overwinter_Korres_18_cohort_list <- lapply(split(overwinter_Korres_18_deduced, overwinter_Korres_18_deduced$matrix_id),
       function(x)(rbind(cbind(diag(x$Seed_overwinter_prop, 2), 
                         matrix(rep(0,12), nrow = 2)),
                   matrix(rep(0,48), nrow = 6))))
                         
off_season_scenario2 <- overwinter_Korres_18_cohort_list

# saveRDS(off_season_scenario2, file="../2-Data/Clean/winter-seed-survival-scenario2.RData")


#################### Scenario 3: corn and soybean deduced from Steckel et al. 2007, oat and alfalfa from Scenario 1's numbers.  

overwinter_Steckel_07 <- read_csv("../2-Data/Raw/winter-seed-survival-Steckel-2007.csv", show_col_types = FALSE)
  
overwinter_Steckel_07_cohort <- overwinter_Steckel_07 %>%
  mutate(Seed_overwinter_prop = Seed_overwinter_pct/100,
         matrix_id = paste(from_Phase, Corn_weed_management, sep = "_")) %>%
  arrange(matrix_id, desc(Stratum))


overwinter_Steckel_07_cohort_list <- lapply(split(overwinter_Steckel_07_cohort, overwinter_Steckel_07_cohort$matrix_id),
       function(x)(rbind(cbind(diag(x$Seed_overwinter_prop, 2), 
                         matrix(rep(0,12), nrow = 2)),
                   matrix(rep(0,48), nrow = 6))))
                         
off_season_scenario3 <- overwinter_Steckel_07_cohort_list 

# saveRDS(off_season_scenario3, file="../2-Data/Clean/winter-seed-survival-scenario3.RData")

```

### Prep-planting tillage matrices  
The field cultivator mean and variance matrices were calculated using the same procedure as for chisel and moldboard plow.

```{r pre-planting-tillage-matrix-for-model}
#consider dataframe instead of matrix

field_cultivator_mean <- as.data.frame(matrix(c(0.594855305,	0.149901381,
0.405144695,	0.850098619), byrow = TRUE, nrow = 2))

field_cultivator_var <- as.data.frame(matrix(c(0.000774928,	-0.000125672,
-0.000774928,	0.000125672), byrow = TRUE, nrow = 2))

# Make a long dataframe of field cultivator and notill according to the crop id

X1 = c(rep(field_cultivator_mean[,1],16), rep(notill_mean[,1],2))
X2 = c(rep(field_cultivator_mean[,2],16), rep(notill_mean[,2],2))
matrix_id <- c(rep("C2_low", 2), rep("C2_conv",2),
                                    rep("C3_low", 2), rep("C3_conv",2),
                                    rep("C4_low", 2), rep("C4_conv",2),
                                    rep("S2_low", 2), rep("S2_conv",2),
                                    rep("S3_low", 2), rep("S3_conv",2),
                                    rep("S4_low", 2), rep("S4_conv",2),
                                    rep("O4_low", 2), rep("O4_conv",2),
                                    rep("O3_low", 2), rep("O3_conv",2),
                                    rep("A4_low", 2), rep("A4_conv",2))

mean_pre_planting_tillage_df <- data_frame(X1,X2, matrix_id)

## save mean list
mean_pre_planting_tillage_list <- lapply(split(mean_pre_planting_tillage_df, mean_pre_planting_tillage_df$matrix_id),
      function(x) rbind(cbind(matrix(c(x$X1, x$X2), byrow = FALSE, nrow=2),
                   matrix(rep(0,12), nrow = 2)),
                   matrix(rep(0,48), nrow = 6)))

#saveRDS(mean_pre_planting_tillage_list, file="../2-Data/Clean/mean-pre-planting-tillage.RData")

## save var list
V1 = c(rep(field_cultivator_var[,1],16), rep(notill_var[,1],2))
V2 = c(rep(field_cultivator_var[,2],16), rep(notill_var[,2],2))

var_pre_planting_tillage_df <- data_frame(V1,V2, matrix_id)

## save mean list
var_pre_planting_tillage_list <- lapply(split(var_pre_planting_tillage_df, var_pre_planting_tillage_df$matrix_id),
      function(x) rbind(cbind(matrix(c(x$V1, x$V2), byrow = FALSE, nrow=2),
                   matrix(rep(0,12), nrow = 2)),
                   matrix(rep(0,48), nrow = 6)))

# saveRDS(var_pre_planting_tillage_list, file="../2-Data/Clean/var-pre-planting-tillage.RData")
```

```{r pre-planting-tillage-matrix, include=FALSE}
# Pre-planting tillage

field_cultivator <- read_csv("../2-Data/Raw/Tillage-resized/fieldcultivator_2by2.csv", col_names = FALSE, show_col_types = FALSE)

pre_planting_tillage_for_model <- rbindlist(list(field_cultivator, field_cultivator , field_cultivator, field_cultivator,
                                               field_cultivator, field_cultivator , field_cultivator, field_cultivator,
                                               field_cultivator, field_cultivator , field_cultivator, field_cultivator,
                                               field_cultivator, field_cultivator , field_cultivator, field_cultivator,
                                               no_till, no_till),
                              use.names = TRUE, fill = FALSE)

# All phases received pre-planting field cultivator, except A4
pre_planting_tillage_for_model$matrix_id <- c(rep("C2_low", 2), rep("C2_conv",2),
                                    rep("C3_low", 2), rep("C3_conv",2),
                                    rep("C4_low", 2), rep("C4_conv",2),
                                    rep("S2_low", 2), rep("S2_conv",2),
                                    rep("S3_low", 2), rep("S3_conv",2),
                                    rep("S4_low", 2), rep("S4_conv",2),
                                    rep("O4_low", 2), rep("O4_conv",2),
                                    rep("O3_low", 2), rep("O3_conv",2),
                                    rep("A4_low", 2), rep("A4_conv",2))

pre_planting_tillage_list_for_model <- lapply(split(pre_planting_tillage_for_model, pre_planting_tillage_for_model$matrix_id),
      function(x)(rbind(cbind(matrix(c(x$X1, x$X2), byrow = FALSE, nrow=2),
                   matrix(rep(0,12), nrow = 2)),
                   matrix(rep(0,48), nrow = 6))))

#saveRDS(pre_planting_tillage_list_for_model, file="../2-Data/Clean/pre-planting-tillage.RData")
```


```{r pre-planting-tillage-matrix-for-recruitment-prop}
## Sequence: copy from mean_pre_planting_tillage_df

pre_planting_tillage_for_recruitment <- mean_pre_planting_tillage_df

# All phases received pre-planting field cultivator, except A4
pre_planting_tillage_for_recruitment$from_Phase <- c(rep("C2_low", 2), rep("C2_conv", 2),
                                    rep("C3_low", 2), rep("C3_conv", 2),
                                    rep("C4_low", 2), rep("C4_conv", 2),
                                    rep("S2_low", 2), rep("S2_conv", 2),
                                    rep("S3_low", 2), rep("S3_conv", 2),
                                    rep("S4_low", 2), rep("S4_conv", 2),
                                    rep("O4_low", 2), rep("O4_conv", 2),
                                    rep("O3_low", 2), rep("O3_conv", 2),
                                    rep("A4_low", 2), rep("A4_conv", 2))

pre_planting_tillage_for_recruitment$to_Phase <-c(rep("S2_low", 2), rep("S2_conv", 2),
                                                  rep("S3_low", 2), rep("S3_conv", 2),
                                                  rep("S4_low", 2), rep("S4_conv", 2),
                                                  rep("C2_low", 2), rep("C2_conv", 2),
                                                  rep("O3_low", 2), rep("O3_conv", 2),
                                                  rep("O4_low", 2), rep("O4_conv", 2),
                                                  rep("A4_low", 2), rep("A4_conv", 2),
                                                  rep("C3_low", 2), rep("C3_conv", 2),
                                                  rep("C4_low", 2), rep("C4_conv", 2))

# The matrix_id is identified by the phase that the cycle is moving from 
pre_planting_tillage_for_recruitment$matrix_id <- pre_planting_tillage_for_recruitment$from_Phase



pre_planting_tillage_list_for_recruitment <- lapply(split(pre_planting_tillage_for_recruitment, pre_planting_tillage_for_recruitment$matrix_id),
       function(x) rbind(cbind(matrix(c(x$X1, x$X2), byrow = FALSE, nrow = 2),
                   matrix(rep(0,12), nrow = 2)),
                   matrix(rep(0,48), nrow = 6)))

# saveRDS(pre_planting_tillage_list_for_recruitment, file="../2-Data/Raw/pre-planting-tillage-for-recruitment.RData")

#This list is continued in Script FLOW-recruitment-matrix-organization
```

```{r total-prob-calculation}
# Calculate total probability by assuming a seed column of 50 at the top and 50 at the bottom, then apply the relevant tillage 
seed_column_example <- matrix(50, nrow =2)


as.matrix(chisel) %*% seed_column_example

as.matrix(moldboard) %*% seed_column_example

as.matrix(field_cultivator) %*% seed_column_example
```

```{r summer-seed-survival-matrix}
# Summer seed consumption on the soil surface,  to be incorporated with recruitment matrix by summation

####################### Scenario 1: 

# Calculations are from Sosnoskie et al., 2013, using the same rationale as in the winter survival, but different values of the time lapse to account for cool-seasoned crops effects. 
# The gap month is the number of months that a crop is in active growth.  This arrangement will account for dormant seeds viability loss.

summer_seed_survival_Sosnoskie_13 <- overwinter_Sosnoskie_13 

# The same coefficients a and b were used on summer seed survival
summer_seed_survival_Sosnoskie_13_cohort <- summer_seed_survival_Sosnoskie_13  %>%
  mutate(Seed_summer_survive = Coefficient_a*exp(Coefficient_b*summer_month),
         Seed_summer_prop = Seed_summer_survive/100,
          var_summer_survival = ifelse(Stratum == "Top", 0.025, 0.049),
         matrix_id = paste(from_Phase,Corn_weed_management, sep = "_")) %>%
  arrange(matrix_id, desc(Stratum)) #Matrices are identified by the crop species that presented before the month of December in a particular plot.

## fill the bottom left 6x6 matrix's diagonal with 1's to perform element wise multiplication with female-survival in prospective-dynamics.Rmd 
mean_summer_seed_survival_Sosnoskie_13_cohort_list  <- lapply(split(summer_seed_survival_Sosnoskie_13_cohort , summer_seed_survival_Sosnoskie_13_cohort$matrix_id),
       function(x) rbind(cbind(diag(x$Seed_summer_prop, 2), 
                         matrix(rep(0,12), nrow = 2)), cbind(matrix(0,nrow = 6, ncol = 2), diag(1,6))))

## Save summer seed survival means

# saveRDS(mean_summer_seed_survival_Sosnoskie_13_cohort_list, file="../2-Data/Clean/mean-summer-seed-survival-scenario1.RData")

## fill the bottom left 6x6 matrix's diagonal with 1's to perform element wise multiplication with female-survival in prospective-dynamics.Rmd 
var_summer_seed_survival_Sosnoskie_13_cohort_list  <- lapply(split(summer_seed_survival_Sosnoskie_13_cohort , summer_seed_survival_Sosnoskie_13_cohort$matrix_id),
       function(x) rbind(cbind(diag(x$var_summer_survival, 2), 
                         matrix(rep(0,12), nrow = 2)), cbind(matrix(0,nrow = 6, ncol = 2), diag(1,6))))

## Save summer seed survival variance

# saveRDS(var_summer_seed_survival_Sosnoskie_13_cohort_list , file="../2-Data/Clean/var-summer-seed-survival-scenario1.RData")

####################### Scenario 2: Buhler and Harzler 2001  findings are actually survival after one year. These numbers might be inflated if an overwinter matrix is already in place. This matrix might serve as a combined matrix of seed survival for the whole year. 
summer_seed_survival_scenario2 <- expand.grid(Crop_ID = c("C2","S2",
                                                "C3","S3","O3",
                                                "C4", "S4", "O4", "A4"),
                                    Corn_weed_management = c("conv","low"),
                                    Stratum = c("Top", "Bottom"))

## Fill in the proportion of retrieved seeds and viability of retrieved seeds 1 year after burial ## Assuming no crop effects
summer_seed_survival_scenario2_filled <- summer_seed_survival_scenario2 %>%
  mutate(retrieved_prop = 0.65,
         number_buried = 2000, #2000 seed buried when exp started
         retrieved_viability = 0.8,
         number_retrieved = retrieved_prop*2000,
         number_viable =  retrieved_viability*number_retrieved,
       #  var_retrieved = number_buried*retrieved_prop*(1-retrieved_prop),
      #   var_viable = number_viable* retrieved_viability*(1-retrieved_viability),
         original_survival_prop = retrieved_prop*retrieved_viability,
         var_original_survival = number_buried*original_survival_prop*(1-original_survival_prop ),
         matrix_id = paste(Crop_ID, Corn_weed_management, sep = "_")) %>%
    arrange(matrix_id, desc(Stratum))

mean_summer_seed_survival_scenario2_filled_list <- lapply(split(summer_seed_survival_scenario2_filled, summer_seed_survival_scenario2_filled$matrix_id),
       function(x) rbind(cbind(diag(x$original_survival_prop, 2), 
                         matrix(rep(0,12), nrow = 2)), cbind(matrix(0,nrow = 6, ncol = 2), diag(1,6))))

# saveRDS(mean_summer_seed_survival_scenario2_filled_list, file="../2-Data/Clean/mean-summer-seed-survival-scenario2.RData")

var_summer_seed_survival_scenario2_filled_list <- lapply(split(summer_seed_survival_scenario2_filled, summer_seed_survival_scenario2_filled$matrix_id),
       function(x) rbind(cbind(diag(x$var_original_survival, 2), 
                         matrix(rep(0,12), nrow = 2)), cbind(matrix(0,nrow = 6, ncol = 2), diag(1,6))))

# saveRDS(var_summer_seed_survival_scenario2_filled_list, file="../2-Data/Clean/var-summer-seed-survival-scenario2.RData")
```

Assuming that the probability of a seed retrieved(X) is independent of its viability(Y): $Var(XY) = E(X^2)Var(Y) + E(Y^2)Var(X)$. Following this formula returned Var(viable) in the billions.  <https://stats.stackexchange.com/questions/397839/varxy-if-x-and-y-are-independent-random-variables>

Variance of a binomial variable is calculated as $Var(X_i) = np_i(1-p_i)$. So Vat(survival) = n * survival * mortal, with n being the number of buried seeds and anything that is not found can be considered mortal. It is acceptable in this case because the seeds that are not "found" would not join the tracked seedling population, and further, the mature plant population.   


```{r summer-seed-predation}
## may be optional 
summer_seed_pred <- read_csv("../2-Data/Raw/seed-predation-van-der-Laat-2015.csv", na = ".",
                             show_col_types = FALSE)

summer_seed_pred_clean <- summer_seed_pred %>%
  rename(Event = ev,
         Block = blk,
         Sample = sam,
         Consumed_pct = `%pred`) %>%
  mutate(Year = ifelse(yr == 1, "2003", "2004"),
         Species = ifelse(sp == "W", "Waterhemp", "NA"),
         Tillage = ifelse(trt =="CP", "chisel",
                          ifelse(trt == "NT", "no till", "moldboard")),
         Exclusion = ifelse(type == "w", "vertebrate exclusion",
                            ifelse(type == "c", "none", "vertebrate and invertebreate exclusion")),
         Consumed_prop = Consumed_pct/100)

#summer_seed_pred_clean[,c(1:7, 9:13)] <- lapply(summer_seed_pred_clean[,c(1:7, 9:13)], factor)

## Since the authors did not find consistent patterns of tillage effects on seed predation, I used a data-based (instead of model-based) sets of mean seed consumption for a starting point of my population dynamics model.  
summer_seed_pred_summ <- summer_seed_pred_clean %>%
  group_by(Tillage, Species, Exclusion, Year, Event) %>%
  summarize(Consumed_prop = mean(Consumed_prop, na.rm = TRUE))

## The summer predation averaged over 10 timestamps and 2 year
summer_seed_pred_grand_summ <- summer_seed_pred_clean %>%
  group_by(Tillage, Species) %>%
  summarize(Consumed_prop = mean(Consumed_prop, na.rm = TRUE)) %>%
  mutate(Stratum = "Top") #add stratum identification to use later


```

```{r seedling-survival-Hartzler-et-al}
# a corn table: from Nordby and Hartzler 2004: Table 3, 100 - Mortality for 76 cm row spacing in corn
# note: this pub has height, shoot dry wt and individual fecundity, too
# Number of plant/cohort = 30 (highlighted, pg 257)
# Assume WH cohort 1 at VE, cohorts 2 and 3 at V3, cohorts 4 and 5 at V5, cohort 6 at V8
# Var = SE^2

seedling_survival_corn <- expand.grid(Crop_ID = c("C2","C3","C4"),
                                    Corn_weed_management = c("conv","low"),
                                    Cohort = c(1:6))

seedling_survival_corn_filled <- seedling_survival_corn %>%
  mutate(mean_seedling_survival = ifelse(Cohort == 1, 0.84,
                                    ifelse(Cohort %in% c(2,3), 0.48,
                                           ifelse(Cohort %in% c(4,5), 0.04, 0.01)))) %>%
  mutate(var_seedling_survival = ifelse(Cohort == 1, 0.09,
                                    ifelse(Cohort %in% c(2,3), 0.09,
                                           ifelse(Cohort %in% c(4,5), 0.09, 0.09)))) %>%
  mutate(matrix_id = paste(Crop_ID, Corn_weed_management, sep = "_")) %>%
  arrange(matrix_id, Cohort)

# a soybean table: from Nordby and Hartzler 2004: fig 1, pg 244
# note: this pub has height, shoot dry wt and individual fecundity, too
# Number of plant/cohort = 60 (highlighted, pg 243)
# Survival (Y) = 95.4 + 0.058x - 0.034x^2 (X = days after planting)
# one cohort is 15 days after the other
# Var = SE^2
# Cohort i (i from 1 to 6): cohort 1 at x = 14, cohort i+1 at x + 14. 

seedling_survival_soybean <- expand.grid(Crop_ID = c("S2","S3","S4"),
                                    Corn_weed_management = c("conv","low"),
                                    Cohort = c(1:6))


#View(seedling_survival_soybean)


seedling_survival_soybean_filled <- seedling_survival_soybean %>%
  mutate(days_after_planting = ifelse(Cohort == 1, 14,
                                      ifelse(Cohort == 2, 28,
                                             ifelse(Cohort == 3, 42,
                                                    ifelse(Cohort == 4, 56,
                                                           ifelse(Cohort == 5, 70, 84)))))) %>%
  mutate(matrix_id = paste(Crop_ID, Corn_weed_management, sep = "_")) %>%
  arrange(matrix_id, Cohort) %>%
  mutate(const. = 95.4, # the constant variable in the y = a + bx + cx^2
         x_squared = days_after_planting^2,
         mean_seedling_survival = (const. + 0.058*days_after_planting - 0.034*(x_squared))/100 ,
         number_of_plant = 60) 

seedling_survival_soybean_filled$mean_seedling_survival<- pmax(seedling_survival_soybean_filled$mean_seedling_survival,0)
    
seedling_survival_soybean_filled <- seedling_survival_soybean_filled %>%
  mutate(var_seedling_survival = ifelse(mean_seedling_survival > 0,
                                         mean_seedling_survival*(1-mean_seedling_survival)/number_of_plant, 0))

seedling_survival_soybean_filled_select <- seedling_survival_soybean_filled%>%
  select(Crop_ID, Corn_weed_management, Cohort, mean_seedling_survival, var_seedling_survival, matrix_id)

### oat table

seedling_survival_oat <- expand.grid(Crop_ID = c("O3","O4"),
                                    Corn_weed_management = c("conv","low"),
                                    Cohort = c(1:6))


## assuming the cohort-specific waterhemp seedling survival rates in oat was reciprocal to those in corn,
## in which the earlier cohorts in oat experience lower survival rate 
seedling_survival_oat_filled <- seedling_survival_oat %>%
  mutate(mean_seedling_survival = ifelse(Cohort %in% c(1,2), 0.1,
                                         ifelse(Cohort %in% c(3,4), 0.5, 0.9)),
    var_seedling_survival = 0.09) %>%
    mutate(matrix_id = paste(Crop_ID, Corn_weed_management, sep = "_")) %>%
  arrange(matrix_id, Cohort)


### alfalfa table

## assuming the first 4 cohorts survival rate was 0.1 and the last two 0.5
## 0.1 because of frequent hay cut and shading
## 0.5 after the last hay cut, but colder and shading 
## assuming the same var(survival) in corn as in alfalfa
seedling_survival_alfalfa <- expand.grid(Crop_ID = c("A4"),
                                    Corn_weed_management = c("conv","low"),
                                    Cohort = c(1:6))

seedling_survival_alfalfa_filled <- seedling_survival_alfalfa %>%
  mutate(mean_seedling_survival = ifelse(Cohort %in% c(1, 2, 3, 4), 0.1, 0.5),
    var_seedling_survival = 0.09)%>%
    mutate(matrix_id = paste(Crop_ID, Corn_weed_management, sep = "_")) %>%
  arrange(matrix_id, Cohort)


## combine 4 crop-specific tables
seedling_survival_Hartzler_et_al <- rbind(seedling_survival_corn_filled,
                                          seedling_survival_soybean_filled_select,
                                          seedling_survival_oat_filled,
                                          seedling_survival_alfalfa_filled)

# View(seedling_survival_Hartzler_et_al)

## need to check top 2x2
mean_seedling_survival_Hartzler_et_al_raw_list <- lapply(split(seedling_survival_Hartzler_et_al, seedling_survival_Hartzler_et_al$matrix_id),
       function(x) rbind(cbind(diag(1,2), matrix(0,nrow = 2, ncol = 6)),
         cbind(matrix(0,nrow=6,ncol=2), diag(x$mean_seedling_survival))))

# saveRDS(mean_seedling_survival_Hartzler_et_al_raw_list, file="../2-Data/Clean/mean-summer-seedling-survival-scenario2.RData")

var_seedling_survival_Hartzler_et_al_raw_list <- lapply(split(seedling_survival_Hartzler_et_al, seedling_survival_Hartzler_et_al$matrix_id),
       function(x) rbind(cbind(diag(1,2), matrix(0,nrow = 2, ncol = 6)),
         cbind(matrix(0,nrow=6,ncol=2), diag(x$var_seedling_survival))))

# saveRDS(var_seedling_survival_Hartzler_et_al_raw_list, file="../2-Data/Clean/var-summer-seedling-survival-scenario2.RData")
```

```{r}
## After model run: where does herbicide survival fit in?
## Using 2018 and 2019 survival as a starting point, partition matrices into HR and non-HR survival rate, sum them up
```