---
output:
    bookdown::word_document2:
      toc: false
      fig_caption: yes
    reference_docx: style_template.docx
bibliography: prospective-pop-dynamics.bib
---


```{r, include=FALSE}
library(readr)
library(flextable)
```
The date used in this excercise were collected from empirical experiment and the literature.

### Experiment design {-}
The general experiment design and diagram were provided in [@nguyenImpactCroppingSystem2022], but an updated diagram is provided below to cover tillage regimes.  


The empirical data was organized to distinguish historical herbicide information for soybean plots because the herbicide regime change was recent while the legacy effect of weed control program on each subplot may last for a few years [@rasmussenLegacyPesticidePollution2015; @gibsonBenchmarkStudyGlyphosateresistant2016].  

*Population transition matrices were calculated using Wood's quadratic programming method [Section 6.2, @caswellMatrixPopulationModels2001].*

The waterhemp population is considered size-structured with overlapping cohorts.
Cohorts in this experiment reflected the time stamps when the waterhemp sub-populations were surveyed.
Two consecutive surveys were approximately two to three weeks apart, as the weather permitted.

### Matrix form {-}   

Each transition matrix of eight rows by eight columns (8 x 8) consists of four blocks using the format of matrix 4.8 in [Chapter 4, @caswellMatrixPopulationModels2001]. 
Each rotational transition matrix, $A_r$, is the product of two, three or four annual projection matrices, $P_c$ (c = 2, 3, or 4, corresponding to the number of crop phases in each rotation). Each annual projection matrices, corresponding to a crop environment, $P_c$, is the product of six sub-annual matrices, $B_k$ with $k = {1,2,...,6}$.s 

The general form of any matrix (periodic sub-annual ($B_k$), annual ($A_r$), or rotational ($P_c$)) is 

$$
\left[\begin{array}{c|c}
  M_{s} & M_{p\,\,to\,\,s}  \\ \hline
  M_{s\,\,to\,\,p} & M_{p} 
\end{array} \right]
$$

where,  
$M_s$, 2 x 2, is the transition within the seedbank population,   
$M_p$, 6 x 6, is the transition within the plant population,  
$M_{s\,\,to\,\,p}$, 6 x 2, is the transition from the seedbank to the plant population (emergence), and    
$M_{p\,\,to\,\,s}$, 2 x 6, is the transition from the plant to the seedbank population (distribution of newly produced seeds to the soil seedbank)   

Within a year, waterhemp population projection from time $k$ to $k+1$ was reflected by the changes in number of seeds and plants and calculated as follow [Chapters 2, 3 and 4, @caswellMatrixPopulationModels2001].  
$$N_{k+1} = B_k*N_k$$  
where,  
$\mathcal{B_k}$ is the square transition matrix from time $k$ to time $k+1$, and   
$N$ is the population vector (in column matrix form) of eight rows and one column.  
Waterhemp population transition in one crop phase, from pre-planting tillage to winter dormancy is calculated with 

$$ N_{r_{c,k+1}} = N_{r_{c,h}}*\prod_{k=1}^6 B_k  \;\;\;\;\;  [1]$$   
where,   
$r_{c}$ is the index for the $c^{th}$ crop phase in a particular rotation,   
$N_{r_{c}}$ is the Leslie's population matrix in crop phase $c^{th}$, and   [@leslieUseMatricesCertain1945]   
$B_k$ is the population transition matrix during summer    

The main factors that contribute to the success of the control practices are tillage regime, chemical herbicide, cultivation practice and crops' competitiveness. 
Among those four factors, crop competitiveness will not be factorized because this factor is dependent on other factors.  

The following parameters were collected and plugged into a general population model to examine when and where large changes in population dynamics might occur. 
The list of parameters and the corresponding formula is presented in Table 4.1 below.    

Table 4.1 - Model parameters  

| Parameter | Formula | Notation | Definition |  Reference | 
| --------- | ------------------------- | ------------------- | ----------------------------- | -------------- |   
| Common indexes | | $i=\{1,2,...6\}$ | Cohort number, 1 to 6 |   
| | | $d=\{1,2\}$ | Top (1) and bottom (2) soil strata |  
| | | $k$ | The time between two consecutive measurements |  
| | | $c$ | Crop: corn, soybean, oat, alfalfa |   
| | | $r$ | Rotation: 2-,3-,4-year |   
| | | $h$ | Corn weed management: low or conventional |  
| | | $N_{s_d}$ | Seedbank density at stratum $d^{th}$ |  
| | | $N_{p_i}$ | Cohort $i^{th}$ plant density |  
| --------- | ------------------------- | ------------------- | ----------------------------- | -------------- |   
| Seedling establishment | $$ e_{i,(r,c,h)} = \frac{N_{p_i,(r,c,h)}} {N_{s,(r,c,h)}}$$ |  | The proportion of the seedbank's top stratum to germinate to cohort $i^{th}$ |   
| Surface seedbank density | | $N_{s_1,(r,c,h)}$ | in Crop $c$, rotation $r$, herbicide $h$ |  
| Seedling density | | $N_{p_i,(r,c,h)}(k)$ | in Crop $c$, rotation $r$, herbicide $h$ |  
| --------- | ------------------------- | ------------------- | ----------------------------- | -------------- |      
| Plant survival | $$ l_{i,(r,c,h)} = \frac{N_{p_i,(r,c,h)}(k+1)} {N_{p_i,(r,c,h)}(k)}$$ |  | The proportion seedling in each cohort that survives until the end of the crop season |    
| Mature plant density | | $N_{p_i,(r,c,h)}(k+1)$ | in Crop $c$, rotation $r$, herbicide $h$ | 
| --------- | ------------------------- | ------------------- | ----------------------------- | -------------- |   
| | | $\phi$ | fecundity of a surviving plant (presumably resistant) relative to a susceptible one | 


### Model assumptions  

    

```{r assumptions, echo=FALSE}
#https://ardata-fr.github.io/flextable-book/captions-and-cross-references.html

assumptions <- read.csv("assumption-table.csv")

assumptions_ft <-flextable(assumptions)

assumptions_ft <- set_caption(assumptions_ft,
                              "Female-only population dynamics model assumptions for common waterhemp (*Amaranthus tuberculatus*)")

assumptions_ft 
```

### Parameterization {-}

#### Vertical redistribution of seeds, in pre-planting and post-harvest tillage {-}

In fall 2019, 36 soil cores were collected for seedbank densities estimation at two soil strata, 0-2cm as the top (1), and 2-20 cm depth as the bottom (2). 


The redistribution of female seeds that were produced at the end of the crop season is calculated per each crop phase under each rotation using the transition matrices corresponding to each tillage regime by @spokasSeedChaserVerticalSoil2007. No-till is represented by an identical matrix after  @cousensModelEffectsCultivation1990.    

#### Seedling recruitment {-}
Seedling recruitment was calculated as the proportion of emerged seedlings with regards to the top 0-2 cm top soil stratum: $N_{Apr\,through\,Oct\,2020}/N_{Nov\,2019}$, where $N_{in\,season\,2020}$ 

#### In-season survival {-}

#### Plant fecundity {-}

#### Overwinter survival {-} 
what assumptions, list in previous table

### Population growth rate and its variance {-}  

The variance of population growth rate is calculated with $Var(\lambda)= \sum_i^b S_b^TV_bS_b$ where $V$ is the variance-covariance matrix of each periodic matrix.   

Variance of annualized population growth rate is calculated with Taylor series expansion using the general formula: $Var[f(\lambda)] = {f'(E(\lambda))}^2 Var(\lambda)$, in which $f(\lambda)$ is the nth-root function used to annualize the rotation-wise growth rates. For example, the variance of annualized population growth rate in the 2-year rotation was $Var(\lambda_{2}) = \frac{1}{4\lambda_{2r}}Var(\lambda_{2r})$, where $\lambda_2 = \sqrt{\lambda_{2r}}$.  

 


### Elasticity of population growth rate on each lower-level parameter {-}
