---
output:
    bookdown::word_document2:
      toc: false
      fig_caption: yes
    reference_docx: style_template.docx
---


```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(purrr) # for map2()
library(pracma)
library(popbio)
library(tidyverse)
library(magrittr) 
```

```{r periodic-matrices-phase-mean-and-var, include=FALSE}
## Pre-planting tillage 
mean_spring_tillage <-  readRDS("../2-Data/Clean/mean-pre-planting-tillage.RData")

var_spring_tillage <-  readRDS("../2-Data/Clean/var-pre-planting-tillage.RData")

## Emergence 
mean_emergence_scenario1 <- readRDS("../2-Data/Clean/mean-emergence-prop-scenario1.RData")

var_emergence_scenario1 <- readRDS("../2-Data/Clean/var-emergence-prop-scenario1.RData")

## Seed survival and plant survival 
#### seed survival 

mean_summer_seed_survival_scenario1  <- readRDS("../2-Data/Clean/mean-summer-seed-survival-scenario1.RData")

var_summer_seed_survival_scenario1  <- readRDS("../2-Data/Clean/var-summer-seed-survival-scenario1.RData")

#### plant survival: literature data
#female_survival <- readRDS("../2-Data/Clean/female-survival-rate-cohort-equal.RData")

mean_female_survival <- readRDS("../2-Data/Clean/mean-summer-seedling-survival-scenario2.RData")

var_female_survival <- readRDS("../2-Data/Clean/var-summer-seedling-survival-scenario2.RData")

#### combine seed and plant survivals into one matrix with element-wise multiplication: this is not computing, but arranging using 1's as placeholders 
mean_summer_survival_scenario1 <- purrr::map2(mean_summer_seed_survival_scenario1 , mean_female_survival, `*`)

var_summer_survival_scenario1 <- purrr::map2(var_summer_seed_survival_scenario1 , var_female_survival, `*`)

## Fecundity: next chunk


## Post-harvest tillage
mean_post_harvest_tillage <- readRDS("../2-Data/Clean/mean-post-harvest-tillage.RData")

var_post_harvest_tillage <- readRDS("../2-Data/Clean/var-post-harvest-tillage.RData")

## Overwinter survival 
mean_overwinter_scenario1 <- readRDS("../2-Data/Clean/mean-winter-seed-survival-scenario1.RData")

var_overwinter_scenario1 <- readRDS("../2-Data/Clean/var-winter-seed-survival-scenario1.RData")
```


```{r phase-wise-projection-1, include=FALSE}
## Circle starts at spring tillage Fecundity scenario 1 - 2019 rates

mean_fecundity_2019 <- readRDS("../2-Data/Clean/mean-fecundity-19-cohort.RData")

var_fecundity_2019 <- readRDS("../2-Data/Clean/var-fecundity-19-cohort.RData")

scenario1_projection_by_matrix_id <- tibble::lst(mean_overwinter_scenario1,
                                                 mean_post_harvest_tillage,
                                                 mean_fecundity_2019,
                                                 mean_summer_survival_scenario1,
                                               mean_emergence_scenario1,
                                               mean_spring_tillage)

## List all periodic matrices within the same crop ID together  
scenario1_projection_by_matrix_id_transpose <- purrr::transpose(scenario1_projection_by_matrix_id )

# multiply all the periodic matrices within a crop ID 

scenario1_phase_projection <- scenario1_projection_by_matrix_id_transpose  %>%
  map(., ~{Reduce( "%*%", .)}) 

# check 
#lapply(scenario1_projection_by_matrix_id_transpose, 
#       \(x) x[[1]] %*% x[[2]] %*% x[[3]] %*% x[[4]] %*% x[[5]] %*% x[[6]])

```

```{r rotation-group, include=FALSE}
#2year, conventional
scenario1_rotation_wise_2yr_conv <- scenario1_phase_projection[c("S2_conv", "C2_conv")]

#2year, low
scenario1_rotation_wise_2yr_low <- scenario1_phase_projection[c("S2_low","C2_low")]
#3-year, conventional
scenario1_rotation_wise_3yr_conv <- scenario1_phase_projection[c("O3_conv", "S3_conv", "C3_conv")]
#View(scenario1_rotation_wise_3yr_conv)

#3-year, low
scenario1_rotation_wise_3yr_low <- scenario1_phase_projection[c("O3_low", "S3_low", "C3_low")]

scenario1_rotation_wise_4yr_conv <- scenario1_phase_projection[c("A4_conv", "O4_conv", "S4_conv", "C4_conv")]

#4-year, low
scenario1_rotation_wise_4yr_low <- scenario1_phase_projection[c("A4_low", "O4_low", "S4_low", "C4_low")]


scenario1_rotation_group <- tibble::lst(scenario1_rotation_wise_2yr_conv,
                                     scenario1_rotation_wise_2yr_low,
                                     scenario1_rotation_wise_3yr_conv,
                                     scenario1_rotation_wise_3yr_low,
                                     scenario1_rotation_wise_4yr_conv,
                                     scenario1_rotation_wise_4yr_low)

scenario1_rotation_wise_projection <- scenario1_rotation_group %>%
  map(., ~{Reduce( "%*%", .)}) 

scenario1_rotation_wise_eigen <- scenario1_rotation_wise_projection %>%
  map(., ~{eigen.analysis(.)}) 

```


```{r scenario1-rotation-lambda, echo=FALSE}
scenario1_rotation_wise_lambda <- scenario1_rotation_wise_eigen %>%
  map(., ~{head(.,1)}) 

# make a df to admend to the next chunk's df 

scenario1_rotation_wise_lambda_df <- scenario1_rotation_wise_lambda %>%
   plyr::ldply(., data.frame)
```

Using 2019 fecundity rates, waterhemp populations in all crop identity is decreasing in all rotation, fastest in the 3-year rotation (Figures \@ref(fig:scenario1-lambda-annualized)). 
```{r scenario1-lambda-annualized-df, echo=FALSE}
scenario1_annualized_lambda_2yr <- rapply(scenario1_rotation_wise_lambda[1:2],
                                             function(x){nthroot(x,2)}, how = "list")    

scenario1_annualized_lambda_3yr <- rapply(scenario1_rotation_wise_lambda[3:4],
                                             function(x){nthroot(x,3)}, how = "list")

scenario1_annualized_lambda_4yr <- rapply(scenario1_rotation_wise_lambda[5:6],
                                             function(x){nthroot(x,4)}, how = "list")
   
scenario1_annualized_lambda_df <- tibble::lst(scenario1_annualized_lambda_2yr,
                                   scenario1_annualized_lambda_3yr,
                                   scenario1_annualized_lambda_4yr) %>%
   plyr::ldply(., data.frame) %>% 
  rename(conv = "lambda1", low = "lambda1.1")

scenario1_annualized_lambda_df_long <- scenario1_annualized_lambda_df %>%
  pivot_longer(!.id, names_to = "Corn_weed_management", values_to = "annualized_lambda") 

View(scenario1_annualized_lambda_df_long)

## add the rotation-wise lambdas to the annualized lambdas 
scenario1_annualized_lambda_df_long$rot_wise_lambda <-  scenario1_rotation_wise_lambda_df$lambda1


```




```{r functions, include=FALSE}
# https://stackoverflow.com/questions/15795318/efficient-way-to-create-a-circulant-matrix-in-r

circ<-function(x) { 
    n<-length(x)
    matrix(x[matrix(1:n,n+1,n+1,byrow=T)[c(1,n:2),1:n]],n,n)
}
```

```{r circ-2yr-arrange}
### conventional 

# pick all phases in the 2yr conventional treatment and arrange in backward order, i.e., first subannual matrix in the first crop phase at the right-most (or bottom) position of the list
all_periods_2yr_conv <- scenario1_projection_by_matrix_id_transpose[c("S2_conv", "C2_conv")]
#View(all_periods_2yr_conv)

# remove the phase bounded structure in a chain, making a chain of 12 subannual matrices for the 2-year rotation
all_periods_2yr_conv_chain <- unlist(all_periods_2yr_conv, recursive = FALSE)

# make a reference grid, based on a chain starting at the spring_tillage matrix
ref_2yr_conv <- names(all_periods_2yr_conv_chain)

# create circular rotation of the first subannual matrix in each chain
circ_2yr_conv <- circ(ref_2yr_conv)

full_circ_2yr_conv <- purrr::map(1:12,~all_periods_2yr_conv_chain[circ_2yr_conv[.x,]])
# name the nested list by the last element in each nested list
# the last element in each nested list appear in the same order as circ_2yr_conv[,12]

names(full_circ_2yr_conv) <- paste0(circ_2yr_conv[,12],"_contribution") 

# check if nested lists names matches their last elements' names: YES
# View( full_circ_2yr_conv) check

# remove the last subannual matrix from the chain in full_circ_2yr_conv
until_circ_2yr_conv <- full_circ_2yr_conv  %>%
  map(., ~{head(.,11)})
# until_circ_2yr_conv contains 12 chains of 11 matrices each. In term of contribution to changes in lambda, each chain starts at the subannual matrix following the matrix of interest. The `until` and `though` designation is detailed in Caswell and Trevisan 1994. 
### low

# pick all phases in the 2yr conventional treatment
all_periods_2yr_low <- scenario1_projection_by_matrix_id_transpose[c("S2_low", "C2_low")]
#View(all_periods_2yr_conv)

# rearrange the phases in a chain
all_periods_2yr_low_chain <- unlist(all_periods_2yr_low, recursive = FALSE)

# make a reference grid 
ref_2yr_low <- names(all_periods_2yr_low_chain)

circ_2yr_low <- circ(ref_2yr_low)

full_circ_2yr_low <- purrr::map(1:12,~all_periods_2yr_low_chain[circ_2yr_low[.x,]])
# name the nested list by the last element in each nested list
# the last element in each nested list appear in the same order as circ_2yr_low[,12]

names(full_circ_2yr_low) <- paste0(circ_2yr_low[,12],"_contribution") 

# check if nested lists names matches their last elements' names: YES
# View( full_circ_2yr_low): each chain's product is identified by the right-most matrix 

# remove the last subannual matrix from the chain in full_circ_2yr_low
until_circ_2yr_low <- full_circ_2yr_low %>%
  map(., ~{head(.,11)})
```

In all LTRE procedures presented here, the conventional corn weed management treatment is the reference treatment and low herbicide the treatment of interest. Following @caswellSensitivityAnalysisPeriodic1994's notions:   

The sensitivity of $\lambda$ to each element of each sub-annual periodic matrix is calculated with  

$$
S_{B(h)} = {(B_{(h-1)}...B_{(1)}...B_{(h+1)})}^TS_{A(h)} = D^T S_{A(h)} \,\,\,(1)
$$
where,  
$B_{(h)}$ is the periodic projection matrix for sub-annual period h; h = {1,...,6},  
$D^T$ is the transpose of the matrix product of all the $B_{(h)}$, and   
$S_{A(h)}$ is the sensitivity of $\lambda$ to each element of $A^*_{(h)}$ (the average annual projection matrix between the reference treatment and the treatment of interest)  

The elasticities of $\lambda$ to each element of a sub-annual projection matrix is calculated with  
$$
E_{B(h)} = \frac{b_{ijk}}{\lambda} S_{B(h)} \,\,\,(2)
$$
where,  

$b_{ijk}$ is the entry at row i column j of matrix $B_{(h)}$, and other elements as defined in Equation (1).  

```{r LTRE-2yr}
## In all LTRE procedures, the conventional treatment is the reference matrix and low the treatment of interest

# left matrix multiplication of all chains (as defined by starting point, aka the right-most matrix of the chain) THROUGH the right-most matrix --> matrix A's

# The product of each list in the full_circ_2yr_xx is A(hk),
# A(hk) is the product of all sub-annual period starting at phase k, period h
A2_conv_contribution <- full_circ_2yr_conv %>%  
  map(., ~{Reduce( "%*%", .)}) 

A2_low_contribution <- full_circ_2yr_low %>% 
  map(., ~{Reduce( "%*%", .)}) 

A2_contribution <- lst(A2_conv_contribution, A2_low_contribution)

# calculate the A* matrix, average of the two products representing conventional and low corn weed management treatments 

A2_avg_contribution <- lapply(Reduce(\(...)Map('+', ...), A2_contribution), '/', length(A2_contribution))

# Matrix S_A(h)'s: sensitivity of lambda to each element of A(h), i.e. a(ijh), using matrix A*
A2_avg_sens <- A2_avg_contribution %>%
  map(., ~{sensitivity(.)})

# Matrix D: the product of all sub-annual matrices in backward order, until the B(h) matrix appear. 
# other words: left matrix multiplication of all chains (as defined by starting point, B(h), UNTIL the B(h) appear, the chain that makes up D is one matrix fewer than one that makes S

D2_conv_contribution <- until_circ_2yr_conv %>% 
  map(., ~{Reduce( "%*%", .)}) 

D2_low_contribution <- until_circ_2yr_low %>% 
  map(., ~{Reduce( "%*%", .)}) 

D2_contribution <- lst(D2_conv_contribution, D2_low_contribution)

### Do I really need D_avg? ### Yes

D2_avg_contribution <- lapply(Reduce(\(...)Map('+', ...), D2_contribution), '/', length(D2_contribution))

D2_avg_contribution_transposed <- D2_avg_contribution %>%
  map(., ~{t(.)})

# Matrix S_B(h):  sensitivity of lambda to changes in each element of the subannual matrix B(h)  
S_B_contribution_2yr <- purrr::map2(D2_avg_contribution_transposed, A2_avg_sens, `%*%`)

# remove "conv" in all names, conv was picked up from D2_conv_contribution because it was listed first in D2_contribution. 

## pick out the names
contribution_2yr_raw_names <- names( S_B_contribution_2yr)

## remove "conv" in all names
contribution_2yr_names <- gsub('conv.', '', contribution_2yr_raw_names)

names(S_B_contribution_2yr) <- contribution_2yr_names
```

In the corn phase of the 2-year rotation, the mobility of seeds at the bottom stratum to seedling cohort 2 was the highest contributor to changes in the population growth ($\lambda$); to the seedling cohorts 1, 3 and 4 were immediate contributors to changes in the population growth. The same pattern was observed for the top-stratum seeds in the corn phase of the 2-year rotation, but their contribution to changes in $\lambda$ were lower than those of the bottom-stratum seeds. 

In the soybean phase of the 2-year rotation, emergence minimally contributed to changes in population growth. 

*need a table, as graphs will be too many*


In 2019, as $\lambda$'s were under 1 for all treatments, fecundity did not contribute substantially to $\lambda$ changes. 

```{r S-B-spring-tillage-contribution-2yr, fig.cap="Contribution of each entry element in the spring tillage sub-annual matrix to lambda.", include=FALSE}
## subletting removes names
S_B_contribution_2yr_spring_tillage <- S_B_contribution_2yr[c("C2_mean_spring_tillage_contribution", "S2_mean_spring_tillage_contribution")]

# https://stackoverflow.com/questions/62949713/unlisting-nested-lists-and-plotting-using-ggplot
# melt the list will return Var1 = row 1 in the original matrix, Var = 2 column in the original matrix, and L1 = list name in the original matrix 
reshape2::melt(S_B_contribution_2yr_spring_tillage) %>%
 rename(original_matrix_row = Var1, original_matrix_col = Var2) %>%
  filter(original_matrix_col < 3) %>%
  ggplot(aes(x =  original_matrix_row, y = value, color = as.factor(original_matrix_col))) +
  #geom_point() +
  geom_jitter() +
  facet_grid(~ L1) +
  xlab("to seed stratum or plant group") +
  labs(colour = "from seed stratum or plant group") + 
  ylab("contribution to lambda sensitivity") +
  scale_color_discrete(breaks=c("1", "2"),
                           labels=c("Top", "Bottom")) +
  scale_x_discrete(limits=c("top st. seed","bottom  st. seed","seedling co. 1", "seedling co. 2",
                            "seedling co. 3", "seedling co. 4", "seedling co. 5", "seedling co. 6")) +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.position="top")
```

```{r S-B-emerge-contribution-2yr, fig.cap="Contribution of each entry element in the emergence sub-annual matrix to lambda.", include=FALSE}
## subletting removes names
S_B_contribution_2yr_emerge <- S_B_contribution_2yr[c("C2_mean_emergence_scenario1_contribution", "S2_mean_emergence_scenario1_contribution")]

# https://stackoverflow.com/questions/62949713/unlisting-nested-lists-and-plotting-using-ggplot
# melt the list will return Var1 = row 1 in the original matrix, Var = 2 column in the original matrix, and L1 = list name in the original matrix 
reshape2::melt(S_B_contribution_2yr_emerge) %>%
 rename(original_matrix_row = Var1, original_matrix_col = Var2) %>%
  filter(original_matrix_col < 3) %>%
  ggplot(aes(x =  original_matrix_row, y = value, color = as.factor(original_matrix_col))) +
  #geom_point() +
  geom_jitter() +
  facet_grid(~ L1) +
  xlab("to seed stratum or plant group") +
  labs(colour = "from seed stratum or plant group") + 
  ylab("contribution to lambda sensitivity") +
  scale_color_discrete(breaks=c("1", "2"),
                           labels=c("Top", "Bottom")) +
  scale_x_discrete(limits=c("top st. seed","bottom  st. seed","seedling co. 1", "seedling co. 2",
                            "seedling co. 3", "seedling co. 4", "seedling co. 5", "seedling co. 6")) +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.position="top")
```

```{r S-B-summer-survival-contribution-2yr, fig.cap="Contribution of each entry element in the emergence sub-annual matrix to lambda.", include=FALSE}
## subletting removes names
S_B_contribution_2yr_summer_survival <- S_B_contribution_2yr[c("C2_mean_summer_survival_scenario1_contribution", "S2_mean_summer_survival_scenario1_contribution")]

# https://stackoverflow.com/questions/62949713/unlisting-nested-lists-and-plotting-using-ggplot
# melt the list will return Var1 = row 1 in the original matrix, Var = 2 column in the original matrix, and L1 = list name in the original matrix 
reshape2::melt(S_B_contribution_2yr_summer_survival) %>%
 rename(original_matrix_row = Var1, original_matrix_col = Var2) %>%
  filter(original_matrix_col < 3) %>%
  ggplot(aes(x =  original_matrix_row, y = value, color = as.factor(original_matrix_col))) +
  #geom_point() +
  geom_jitter() +
  facet_grid(~ L1) +
  xlab("to seed stratum or plant group") +
  labs(colour = "from seed stratum or plant group") + 
  ylab("contribution to lambda sensitivity") +
  scale_color_discrete(breaks=c("1", "2"),
                           labels=c("Top", "Bottom")) +
  scale_x_discrete(limits=c("top st. seed","bottom  st. seed","seedling co. 1", "seedling co. 2",
                            "seedling co. 3", "seedling co. 4", "seedling co. 5", "seedling co. 6")) +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.position="top")
```

```{r S-B-post-harvest-contribution-2yr, fig.cap="Contribution of each entry element in the post harvest tillage sub-annual matrix to lambda.", include = FALSE}
S_B_contribution_2yr_post_harvest_tillage <- S_B_contribution_2yr[c("C2_mean_post_harvest_tillage_contribution", "S2_mean_post_harvest_tillage_contribution")]

# https://stackoverflow.com/questions/62949713/unlisting-nested-lists-and-plotting-using-ggplot
# melt the list will return Var1 = row 1 in the original matrix, Var = 2 column in the original matrix, and L1 = list name in the original matrix 
reshape2::melt(S_B_contribution_2yr_post_harvest_tillage) %>%
  rename(original_matrix_row = Var1, original_matrix_col = Var2) %>%
  filter(original_matrix_col < 3) %>%
  ggplot(aes(x =  original_matrix_row, y = value, color = as.factor(original_matrix_col))) +
  #geom_point() +
  geom_jitter() +
  facet_grid(~ L1) +
  xlab("to seed or plant group") +
  labs(colour = "from seed stratum or plant group") + 
  ylab("contribution to lambda sensitivity") +
  scale_color_discrete(breaks=c("1", "2"),
                           labels=c("Top", "Bottom")) +
  scale_x_discrete(limits=c("top st. seed","bottom st. seed","plant co. 1", "plant co. 2",
                            "plant co. 3", "plant co. 4", "plant co. 5", "plant co. 6")) +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.position="top")

```

```{r S-B-overwinter-contribution-2yr, fig.cap="Contribution of each entry element in the overwinter sub-annual matrix to lambda.", include=FALSE}
## subletting removes names
S_B_contribution_2yr_overwinter <- S_B_contribution_2yr[c("C2_mean_overwinter_scenario1_contribution", "S2_mean_overwinter_scenario1_contribution")]

# https://stackoverflow.com/questions/62949713/unlisting-nested-lists-and-plotting-using-ggplot
# melt the list will return Var1 = row 1 in the original matrix, Var = 2 column in the original matrix, and L1 = list name in the original matrix 
reshape2::melt(S_B_contribution_2yr_overwinter) %>%
 rename(original_matrix_row = Var1, original_matrix_col = Var2) %>%
  filter(original_matrix_col < 3) %>%
  ggplot(aes(x =  original_matrix_row, y = value, color = as.factor(original_matrix_col))) +
  #geom_point() +
  geom_jitter() +
  facet_grid(~ L1) +
  xlab("to seed stratum or plant group") +
  labs(colour = "from seed stratum or plant group") + 
  ylab("contribution to lambda sensitivity") +
  scale_color_discrete(breaks=c("1", "2"),
                           labels=c("Top", "Bottom")) +
  scale_x_discrete(limits=c("top st. seed","bottom  st. seed","seedling co. 1", "seedling co. 2",
                            "seedling co. 3", "seedling co. 4", "seedling co. 5", "seedling co. 6")) +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.position="top")
```


The variance of population growth rate is calculated with $Var(\lambda)= \sum_i^b S_b^TV_bS_b$ where $V$ is the variance-covariance matrix of each periodic matrix.  

```{r var-list}

## Combine all variance lists into one grand list, similar to the means list

## Variance list needed for each rotation, but in the oposit order as in the mean list, to accommodate the order of the S_B list

var_scenario1_projection_by_matrix_id <- tibble::lst(var_spring_tillage,
                                                     var_emergence_scenario1,
                                                      var_summer_survival_scenario1,
                                                       var_fecundity_2019,
                                                        var_post_harvest_tillage,
                                                     var_overwinter_scenario1)
```

```{r var-lambda-2yr}
## The list of S_B for each subannual matrix in the 2-year rotation is in S_B_contribution_2yr

## List Variance of each subannual matrix in the 2-year rotation
var_scenario1_projection_2yr_conv <- var_scenario1_projection_by_matrix_id %>%
  map(., ~{.[c(3,13)]})

var_scenario1_projection_2yr_low <- var_scenario1_projection_by_matrix_id %>%
  map(., ~{.[c(4,14)]})

## Transpose the 2yr var list
var_scenario1_projection_2yr_conv_transpose <- purrr::transpose(var_scenario1_projection_2yr_conv)

var_scenario1_projection_2yr_low_transpose <- purrr::transpose(var_scenario1_projection_2yr_low)

## Remove nesting factor of the var list

var_scenario1_projection_2yr_conv_unlist <- unlist(var_scenario1_projection_2yr_conv_transpose, recursive = FALSE)

var_scenario1_projection_2yr_low_unlist <- unlist(var_scenario1_projection_2yr_low_transpose, recursive = FALSE)

## Keep the first column in the S_B list

S_B_contribution_2yr_first <- lapply(S_B_contribution_2yr, function(x) x[,1])
  

## Transpose of each S_B_first_col matrix, named _transposed_each to distinguish with transposing the whole list
S_B_contribution_2yr_first_transposed_each <- lapply(S_B_contribution_2yr_first, function(x) t(x))

## List t(S_B), V, and S_B in a nested list 

var_elements_2yr_lambda_list <- tibble::lst(S_B_contribution_2yr_first_transposed_each,
                               var_scenario1_projection_2yr_conv_unlist,
                               S_B_contribution_2yr_first)

View(var_elements_2yr_lambda_list)

## Multiply t(S_B), V, and S_B 

var_2yr_conv_lambda  <- Map(\(x, y, z) Reduce(`%*%`,
                      list(x, y, z)),
    S_B_contribution_2yr_first_transposed_each,
    var_scenario1_projection_2yr_conv_unlist,
    S_B_contribution_2yr_first)

var_2yr_low_lambda  <- Map(\(x, y, z) Reduce(`%*%`,
                      list(x, y, z)),
    S_B_contribution_2yr_first_transposed_each,
    var_scenario1_projection_2yr_low_unlist,
    S_B_contribution_2yr_first)
```

```{r elasticity-2yr}
# lambda atscenario1_annualized_lambda_df_long$annualized_lambd
lambda_2yr_conv <-scenario1_annualized_lambda_df_long$annualized_lambda[1]

lambda_2yr_low <-scenario1_annualized_lambda_df_long$annualized_lambda[2]
# Elasticity contribution left: check if all_periods_2yr_conv_chain is the appropriate element 
# The right-most matrix of all_periods_2yr_conv_chain is C2_spring_tillage
# why is the order of "all_periods_2yr_conv_chain" and "sens_contribution_2yr" opposite?

# Elasticity contribution left component: b_hij_2/lambda, or _3, _4 in the later chunks 
b_hij_2_over_lambda_conv <-  all_periods_2yr_conv_chain %>%
  map(., ~{./lambda_2yr_conv})

b_hij_2_over_lambda_low <-  all_periods_2yr_low_chain %>%
  map(., ~{./lambda_2yr_low})

# Elasticity contribution right component: S_B's 

# Multiply the left and right components 
E_B_contribution_2yr_conv <- Map(\(x, y) Reduce(`%*%`, list(x, y)), b_hij_2_over_lambda_conv, S_B_contribution_2yr)


E_B_contribution_2yr_low <- Map(\(x, y) Reduce(`%*%`, list(x, y)), b_hij_2_over_lambda_low, S_B_contribution_2yr)
```



```{r circ-3yr-arrange}
### conventional 

# pick all phases in the 3yr conventional treatment
all_periods_3yr_conv <- scenario1_projection_by_matrix_id_transpose[c("O3_conv", "S3_conv", "C3_conv")]
#View(all_periods_3yr_conv)

# rearrange the phases in a chain
all_periods_3yr_conv_chain <- unlist(all_periods_3yr_conv, recursive = FALSE)

# make a reference grid 
ref_3yr_conv <- names(all_periods_3yr_conv_chain)

circ_3yr_conv <- circ(ref_3yr_conv)

full_circ_3yr_conv <- purrr::map(1:18,~all_periods_3yr_conv_chain[circ_3yr_conv[.x,]])
# name the nested list by the last element in each nested list
# the last element in each nested list appear in the same order as circ_3yr_conv[,12]

names(full_circ_3yr_conv) <- paste0(circ_3yr_conv[,18],"_contribution") 

# check if nested lists names matches their last elements' names: YES
# View( full_circ_2yr_conv) check

# remove the last subannual matrix from the chain in full_circ_3yr_conv
until_circ_3yr_conv <- full_circ_3yr_conv  %>%
  map(., ~{head(.,11)})

### low

# pick all phases in the 3yr low treatment
all_periods_3yr_low <- scenario1_projection_by_matrix_id_transpose[c("O3_low", "S3_low", "C3_low")]
#View(all_periods_2yr_conv)

# rearrange the phases in a chain
all_periods_3yr_low_chain <- unlist(all_periods_3yr_low, recursive = FALSE)

# make a reference grid 
ref_3yr_low <- names(all_periods_3yr_low_chain)

circ_3yr_low <- circ(ref_3yr_low)

full_circ_3yr_low <- purrr::map(1:18,~all_periods_3yr_low_chain[circ_3yr_low[.x,]])
# name the nested list by the last element in each nested list
# the last element in each nested list appear in the same order as circ_3yr_low[,12]

names(full_circ_3yr_low) <- paste0(circ_3yr_low[,18],"_contribution") 

# check if nested lists names matches their last elements' names: YES
# View( full_circ_3yr_low) 

# remove the last subannual matrix from the chain in full_circ_3yr_low
until_circ_3yr_low <- full_circ_3yr_low %>%
  map(., ~{head(.,11)})
```

```{r LTRE-3yr}
## In all LTRE procedures, the conventional treatment is the reference matrix and low the

# left matrix multiplication of all chains (as defined by starting point, aka the right-most matrix of the chain) THROUGH the right-most matrix --> matrix A's
A3_conv_contribution <- full_circ_3yr_conv %>% 
  map(., ~{Reduce( "%*%", .)}) 

A3_low_contribution <- full_circ_3yr_low %>% 
  map(., ~{Reduce( "%*%", .)}) 

A3_contribution <- lst(A3_conv_contribution, A3_low_contribution)

A3_avg_contribution <- lapply(Reduce(\(...)Map('+', ...), A3_contribution), '/', length(A3_contribution))

# Sensitivities with regards to the right-most matrix --> matrix S_a's
A3_avg_sens <- A3_avg_contribution %>%
  map(., ~{sensitivity(.)})

# left matrix multiplication of all chains (as defined by starting point, aka the right-most matrix of the chain) UNTIL the right-most matrix appear --> matrix D's

D3_conv_contribution <- until_circ_3yr_conv %>% 
  map(., ~{Reduce( "%*%", .)}) 

D3_low_contribution <- until_circ_3yr_low %>% 
  map(., ~{Reduce( "%*%", .)}) 

D3_contribution <- lst(D3_conv_contribution, D3_low_contribution)

D3_avg_contribution <- lapply(Reduce(\(...)Map('+', ...), D3_contribution), '/', length(D3_contribution))

S_B_contribution_3yr <- purrr::map2(D3_avg_contribution, A3_avg_sens, `%*%`)

# remove "conv" in all names
contribution_3yr_raw_names <- names(S_B_contribution_3yr )

contribution_3yr_names <- gsub('conv.', '', contribution_3yr_raw_names)

names(S_B_contribution_3yr) <- contribution_3yr_names
```

```{r var-lambda-3yr}
## The list of S_B for each subannual matrix in the 2-year rotation is in S_B_contribution_2yr

## List Variance of each subannual matrix in the 2-year rotation
var_scenario1_projection_3yr_conv <- var_scenario1_projection_by_matrix_id %>%
  map(., ~{.[c(5, 15, 9)]}) # arranged by sequence C - S - O

var_scenario1_projection_3yr_low <- var_scenario1_projection_by_matrix_id %>%
  map(., ~{.[c(6, 16, 10)]})

## Transpose the 3yr var list
var_scenario1_projection_3yr_conv_transpose <- purrr::transpose(var_scenario1_projection_3yr_conv)

var_scenario1_projection_3yr_low_transpose <- purrr::transpose(var_scenario1_projection_3yr_low)

## Remove nesting factor of the var list

var_scenario1_projection_3yr_conv_unlist <- unlist(var_scenario1_projection_3yr_conv_transpose, recursive = FALSE)

var_scenario1_projection_3yr_low_unlist <- unlist(var_scenario1_projection_3yr_low_transpose, recursive = FALSE)

## Keep the first column in the S_B list

S_B_contribution_3yr_first <- lapply(S_B_contribution_3yr, function(x) x[,1])
  

## Transpose of each S_B_first_col matrix, named _transposed_each to distinguish with transposing the whole list
S_B_contribution_3yr_first_transposed_each <- lapply(S_B_contribution_3yr_first, function(x) t(x))

## List t(S_B), V, and S_B in a nested list 

var_elements_3yr_lambda_list <- tibble::lst(S_B_contribution_3yr_first_transposed_each,
                               var_scenario1_projection_3yr_conv_unlist,
                               S_B_contribution_3yr_first)

# View(var_elements_3yr_lambda_list)

## Multiply t(S_B), V, and S_B 

var_3yr_conv_lambda  <- Map(\(x, y, z) Reduce(`%*%`,
                      list(x, y, z)),
    S_B_contribution_3yr_first_transposed_each,
    var_scenario1_projection_3yr_conv_unlist,
    S_B_contribution_3yr_first)

var_3yr_low_lambda  <- Map(\(x, y, z) Reduce(`%*%`,
                      list(x, y, z)),
    S_B_contribution_3yr_first_transposed_each,
    var_scenario1_projection_3yr_low_unlist,
    S_B_contribution_3yr_first)
```

```{r elasticity-3yr}
# lambda at scenario1_rotation_wise_lambda 
lambda_3yr_conv <- scenario1_annualized_lambda_df_long$annualized_lambda[3]

lambda_3yr_low <- scenario1_annualized_lambda_df_long$annualized_lambda[4]


# Elasticity contribution left component: b_hij_3/lambda,
b_hij_3_over_lambda_conv <-  all_periods_3yr_conv_chain %>%
  map(., ~{./lambda_3yr_conv})

b_hij_3_over_lambda_low <-  all_periods_3yr_low_chain %>%
  map(., ~{./lambda_3yr_low})

# Elasticity contribution right component: S_B's 

# Multiply the left and right components 
E_B_contribution_3yr_conv <- Map(\(x, y) Reduce(`%*%`, list(x, y)), b_hij_3_over_lambda_conv, S_B_contribution_3yr)


E_B_contribution_3yr_low <- Map(\(x, y) Reduce(`%*%`, list(x, y)), b_hij_3_over_lambda_low, S_B_contribution_3yr)
```



```{r circ-4yr-arrange}
### conventional 

# pick all phases in the 4yr conventional treatment
all_periods_4yr_conv <- scenario1_projection_by_matrix_id_transpose[c("A4_conv", "O4_conv", "S4_conv", "C4_conv")]
# View(all_periods_4yr_conv)

# rearrange the phases in a chain
all_periods_4yr_conv_chain <- unlist(all_periods_4yr_conv, recursive = FALSE)

# make a reference grid 
ref_4yr_conv <- names(all_periods_4yr_conv_chain)

circ_4yr_conv <- circ(ref_4yr_conv)

full_circ_4yr_conv <- purrr::map(1:24,~all_periods_4yr_conv_chain[circ_4yr_conv[.x,]])
# name the nested list by the last element in each nested list
# the last element in each nested list appear in the same order as circ_4yr_conv[,12]

names(full_circ_4yr_conv) <- paste0(circ_4yr_conv[,24],"_contribution") 

# check if nested lists names matches their last elements' names: YES
# View( full_circ_4yr_conv) check

# remove the last subannual matrix from the chain in full_circ_3yr_conv
until_circ_4yr_conv <- full_circ_4yr_conv  %>%
  map(., ~{head(.,11)})

### low

# pick all phases in the 3yr low treatment
all_periods_4yr_low <- scenario1_projection_by_matrix_id_transpose[c("A4_low", "O4_low", "S4_low", "C4_low")]
#View(all_periods_4yr_conv)

# rearrange the phases in a chain
all_periods_4yr_low_chain <- unlist(all_periods_4yr_low, recursive = FALSE)

# make a reference grid 
ref_4yr_low <- names(all_periods_4yr_low_chain)

circ_4yr_low <- circ(ref_4yr_low)

full_circ_4yr_low <- purrr::map(1:24,~all_periods_4yr_low_chain[circ_4yr_low[.x,]])
# name the nested list by the last element in each nested list
# the last element in each nested list appear in the same order as circ_4yr_low[,12]

names(full_circ_4yr_low) <- paste0(circ_4yr_low[,24],"_contribution") 

# check if nested lists names matches their last elements' names: YES
# View( full_circ_4yr_low) 

# remove the last subannual matrix from the chain in full_circ_4yr_low
until_circ_4yr_low <- full_circ_4yr_low %>%
  map(., ~{head(.,11)})
```


```{r LTRE-4yr}
## In all LTRE procedures, the conventional treatment is the reference matrix and low the

# left matrix multiplication of all chains (as defined by starting point, aka the right-most matrix of the chain) THROUGH the right-most matrix --> matrix A's
A4_conv_contribution <- full_circ_4yr_conv %>% 
  map(., ~{Reduce( "%*%", .)}) 

A4_low_contribution <- full_circ_4yr_low %>% 
  map(., ~{Reduce( "%*%", .)}) 

A4_contribution <- lst(A4_conv_contribution, A4_low_contribution)

A4_avg_contribution <- lapply(Reduce(\(...)Map('+', ...), A4_contribution), '/', length(A4_contribution))

# Sensitivities with regards to the right-most matrix --> matrix S_A's
A4_avg_sens <- A4_avg_contribution %>%
  map(., ~{sensitivity(.)})

# left matrix multiplication of all chains (as defined by starting point, aka the right-most matrix of the chain) UNTIL the right-most matrix appear --> matrix D's

D4_conv_contribution <- until_circ_4yr_conv %>% 
  map(., ~{Reduce( "%*%", .)}) 

D4_low_contribution <- until_circ_4yr_low %>% 
  map(., ~{Reduce( "%*%", .)}) 

D4_contribution <- lst(D4_conv_contribution, D4_low_contribution)

D4_avg_contribution <- lapply(Reduce(\(...)Map('+', ...), D4_contribution), '/', length(D4_contribution))

D4_avg_contribution_transposed <- D4_avg_contribution %>% map(.,~{t(.)})

## contribution of each sub-annual matrix to lambda's sensitivities, matrix S_B's                                     
S_B_contribution_4yr <- purrr::map2(D4_avg_contribution_transposed, A4_avg_sens, `%*%`)
# remove "conv" in all names
S_B_contribution_4yr_raw_names <- names( S_B_contribution_4yr )

S_B_contribution_4yr_names <- gsub('conv.', '', S_B_contribution_4yr_raw_names)

names(S_B_contribution_4yr) <- S_B_contribution_4yr_names
```

```{r var-lambda-4yr}
## The list of S_B for each subannual matrix in the 4-year rotation is in S_B_contribution_2yr

## List Variance of each subannual matrix in the 4-year rotation
var_scenario1_projection_4yr_conv <- var_scenario1_projection_by_matrix_id %>%
  map(., ~{.[c(7, 17, 11, 1)]}) # arranged by sequence C - S - O - A

var_scenario1_projection_4yr_low <- var_scenario1_projection_by_matrix_id %>%
  map(., ~{.[c(8, 18, 12, 2)]})

## Transpose the 4yr var list
var_scenario1_projection_4yr_conv_transpose <- purrr::transpose(var_scenario1_projection_4yr_conv)

var_scenario1_projection_4yr_low_transpose <- purrr::transpose(var_scenario1_projection_4yr_low)

## Remove nesting factor of the var list

var_scenario1_projection_4yr_conv_unlist <- unlist(var_scenario1_projection_4yr_conv_transpose, recursive = FALSE)

var_scenario1_projection_4yr_low_unlist <- unlist(var_scenario1_projection_4yr_low_transpose, recursive = FALSE)

## Keep the first column in the S_B list

S_B_contribution_4yr_first <- lapply(S_B_contribution_4yr, function(x) x[,1])
  

## Transpose of each S_B_first_col matrix, named _transposed_each to distinguish with transposing the whole list
S_B_contribution_4yr_first_transposed_each <- lapply(S_B_contribution_4yr_first, function(x) t(x))

## List t(S_B), V, and S_B in a nested list 

var_elements_4yr_lambda_list <- tibble::lst(S_B_contribution_4yr_first_transposed_each,
                               var_scenario1_projection_4yr_conv_unlist,
                               S_B_contribution_4yr_first)

# View(var_elements_4yr_lambda_list)

## Multiply t(S_B), V, and S_B 

var_4yr_conv_lambda  <- Map(\(x, y, z) Reduce(`%*%`,
                      list(x, y, z)),
    S_B_contribution_4yr_first_transposed_each,
    var_scenario1_projection_4yr_conv_unlist,
    S_B_contribution_4yr_first)

var_4yr_low_lambda  <- Map(\(x, y, z) Reduce(`%*%`,
                      list(x, y, z)),
    S_B_contribution_4yr_first_transposed_each,
    var_scenario1_projection_4yr_low_unlist,
    S_B_contribution_4yr_first)
```

```{r elasticity-4yr}
## calculate elasticity, see p. 121

## NEED NEW NAMES FOR ALL THE LIST
# lambda at scenario1_rotation_wise_lambda 
lambda_4yr_conv <-scenario1_annualized_lambda_df_long$annualized_lambda[5]

lambda_4yr_low <-scenario1_annualized_lambda_df_long$annualized_lambda[6]
# Elasticity contribution left: check if all_periods_4yr_conv_chain is the appropriate element 
# The right-most matrix of all_periods_4yr_conv_chain is C4_spring_tillage
# why is the order of "all_periods_4yr_conv_chain" and "sens_contribution_4yr" opposite?


A4_conv_elas_left <-  all_periods_4yr_conv_chain %>%
  map(., ~{./lambda_4yr_conv})

A4_low_elas_left <-  all_periods_4yr_low_chain %>%
  map(., ~{./lambda_4yr_low})

# Elasticity contribution right component: S_B's sens_contribution_4yr

# Elasticity
elas_contribution_4yr_conv <- Map(\(x, y) Reduce(`%*%`, list(x, y)), A4_conv_elas_left, S_B_contribution_4yr)

```

Variance of annualized population growth rate is calculated with Taylor series expansion on the relevant moment <https://en.wikipedia.org/wiki/Taylor_expansions_for_the_moments_of_functions_of_random_variables> .

```{r scenario1-lambda-annualized, echo=FALSE, fig.cap= "Annualized waterhemp population growth rate using 2019 fecundity rate."}
# move behind variance calculation
scenario1_annualized_lambda_and_var <- scenario1_annualized_lambda_df_long %>%
  mutate(squared_exp_annualized_lambda = annualized_lambda^2)

#View(scenario1_annualized_lambda_and_var)
x_labs <- c("2-year", "3-year", "4-year")

ggplot(scenario1_annualized_lambda_and_var, aes(y = annualized_lambda, x = .id , fill = Corn_weed_management)) + 
  geom_bar(stat = "identity", position = "dodge") +
  ylab("Annualized population growth rate with 2019 fecundity") +
  xlab("Rotation") +
  scale_x_discrete(labels= x_labs) +
  theme_bw()
```
