---
title: "Q1-fecund-prep"
output:
    bookdown::word_document2:
      toc: false
      fig_caption: yes
bibliography: WH-pop-dynamics.bib
csl: apa-no-ampersand.csl 
---

We pooled all the data points in @nguyenImpactCroppingSystem2022 into one regression of ln(individual fecundity +1) against ln(individual aboveground mass + 0.005) and estimated twenty-four means of fecundity on natural logarithm scales using `emmeans(model, ~ Biomass, at = list(Biomass = c()`. The values in `Biomass = c()` were the mean of 24 quantiles of individual plant size on the original scale (gram/individual). 

```{r, include=FALSE}
library(tidyverse)
library(data.table)
library(readr)
library(magrittr) # for %<>%
library(nlme) # for gls
library(emmeans) # for joint_tests
library(patchwork)
library(flextable)
set.seed(722)
```

```{r, include=FALSE}
# bring over the seed - biomass relationship from AMATA-fecundity, 
fecundity18 <- read_csv("../2-Data/Clean/fecundity_18.csv", 
    col_types = cols(Biomass = col_number(), 
        Seed = col_number()), na = "empty")

fecundity18$Crop_ID <- factor(fecundity18$Crop_ID, 
                              levels = c("C2", "S2",
                                         "C3", "S3", "O3",
                                         "C4", "S4", "O4", "A4"))

# Make sure that Block, Crop ID, Corn_weed_management, and bt are factors
fecundity18  <- fecundity18   %<>%
  mutate_at(c("Block","Crop_ID","Corn_weed_management", "bt"),funs(factor(.)))

# sb stands for seed and biomass
fecundity18_sb <- fecundity18[complete.cases(fecundity18$Seed, fecundity18$Biomass), ]

fecundity18_sb_ln_6bin <- fecundity18_sb %>% 
  mutate(ln_Biomass = log(Biomass + 0.005),
         ln_Seed = log(Seed + 1)) %>%
  group_by(Crop_ID, Corn_weed_management) %>% # pool all plants from the same CropID x Corn weed mgt
  mutate(quantile_Biomass = ntile(ln_Biomass , 6),
         quantile_Seed = ntile(ln_Seed, 6)) %>%
  dplyr::mutate(tag = ifelse(quantile_Biomass == quantile_Seed,  "matched", "mismatched"))  # check if Biomass and Seed quantiles are aligned: ONLY 202 are matched.

##
fecundity18_seed_ln_q_mean_6bin <-  fecundity18_sb_ln_6bin %>%
  group_by(Crop_ID, Corn_weed_management, quantile_Seed) %>%
  summarize(mean_ln_Seed = mean(ln_Seed))
```



```{r ln-seed-bar, fig.cap= "Summary of individual fecundity on natural logarithm in six quantiles", include = FALSE}
## since biomass is used to estimate fecundity, use biomass bins to partition data
fecundity18_seed_ln_q_mean_6bin  %>% 
  ggplot(aes(x = mean_ln_Seed, y = as.factor(quantile_Seed))) + 
  facet_grid(Crop_ID ~ Corn_weed_management) +
  geom_bar(stat = "identity") + 
  ylab("Quantile") + 
  xlab(expression(paste("mean of ln( ", Seed ,"+ 1)")))

# fecundity18_seed_ln_q_mean %>% filter(Crop_ID == "A4") # 5.44/2.88 = 1.88; 3.16/2.32 = 1.36
```
```{r biom-seed-eq-copy, include=FALSE}
## overall quantile means, partition the original biomass by its geometric mean's quantile 
fecundity18_sb_ln_whole_6bin <- fecundity18_sb %>% 
  mutate(ln_Biomass = log(Biomass + 0.005),
         ln_Seed = log(Seed + 1)) %>%
  mutate(quantile_Biomass = ntile(ln_Biomass , 6),
         quantile_Seed = ntile(ln_Seed, 6)) %>%
  group_by(quantile_Biomass) %>% 
  summarize(quant_n = n(),
            grand_mean_Biomass = mean(Biomass))


## copy of ancova model 
allcrops.biom.seed.gls <- gls(log(Seed+1) ~ Block + 
                                log(Biomass + 0.005) + 
                                Crop_ID + Corn_weed_management +
                      Crop_ID:Corn_weed_management +
                      Crop_ID:log(Biomass + 0.005) +
                        Corn_weed_management:log(Biomass + 0.005) +
                      Crop_ID:Corn_weed_management:log(Biomass + 0.005),
  correlation=corCompSymm(form= ~1 | bt), #identifies each treatment within block
  data=fecundity18_sb)


## overall, a plant yield about 13000 seeds.  
sb_overall_mean <- emmeans(allcrops.biom.seed.gls, ~ 1) %>%
  as.data.frame() %>%
  mutate(raw_Seed_mean = exp(emmean + .5*SE^2))

## seed crop ID x corn weed management influence on means
sb_emm_6bin <- emmeans(allcrops.biom.seed.gls ,
        ~ Crop_ID * Corn_weed_management | Biomass,
        at = list(Biomass = fecundity18_sb_ln_whole_6bin$grand_mean_Biomass))

sb_emm_6bin_df <- sb_emm_6bin %>% 
  as.data.frame() %>%
  rename(ln_Seed = "emmean",
         SE_Seed = "SE")

# regression with grand Biomass mean fecundity18_sb_ln_whole$grand_mean_ln_Biomass
sb_emm_6bin_grand_mean <- emmeans(allcrops.biom.seed.gls , ~  Biomass,
        at = list(Biomass = fecundity18_sb_ln_whole_6bin$grand_mean_Biomass)) #check with , type = "response": han caculation and automatic conversion are about 1% differ

sb_emm_6bin_grand_mean_df <-  sb_emm_6bin_grand_mean %>% 
  as.data.frame() %>%
  rename(ln_Seed = "emmean",
         SE_ln_Seed = "SE") %>%
  mutate(var_ln_Seed = SE_ln_Seed^2,
         raw_Seed = exp(ln_Seed + var_ln_Seed/2) - 1,
         quant_n = fecundity18_sb_ln_whole_6bin$quant_n, # number of obs in each bin in the original data
         sd_ln_Seed = SE_ln_Seed*sqrt(quant_n))
```

```{r fecund-rlnorm-guess, include=FALSE}
# Guess from sb_emm_6bin_grand_mean_df

## Theoretical relationships: 
# Raw_mean = exp(log scale mean + log scale variance/2).  Note: variance, not sd here.
# (sd/raw_mean)*100 =  Coefficient_of_variation = sqrt( exp(log scale variance) – 1) 

raw_seed_mean = mean(fecundity18_sb$Seed)
ln_seed_mean = mean(log(fecundity18_sb$Seed))
ln_seed_var = var(log(fecundity18_sb$Seed))

raw_seed_sd = sd(fecundity18_sb$Seed) 
CV = 100*(raw_seed_sd/raw_seed_mean)

# gls Model-based ln_seed and SE_seed 
#
#round(,2)


#View(sb_emm_6bin_grand_mean_df %>% select(ln_Seed, sd_ln_Seed))
```

```{r sb-ln-and-ori-long, include=FALSE}
# make a long table 
# fecundity18_sb_ln_long <- fecundity18_sb_ln %>% 
#  select(Block, Corn_weed_management, Crop_ID, Crop,ln_Biomass, ln_Seed) %>%
#  pivot_longer(!c(Block, Corn_weed_management, Crop_ID, Crop),
#               names_to = "Category_ln", values_to = "Value_ln")

#fecundity18_sb_long <- fecundity18_sb_ln %>% 
#  select(Block, Corn_weed_management, Crop_ID, Crop, Biomass, Seed) %>%
#  pivot_longer(!c(Block, Corn_weed_management, Crop_ID, Crop),
#               names_to = "Category_ori", values_to = "Value_ori")


```



```{r 12-bin, include=FALSE}

fecundity18_sb_ln_whole_12bin <- fecundity18_sb %>% 
  mutate(ln_Biomass = log(Biomass + 0.005),
         ln_Seed = log(Seed + 1)) %>%
  mutate(quantile_Biomass = ntile(ln_Biomass , 12),
         quantile_Seed = ntile(ln_Seed, 12)) %>%
  group_by(quantile_Biomass) %>% 
  summarize(quant_n = n(),
            grand_mean_Biomass = mean(Biomass))


sb_emm_12bin_grand_mean <- emmeans(allcrops.biom.seed.gls , ~  Biomass,
        at = list(Biomass = fecundity18_sb_ln_whole_12bin$grand_mean_Biomass))

sb_emm_12bin_grand_mean_raw <- emmeans(allcrops.biom.seed.gls , ~  Biomass,
        at = list(Biomass = fecundity18_sb_ln_whole_12bin$grand_mean_Biomass), type = "response")


sb_emm_12bin_grand_mean_df <-  sb_emm_12bin_grand_mean %>% 
  as.data.frame() %>%
  rename(ln_Seed = "emmean",
         SE_ln_Seed = "SE") %>%
  mutate(var_ln_Seed = SE_ln_Seed^2,
         raw_Seed = exp(ln_Seed + var_ln_Seed/2) - 1,
         quant_n = fecundity18_sb_ln_whole_12bin$quant_n, # number of obs in each bin in the original data
         sd_ln_Seed = SE_ln_Seed*sqrt(quant_n))

#View(sb_emm_12bin_grand_mean_df %>% select(ln_Seed, sd_ln_Seed))
```

```{r 24-bin, echo=FALSE}

fecundity18_sb_ln_whole_24bin <- fecundity18_sb %>% 
  mutate(ln_Biomass = log(Biomass + 0.005),
         ln_Seed = log(Seed + 1)) %>%
  mutate(quantile_Biomass = ntile(ln_Biomass , 24),
         quantile_Seed = ntile(ln_Seed, 24)) %>%
  group_by(quantile_Biomass) %>% 
  summarize(quant_n = n(),
            grand_mean_Biomass = mean(Biomass))


sb_emm_24bin_grand_mean <- emmeans(allcrops.biom.seed.gls , ~  Biomass,
        at = list(Biomass = fecundity18_sb_ln_whole_24bin$grand_mean_Biomass))

sb_emm_24bin_grand_mean_raw <- emmeans(allcrops.biom.seed.gls , ~  Biomass,
        at = list(Biomass = fecundity18_sb_ln_whole_24bin$grand_mean_Biomass), type = "response")


sb_emm_24bin_grand_mean_df <-  sb_emm_24bin_grand_mean %>% 
  as.data.frame() %>%
  rename(ln_Seed = "emmean",
         SE_ln_Seed = "SE") %>%
  mutate(var_ln_Seed = SE_ln_Seed^2,
        Seed = exp(ln_Seed + var_ln_Seed/2) - 1,
         quant_n = fecundity18_sb_ln_whole_24bin$quant_n, # number of obs in each bin in the original data
         sd_ln_Seed = SE_ln_Seed*sqrt(quant_n))

#View(sb_emm_24bin_grand_mean_df %>% select(Biomass, raw_Seed, ln_Seed, sd_ln_Seed))

sb_emm_24bin_grand_mean_df %>% 
  select(Biomass, Seed, ln_Seed, SE_ln_Seed, sd_ln_Seed) %>%
  rename(m = "ln_Seed", s = "sd_ln_Seed", `Aboveground mass` = "Biomass", M = "Seed") %>%
  flextable() %>%
  colformat_double(big.mark=",", digits = 2, na_str = "N/A") %>%
  set_caption("Mean on original scale (M) and mean (m) and standard deviation (s) on natural logarithm scale estimation from a set of mean aboveground mass values. m and s were used to assign the mean and standard deviation values in the rlnorm function in the simulation.")

```