---
output:
 # bookdown::word_document2:
  bookdown::html_document2:
      toc: false
      fig_caption: yes
      keep_md: true
bibliography: WH-pop-dynamics.bib
csl: apa-no-ampersand.csl 
---

The data in the model projection was used in this simulation. 100 iterations of simulation were run per each rotation crossed with corn weed management regime. Different simulation outputs are collected at different chunks: x-lambda-x-x for population growth rate, x-seed-production-per-capita for manipulated seed production per plant, x-plant-density-fixed for mature plant density as empirically measure, and x-all-output for all the outputs combined. 

```{r, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(readr)
library(nlme) # for gls
library(emmeans) # for joint_tests
library(patchwork)
library(pracma) #for nthroot
set.seed(722)
source("seed-production-threshold-simulation-functions.R")
```


```{r, echo=FALSE}
starting_point <-  c(10000, 0, 0, 0, 0, 0, 0, 0) #10000 seeds top soil, 0 seeds deep soil, 0 plants from cohorts 1 through 6
t <- 10000  #number of iteration
```

```{r tillage-sim-a, echo=FALSE}
fall_tillage <-readRDS("../2-Data/Clean/mean-post-harvest-tillage.RData")

spring_tillage <-readRDS("../2-Data/Clean/mean-pre-planting-tillage.RData")
```

```{r emerge-sim-a, echo=FALSE}
emergence <-readRDS("../2-Data/Clean/mean-emergence-prop-adjusted.RData")
```

```{r survive-sim-a, echo=FALSE}
summer_seed_survival <- readRDS("../2-Data/Clean/mean-summer-seed-survival-Sosnoskie.RData")

#### plant survival: 
female_survival <- readRDS("../2-Data/Clean/mean-summer-seedling-survival-Hartzler.RData")

#### combine seed and plant survivals into one matrix with element-wise multiplication: this is not computing new values, but arranging using 1's as placeholders 
summer_survival <- purrr::map2(summer_seed_survival, female_survival, `*`)
```

```{r fecund-sim-a, echo=FALSE}
# First step: use a cumulative fecundity of all cohorts as `
#new_seeds_total`=  mature_plant_total * fecundity_sim
# change fecundity_sim  to get lambda as close to 1 as possible 

# manipulate cohorts 1 through 3 in corn and soybean fecundity only because 1) these cohorts are more likely exposed to control measures and 2) they have higher survival rates than cohorts 3+ 

# cohort 3+ fecundity: estimated fecundity as in population projection 

# Final goal: flatten the fecundity distribution curve in corn and soybean to see seed production "allowance" in cohorts 1 and 2 3

### find good pairs of mean and sd for rlnorm in Q1-fecund-prep

fecundity18 <- readRDS("../2-Data/Clean/mean-fecundity-18-cohort.RData")

## Reference data frame, with the same value as fecundity18 but in data frame format for weighing how much seed production could be in each rlnorm iteration

fecundity18_df_ref <- read.csv("../2-Data/Clean/mean-fecundity-18-cohort-df.csv")
```

```{r overwinter-sim-a, echo=FALSE}
overwinter <- readRDS("../2-Data/Clean/mean-winter-seed-survival-Sosnoskie.RData")
```


```{r 2yr-lambda-sim-a}
# event sequence: seed dropped - field cultivator - emerge - survive - new seed - chisel - overwinter 

# create a function 
# vec: starting seed column
# prt: pre-planting-tillage
# em: emergence
# sv: seed survival rate and seedling to maturity success rate
# seed: fecundity
# poh: post-harvest tillage
# ow: over winter seed survival
##### with corn under conventional weed management {-}

### Output: seedbank at top and bottom strata at the end of each crop phase

N_2yr_conv_lambda <- list() # blank data frame to save loop output 
N_2yr_conv_lambda[[1]] <- starting_point 

for (i in 2:t) { 
  N_2yr_conv_lambda[[i]] = rot_2year_conv_lambda(vec = N_2yr_conv_lambda[[i-1]],
                              poh_C = fall_tillage$C2_conv,
                              ow_C = overwinter$C2_conv,
                              prt_C  = spring_tillage$C2_conv,
                              em_C  = emergence$C2_conv,
                              sv_C = summer_survival$C2_conv,
                              seed_C = fecundity18$C2_conv,
                              
                              #soybean dynamics   
                              poh_S = fall_tillage$S2_conv,
                              ow_S = overwinter$S2_conv,
                              prt_S  = spring_tillage$S2_conv,
                              em_S  = emergence$S2_conv,
                              sv_S = summer_survival$S2_conv,
                              seed_S = fecundity18$S2_conv)
}

N_2yr_conv_lambda_df <-  N_2yr_conv_lambda %>% 
  unlist(recursive = TRUE) %>%
  data.frame() %>%
    dplyr::rename(seedbank_counts = ".") %>%
  dplyr::mutate(stratum = rep(c("top", "bottom", 
                                "cohort_1", "cohort_2", 
                                "cohort_3", "cohort_4",
                                "cohort_5", "cohort_6"),t)) %>%
    filter(stratum %in% c("top", "bottom")) %>%
    unnest(cols = everything())  %>%
    mutate(cycle_no = as.character(rep(1:t, each = 2))) %>%
  pivot_wider(names_from = stratum, values_from = seedbank_counts) %>%
  mutate(total_seedbank_counts = top + bottom) %>%
  mutate(lambda_cycle = total_seedbank_counts/lag(total_seedbank_counts),
         lambda_annualized = sqrt(lambda_cycle),
          Rotation = "2-year",
         Corn_weed_management = "conventional") 

##### with corn under low herbicide weed management {-}
N_2yr_low_lambda <- list() # blank data frame to save loop output 

N_2yr_low_lambda[[1]] <- starting_point 
for (i in 2:t) { 
  N_2yr_low_lambda[[i]] = rot_2year_low_lambda(vec = N_2yr_low_lambda[[i-1]],
                             poh_C = fall_tillage$C2_low,
                             ow_C = overwinter$C2_low,
                             prt_C  = spring_tillage$C2_low,
                             em_C  = emergence$C2_low,
                             sv_C = summer_survival$C2_low,
                             seed_C = fecundity18$C2_low,
                             
                             #soybean dynamics   
                             poh_S = fall_tillage$S2_low,
                             ow_S = overwinter$S2_low,
                             prt_S  = spring_tillage$S2_low,
                             em_S  = emergence$S2_low,
                             sv_S = summer_survival$S2_low,
                             seed_S = fecundity18$S2_low)
}

N_2yr_low_lambda_df <-  N_2yr_low_lambda %>% 
  unlist(recursive = TRUE) %>%
  data.frame() %>%
    dplyr::rename(seedbank_counts = ".") %>%
  dplyr::mutate(stratum = rep(c("top", "bottom", 
                                "cohort_1", "cohort_2", 
                                "cohort_3", "cohort_4",
                                "cohort_5", "cohort_6"),t)) %>%
    filter(stratum %in% c("top", "bottom")) %>%
    unnest(cols = everything())  %>%
    mutate(cycle_no = as.character(rep(1:t, each = 2))) %>%
  pivot_wider(names_from = stratum, values_from = seedbank_counts) %>%
  mutate(total_seedbank_counts = top + bottom) %>%
  mutate(lambda_cycle = total_seedbank_counts/lag(total_seedbank_counts),
         lambda_annualized = sqrt(lambda_cycle),
          Rotation = "2-year",
         Corn_weed_management = "low") 

```


```{r 2yr-plant-density-fixed}
### Output: Mature plant densities until seed production (B_h = sv_C or sv_S)
### 1 iteration only because no randomization at any matrix
N_2yr_conv_mature_plant_density_fixed <- rot_2year_conv_plant_density_fixed(vec = starting_point ,
                              poh_C = fall_tillage$C2_conv,
                              ow_C = overwinter$C2_conv,
                              prt_C  = spring_tillage$C2_conv,
                              em_C  = emergence$C2_conv,
                              sv_C = summer_survival$C2_conv,
                              seed_C = fecundity18$C2_conv,
                              
                              #soybean dynamics   
                              poh_S = fall_tillage$S2_conv,
                              ow_S = overwinter$S2_conv,
                              prt_S  = spring_tillage$S2_conv,
                              em_S  = emergence$S2_conv,
                              sv_S = summer_survival$S2_conv,
                              seed_S = fecundity18$S2_conv)

N_2yr_conv_mature_plant_density_fixed_df <- N_2yr_conv_mature_plant_density_fixed  %>% 
  unlist(recursive = TRUE) %>% # make a long table
  data.frame() %>% #corn was listed before soybean in the customized f.
  dplyr::rename(plant_counts_fixed = ".")  %>%
  dplyr::mutate(cohort = rep(c("1", "2", "3", "4", "5", "6"),2)) %>%
   mutate(Crop_ID = c(rep("C2", 6), rep("S2", 6)))


### Output: Mature plant densities until seed production (B_h = sv_C or sv_S)
### 1 iteration only because no randomization at any matrix

N_2yr_low_mature_plant_density_fixed <- rot_2year_low_plant_density_fixed(vec = starting_point ,
                             poh_C = fall_tillage$C2_low,
                             ow_C = overwinter$C2_low,
                             prt_C  = spring_tillage$C2_low,
                             em_C  = emergence$C2_low,
                             sv_C = summer_survival$C2_low,
                             seed_C = fecundity18$C2_low,
                             
                             #soybean dynamics   
                             poh_S = fall_tillage$S2_low,
                             ow_S = overwinter$S2_low,
                             prt_S  = spring_tillage$S2_low,
                             em_S  = emergence$S2_low,
                             sv_S = summer_survival$S2_low,
                             seed_S = fecundity18$S2_low)

N_2yr_low_mature_plant_density_fixed_df <- N_2yr_low_mature_plant_density_fixed  %>% 
  unlist(recursive = TRUE) %>% # make a long table
  data.frame() %>% #corn was listed before soybean in the customized f.
  dplyr::rename(plant_counts_fixed = ".")  %>%
  dplyr::mutate(cohort = rep(c("1", "2", "3", "4", "5", "6"),2)) %>%
   mutate(Crop_ID = c(rep("C2", 6), rep("S2", 6)))
```

```{r 2yr-seed-production-per-capita}
### CONV Output: seed production, after manipulation by cohort x crop phase x iteration

N_2yr_conv_seed_production_per_capita <- list() # blank data frame to save loop output 

for (i in 1:t) { 
  N_2yr_conv_seed_production_per_capita[[i]] =
    rot_2year_conv_seed_production_per_cap(seed_C = fecundity18$C2_conv,
                                           seed_S = fecundity18$S2_conv)
}

# original per capita seed production in fecundity18$C2_conv[1,3:8] and fecundity18$S2_conv[1,3:8]
N_2yr_conv_seed_production_per_capita_df <- N_2yr_conv_seed_production_per_capita  %>% 
  unlist(recursive = TRUE) %>% # make a long table
  data.frame() %>% #corn was listed before soybean in the customized f.
  dplyr::rename(seed_per_capita_threshold = ".") %>%
  dplyr::mutate(cohort = rep(c("1", "2", "3", "4", "5", "6"), t*2)) %>%
   mutate(cycle_no = as.character(rep(1:t, each = 12))) %>% # 12 = 6 cohorts x 2 phases
   mutate(Crop_ID = rep(c(rep("C2", 6), rep("S2", 6)), t)) %>%
  mutate(seed_per_capita_original = rep(c(fecundity18$C2_conv[1,3:8],
                                          fecundity18$S2_conv[1,3:8]), t)) %>%
  mutate(seed_production_manipulated = ifelse(seed_per_capita_threshold == seed_per_capita_original, "no", "yes"))
  

### LOW output

N_2yr_low_seed_production_per_capita <- list() # blank data frame to save loop output 

for (i in 1:t) { 
  N_2yr_low_seed_production_per_capita[[i]] =
    rot_2year_low_seed_production_per_cap(seed_C = fecundity18$C2_low,
                                          seed_S = fecundity18$S2_low)
}


N_2yr_low_seed_production_per_capita_df <- N_2yr_low_seed_production_per_capita  %>% 
  unlist(recursive = TRUE) %>% # make a long table
  data.frame() %>% #corn was listed before soybean in the customized f.
  dplyr::rename(seed_per_capita_threshold = ".") %>%
  dplyr::mutate(cohort = rep(c("1", "2", "3", "4", "5", "6"), t*2)) %>%
   mutate(cycle_no = as.character(rep(1:t, each = 12))) %>% # 12 = 6 cohorts x 2 phases
   mutate(Crop_ID = rep(c(rep("C2", 6), rep("S2", 6)), t)) %>%
  mutate(seed_per_capita_original = rep(c(fecundity18$C2_low[1,3:8],
                                          fecundity18$S2_low[1,3:8]), t)) %>%
  mutate(seed_production_manipulated = ifelse(seed_per_capita_threshold == seed_per_capita_original, "no", "yes"))
```



```{r 2yr-sim-a-all-output}
# total stratified seedbank density and lambdas in N_2yr_conv_lambda_df
# manipulated seed production in N_2yr_conv_seed_production_per_capita_df
# fixed mature plant density in N_2yr_conv_mature_plant_density_fixed_df 

N_2yr_conv_all_df <- N_2yr_conv_seed_production_per_capita_df %>% 
  left_join(N_2yr_conv_lambda_df, by = "cycle_no") %>%
  left_join(N_2yr_conv_mature_plant_density_fixed_df, by = c("Crop_ID", "cohort")) 


N_2yr_low_all_df <- N_2yr_low_seed_production_per_capita_df %>% 
  left_join(N_2yr_low_lambda_df, by = "cycle_no") %>%
  left_join(N_2yr_low_mature_plant_density_fixed_df, by = c("Crop_ID", "cohort")) 
```



```{r 3yr-lambda-sim-a}
##### with corn under conventional weed management {-}
N_3yr_conv_lambda <- list() # blank dataframe to save loop output 

N_3yr_conv_lambda[[1]] <- starting_point 


for (i in 2:t) { 
  N_3yr_conv_lambda[[i]] = rot_3year_conv_lambda(vec = N_3yr_conv_lambda[[i-1]],
                              poh_C = fall_tillage$C3_conv,
                              ow_C = overwinter$C3_conv,
                              prt_C  = spring_tillage$C3_conv,
                              em_C  = emergence$C3_conv,
                              sv_C = summer_survival$C3_conv,
                              seed_C = fecundity18$C3_conv,
                              
                              #soybean dynamics   
                              poh_S = fall_tillage$S3_conv,
                              ow_S = overwinter$S3_conv,
                              prt_S  = spring_tillage$S3_conv,
                              em_S  = emergence$S3_conv,
                              sv_S = summer_survival$S3_conv,
                              seed_S = fecundity18$S3_conv,
                              
                              #oat dynamics   
                              poh_O = fall_tillage$O3_conv,
                              ow_O = overwinter$O3_conv,
                              prt_O  = spring_tillage$O3_conv,
                              em_O  = emergence$O3_conv,
                              sv_O = summer_survival$O3_conv,
                              seed_O = fecundity18$O3_conv)
}

N_3yr_conv_lambda_df <-  N_3yr_conv_lambda %>% 
  unlist(recursive = TRUE) %>%
  data.frame() %>%
    dplyr::rename(seedbank_counts = ".") %>%
  dplyr::mutate(stratum = rep(c("top", "bottom",
                                "cohort_1", "cohort_2",
                                "cohort_3", "cohort_4",
                                "cohort_5", "cohort_6"),t)) %>%
    filter(stratum %in% c("top", "bottom")) %>%
    unnest(cols = everything() ) %>%
    mutate(cycle_no = as.character(rep(1:t, each = 2))) %>%
  pivot_wider(names_from = stratum, values_from = seedbank_counts) %>%
  mutate(total_seedbank_counts = top + bottom) %>%
  mutate(lambda_cycle = total_seedbank_counts/lag(total_seedbank_counts),
         lambda_annualized = sqrt(lambda_cycle),
          Rotation = "3-year",
         Corn_weed_management = "conventional") 

##### with corn under low herbicide weed management {-} 
N_3yr_low_lambda <- list() # blank dataframe to save loop output 

N_3yr_low_lambda[[1]] <- starting_point 


for (i in 2:t) { 
  N_3yr_low_lambda[[i]] = rot_3year_low_lambda(vec = N_3yr_low_lambda[[i-1]],
                             poh_C = fall_tillage$C3_low,
                             ow_C = overwinter$C3_low,
                             prt_C  = spring_tillage$C3_low,
                             em_C  = emergence$C3_low,
                             sv_C = summer_survival$C3_low,
                             seed_C = fecundity18$C3_low,
                             
                             #soybean dynamics   
                             poh_S = fall_tillage$S3_low,
                             ow_S = overwinter$S3_low,
                             prt_S  = spring_tillage$S3_low,
                             em_S  = emergence$S3_low,
                             sv_S = summer_survival$S3_low,
                             seed_S = fecundity18$S3_low,
                             
                             #oat dynamics   
                             poh_O = fall_tillage$O3_low,
                             ow_O = overwinter$O3_low,
                             prt_O  = spring_tillage$O3_low,
                             em_O  = emergence$O3_low,
                             sv_O = summer_survival$O3_low,
                             seed_O = fecundity18$O3_low)
}

N_3yr_low_lambda_df <-  N_3yr_low_lambda %>% 
  unlist(recursive = TRUE) %>%
  data.frame() %>%
    dplyr::rename(seedbank_counts = ".") %>%
  dplyr::mutate(stratum = rep(c("top", "bottom",
                                "cohort_1", "cohort_2",
                                "cohort_3", "cohort_4",
                                "cohort_5", "cohort_6"),t)) %>%
    filter(stratum %in% c("top", "bottom")) %>%
    unnest(cols = everything() ) %>%
    mutate(cycle_no = as.character(rep(1:t, each = 2))) %>%
  pivot_wider(names_from = stratum, values_from = seedbank_counts) %>%
  mutate(total_seedbank_counts = top + bottom) %>%
  mutate(lambda_cycle = total_seedbank_counts/lag(total_seedbank_counts),
         lambda_annualized = sqrt(lambda_cycle),
          Rotation = "3-year",
         Corn_weed_management = "low") 
```


```{r 3yr-plant-density-fixed}
### Output: Mature plant densities until seed production (B_h = sv_C or sv_S)
### 1 iteration only because no randomization at any matrix
N_3yr_conv_mature_plant_density_fixed <- rot_3year_conv_plant_density_fixed(vec = starting_point ,
                              poh_C = fall_tillage$C3_conv,
                              ow_C = overwinter$C3_conv,
                              prt_C  = spring_tillage$C3_conv,
                              em_C  = emergence$C3_conv,
                              sv_C = summer_survival$C3_conv,
                              seed_C = fecundity18$C3_conv,
                              
                              #soybean dynamics   
                              poh_S = fall_tillage$S3_conv,
                              ow_S = overwinter$S3_conv,
                              prt_S  = spring_tillage$S3_conv,
                              em_S  = emergence$S3_conv,
                              sv_S = summer_survival$S3_conv,
                              seed_S = fecundity18$S3_conv,
                              
                              #oat dynamics   
                              poh_O = fall_tillage$O3_conv,
                              ow_O = overwinter$O3_conv,
                              prt_O  = spring_tillage$O3_conv,
                              em_O  = emergence$O3_conv,
                              sv_O = summer_survival$O3_conv,
                              seed_O = fecundity18$O3_conv)

N_3yr_conv_mature_plant_density_fixed_df <- N_3yr_conv_mature_plant_density_fixed  %>% 
  unlist(recursive = TRUE) %>% # make a long table
  data.frame() %>% #corn was listed before soybean in the customized f.
  dplyr::rename(plant_counts_fixed = ".")  %>%
  dplyr::mutate(cohort = rep(c("1", "2", "3", "4", "5", "6"), 3)) %>% # 3 phases
   mutate(Crop_ID = c(rep("C3", 6), rep("S3", 6), rep("O3", 6)))


### Output: Mature plant densities until seed production (B_h = sv_C or sv_S)
### 1 iteration only because no randomization at any matrix
N_3yr_low_mature_plant_density_fixed <- rot_3year_low_plant_density_fixed(vec = starting_point,
                             poh_C = fall_tillage$C3_low,
                             ow_C = overwinter$C3_low,
                             prt_C  = spring_tillage$C3_low,
                             em_C  = emergence$C3_low,
                             sv_C = summer_survival$C3_low,
                             seed_C = fecundity18$C3_low,
                             
                             #soybean dynamics   
                             poh_S = fall_tillage$S3_low,
                             ow_S = overwinter$S3_low,
                             prt_S  = spring_tillage$S3_low,
                             em_S  = emergence$S3_low,
                             sv_S = summer_survival$S3_low,
                             seed_S = fecundity18$S3_low,
                             
                             #oat dynamics   
                             poh_O = fall_tillage$O3_low,
                             ow_O = overwinter$O3_low,
                             prt_O  = spring_tillage$O3_low,
                             em_O  = emergence$O3_low,
                             sv_O = summer_survival$O3_low,
                             seed_O = fecundity18$O3_low)

N_3yr_low_mature_plant_density_fixed_df <- N_3yr_low_mature_plant_density_fixed  %>% 
  unlist(recursive = TRUE) %>% # make a long table
  data.frame() %>% #corn was listed before soybean in the customized f.
  dplyr::rename(plant_counts_fixed = ".")  %>%
  dplyr::mutate(cohort = rep(c("1", "2", "3", "4", "5", "6"), 3)) %>% #3 phases
   mutate(Crop_ID = c(rep("C3", 6), rep("S3", 6), rep("O3", 6)))
```



```{r 3yr-seed-production-per-capita}
### CONV Output: seed production, after manipulation by cohort x crop phase x iteration
N_3yr_conv_seed_production_per_capita <- list() # blank data frame to save loop output 


for (i in 1:t) { 
  N_3yr_conv_seed_production_per_capita[[i]] =
    rot_3year_conv_seed_production_per_cap(seed_C = fecundity18$C3_conv,
                                           seed_S = fecundity18$S3_conv,
                                           seed_O = fecundity18$O3_conv)
}

# original per capita seed production in fecundity18$C3_conv[1,3:8], fecundity18$S3_conv[1,3:8], and fecundity18$O3_conv
N_3yr_conv_seed_production_per_capita_df <- N_3yr_conv_seed_production_per_capita  %>% 
  unlist(recursive = TRUE) %>% # make a long table
  data.frame() %>% #corn was listed before soybean in the customized f.
  dplyr::rename(seed_per_capita_threshold = ".") %>%
  dplyr::mutate(cohort = rep(c("1", "2", "3", "4", "5", "6"), t*3)) %>% #*3 for 3 phases
   mutate(cycle_no = as.character(rep(1:t, each = 18))) %>% # 18 = 6 cohorts x 3 phases
   mutate(Crop_ID = rep(c(rep("C3", 6), rep("S3", 6), rep("O3", 6)), t)) %>%
  mutate(seed_per_capita_original = rep(c(fecundity18$C3_conv[1,3:8],
                                          fecundity18$S3_conv[1,3:8],
                                          fecundity18$O3_conv[1,3:8]), t)) %>%
  mutate(seed_production_manipulated = ifelse(seed_per_capita_threshold == seed_per_capita_original, "no", "yes"))

### LOW output

N_3yr_low_seed_production_per_capita <- list() # blank data frame to save loop output 


for (i in 1:t) { 
  N_3yr_low_seed_production_per_capita[[i]] =
    rot_3year_low_seed_production_per_cap(seed_C = fecundity18$C3_low,
                                          seed_S = fecundity18$S3_low,
                                          seed_O = fecundity18$O3_low)
}


N_3yr_low_seed_production_per_capita_df <- N_3yr_low_seed_production_per_capita  %>% 
  unlist(recursive = TRUE) %>% # make a long table
  data.frame() %>% #corn was listed before soybean in the customized f.
  dplyr::rename(seed_per_capita_threshold = ".") %>%
  dplyr::mutate(cohort = rep(c("1", "2", "3", "4", "5", "6"), t*3)) %>% #*3 for 3 phases
   mutate(cycle_no = as.character(rep(1:t, each = 18))) %>% # 18 = 6 cohorts x 3 phases
   mutate(Crop_ID = rep(c(rep("C3", 6), rep("S3", 6), rep("O3", 6)), t)) %>%
  mutate(seed_per_capita_original = rep(c(fecundity18$C3_low[1,3:8],
                                          fecundity18$S3_low[1,3:8],
                                          fecundity18$O3_low[1,3:8]), t)) %>%
  mutate(seed_production_manipulated = ifelse(seed_per_capita_threshold == seed_per_capita_original, "no", "yes"))
```


```{r 3yr-sim-a-all-output}
# total stratified seedbank density and lambdas in N_3yr_conv_df
# manipulated seed production in N_3yr_conv_seed_production_per_capita_df
# fixed mature plant density in N_3yr_conv_mature_plant_density_fixed_df 

N_3yr_conv_all_df <- N_3yr_conv_seed_production_per_capita_df %>% 
  left_join(N_3yr_conv_mature_plant_density_fixed_df, by = c("Crop_ID", "cohort")) %>%
  left_join(N_3yr_conv_lambda_df, by = "cycle_no")


N_3yr_low_all_df <- N_3yr_low_seed_production_per_capita_df %>% 
  left_join(N_3yr_low_lambda_df, by = "cycle_no") %>%
  left_join(N_3yr_low_mature_plant_density_fixed_df, by = c("Crop_ID", "cohort"))  
```



```{r 4yr-lambda-sim-a}
##### with corn under conventional weed management {-}
N_4yr_conv_lambda <- list() # blank dataframe to save loop output 

N_4yr_conv_lambda[[1]] <- starting_point 

for (i in 2:t) { 
  N_4yr_conv_lambda[[i]] = rot_4year_conv_lambda(vec = N_4yr_conv_lambda[[i-1]],
                              poh_C = fall_tillage$C4_conv,
                              ow_C = overwinter$C4_conv,
                              prt_C  = spring_tillage$C4_conv,
                              em_C  = emergence$C4_conv,
                              sv_C = summer_survival$C4_conv,
                              seed_C = fecundity18$C4_conv,
                              
                              #soybean dynamics   
                              poh_S = fall_tillage$S4_conv,
                              ow_S = overwinter$S4_conv,
                              prt_S  = spring_tillage$S4_conv,
                              em_S  = emergence$S4_conv,
                              sv_S = summer_survival$S4_conv,
                              seed_S = fecundity18$S4_conv,
                              
                              #oat dynamics   
                              poh_O = fall_tillage$O4_conv,
                              ow_O = overwinter$O4_conv,
                              prt_O  = spring_tillage$O4_conv,
                              em_O  = emergence$O4_conv,
                              sv_O = summer_survival$O4_conv,
                              seed_O = fecundity18$O4_conv,
                              
                              #alfalfa dynamics   
                          poh_A = fall_tillage$A4_conv,
                          ow_A = overwinter$A4_conv,
                          prt_A  = spring_tillage$A4_conv,
                          em_A  = emergence$A4_conv,
                          sv_A = summer_survival$A4_conv,
                          seed_A = fecundity18$A4_conv)
}

N_4yr_conv_lambda_df <-  N_4yr_conv_lambda %>% 
  unlist(recursive = TRUE) %>%
  data.frame() %>%
    dplyr::rename(seedbank_counts = ".") %>%
  dplyr::mutate(stratum = rep(c("top", "bottom",
                                "cohort_1", "cohort_2",
                                "cohort_3", "cohort_4",
                                "cohort_5", "cohort_6"),t)) %>%
    filter(stratum %in% c("top", "bottom")) %>%
    unnest(cols = everything())  %>%
    mutate(cycle_no = as.character(rep(1:t, each = 2))) %>%
  pivot_wider(names_from = stratum, values_from = seedbank_counts) %>%
  mutate(total_seedbank_counts = top + bottom) %>%
  mutate(lambda_cycle = total_seedbank_counts/lag(total_seedbank_counts),
         lambda_annualized = sqrt(lambda_cycle),
          Rotation = "4-year",
         Corn_weed_management = "conventional") 

##### with corn under low herbicide weed management {-} 
N_4yr_low_lambda <- list() # blank dataframe to save loop output 

N_4yr_low_lambda[[1]] <- starting_point 

for (i in 2:t) { 
  N_4yr_low_lambda[[i]] = rot_4year_low_lambda(vec = N_4yr_low_lambda[[i-1]],
                              poh_C = fall_tillage$C4_low,
                              ow_C = overwinter$C4_low,
                              prt_C  = spring_tillage$C4_low,
                              em_C  = emergence$C4_low,
                              sv_C = summer_survival$C4_low,
                              seed_C = fecundity18$C4_low,
                              
                              #soybean dynamics   
                              poh_S = fall_tillage$S4_low,
                              ow_S = overwinter$S4_low,
                              prt_S  = spring_tillage$S4_low,
                              em_S  = emergence$S4_low,
                              sv_S = summer_survival$S4_low,
                              seed_S = fecundity18$S4_low,
                              
                              #oat dynamics   
                              poh_O = fall_tillage$O4_low,
                              ow_O = overwinter$O4_low,
                              prt_O  = spring_tillage$O4_low,
                              em_O  = emergence$O4_low,
                              sv_O = summer_survival$O4_low,
                              seed_O = fecundity18$O4_low,
                              
                              #alfalfa dynamics   
                              poh_A = fall_tillage$A4_low,
                              ow_A = overwinter$A4_low,
                              prt_A  = spring_tillage$A4_low,
                              em_A  = emergence$A4_low,
                              sv_A = summer_survival$A4_low,
                              seed_A = fecundity18$A4_low)
}

N_4yr_low_lambda_df <-  N_4yr_low_lambda %>% 
  unlist(recursive = TRUE) %>%
  data.frame() %>%
    dplyr::rename(seedbank_counts = ".") %>%
  dplyr::mutate(stratum = rep(c("top", "bottom",
                                "cohort_1", "cohort_2",
                                "cohort_3", "cohort_4",
                                "cohort_5", "cohort_6"),t)) %>%
    filter(stratum %in% c("top", "bottom")) %>%
    unnest(cols = everything())  %>%
    mutate(cycle_no = as.character(rep(1:t, each = 2))) %>%
  pivot_wider(names_from = stratum, values_from = seedbank_counts) %>%
  mutate(total_seedbank_counts = top + bottom) %>%
  mutate(lambda_cycle = total_seedbank_counts/lag(total_seedbank_counts),
         lambda_annualized = nthroot(lambda_cycle, 4),
         Rotation = "4-year",
         Corn_weed_management = "low")
```


```{r 4yr-plant-density-fixed}
### Output: Mature plant densities until seed production (B_h = sv_C or sv_S)
### 1 iteration only because no randomization at any matrix
N_4yr_conv_mature_plant_density_fixed <- rot_4year_conv_plant_density_fixed(vec = starting_point ,
                              poh_C = fall_tillage$C4_conv,
                              ow_C = overwinter$C4_conv,
                              prt_C  = spring_tillage$C4_conv,
                              em_C  = emergence$C4_conv,
                              sv_C = summer_survival$C4_conv,
                              seed_C = fecundity18$C4_conv,
                              
                              #soybean dynamics   
                              poh_S = fall_tillage$S4_conv,
                              ow_S = overwinter$S4_conv,
                              prt_S  = spring_tillage$S4_conv,
                              em_S  = emergence$S4_conv,
                              sv_S = summer_survival$S4_conv,
                              seed_S = fecundity18$S4_conv,
                              
                              #oat dynamics   
                              poh_O = fall_tillage$O4_conv,
                              ow_O = overwinter$O4_conv,
                              prt_O  = spring_tillage$O4_conv,
                              em_O  = emergence$O4_conv,
                              sv_O = summer_survival$O4_conv,
                              seed_O = fecundity18$O4_conv,
                                                            
                              #alfalfa dynamics   
                          poh_A = fall_tillage$A4_conv,
                          ow_A = overwinter$A4_conv,
                          prt_A  = spring_tillage$A4_conv,
                          em_A  = emergence$A4_conv,
                          sv_A = summer_survival$A4_conv,
                          seed_A = fecundity18$A4_conv)

N_4yr_conv_mature_plant_density_fixed_df <- N_4yr_conv_mature_plant_density_fixed  %>% 
  unlist(recursive = TRUE) %>% # make a long table
  data.frame() %>% #corn was listed before soybean in the customized f.
  dplyr::rename(plant_counts_fixed = ".")  %>%
  dplyr::mutate(cohort = rep(c("1", "2", "3", "4", "5", "6"), 4)) %>% #4phases
   mutate(Crop_ID = c(rep("C4", 6), rep("S4", 6), rep("O4", 6), rep("A4", 6)))


### Output: Mature plant densities until seed production (B_h = sv_C or sv_S)
### 1 iteration only because no randomization at any matrix


N_4yr_low_mature_plant_density_fixed <- rot_4year_low_plant_density_fixed(vec = starting_point,
                             poh_C = fall_tillage$C4_low,
                             ow_C = overwinter$C4_low,
                             prt_C  = spring_tillage$C4_low,
                             em_C  = emergence$C4_low,
                             sv_C = summer_survival$C4_low,
                             seed_C = fecundity18$C4_low,
                             
                             #soybean dynamics   
                             poh_S = fall_tillage$S4_low,
                             ow_S = overwinter$S4_low,
                             prt_S  = spring_tillage$S4_low,
                             em_S  = emergence$S4_low,
                             sv_S = summer_survival$S4_low,
                             seed_S = fecundity18$S4_low,
                             
                             #oat dynamics   
                             poh_O = fall_tillage$O4_low,
                             ow_O = overwinter$O4_low,
                             prt_O  = spring_tillage$O4_low,
                             em_O  = emergence$O4_low,
                             sv_O = summer_survival$O4_low,
                             seed_O = fecundity18$O4_low,
                              
                              #alfalfa dynamics   
                              poh_A = fall_tillage$A4_low,
                              ow_A = overwinter$A4_low,
                              prt_A  = spring_tillage$A4_low,
                              em_A  = emergence$A4_low,
                              sv_A = summer_survival$A4_low,
                              seed_A = fecundity18$A4_low)

N_4yr_low_mature_plant_density_fixed_df <- N_4yr_low_mature_plant_density_fixed  %>% 
  unlist(recursive = TRUE) %>% # make a long table
  data.frame() %>% #corn was listed before soybean in the customized f.
  dplyr::rename(plant_counts_fixed = ".")  %>%
  dplyr::mutate(cohort = rep(c("1", "2", "3", "4", "5", "6"), 4)) %>% # 4 phases
   mutate(Crop_ID = c(rep("C4", 6), rep("S4", 6), rep("O4", 6), rep("A4", 6)))

```

```{r 4yr-seed-production-per-capita}
### CONV Output: seed production, after manipulation by cohort x crop phase x iteration


N_4yr_conv_seed_production_per_capita <- list() # blank data frame to save loop output 


for (i in 1:t) { 
  N_4yr_conv_seed_production_per_capita[[i]] =
    rot_4year_conv_seed_production_per_cap(seed_C = fecundity18$C4_conv,
                                           seed_S = fecundity18$S4_conv,
                                           seed_O = fecundity18$O4_conv,
                                           seed_A = fecundity18$A4_conv)
}

# original per capita seed production in fecundity18$X4_conv[1,3:8], X = C, S, O, or A
N_4yr_conv_seed_production_per_capita_df <- N_4yr_conv_seed_production_per_capita  %>% 
  unlist(recursive = TRUE) %>% # make a long table
  data.frame() %>% #corn was listed before soybean in the customized f.
  dplyr::rename(seed_per_capita_threshold = ".") %>%
  dplyr::mutate(cohort = rep(c("1", "2", "3", "4", "5", "6"), t*4)) %>% #*4 for 4 phases
   mutate(cycle_no = as.character(rep(1:t, each = 24))) %>% # 24 = 6 cohorts x 4 phases
   mutate(Crop_ID = rep(c(rep("C4", 6), rep("S4", 6), rep("O4", 6), rep("A4", 6)), t)) %>%
  mutate(seed_per_capita_original = rep(c(fecundity18$C4_conv[1,3:8],
                                          fecundity18$S4_conv[1,3:8],
                                          fecundity18$O4_conv[1,3:8],
                                          fecundity18$A4_conv[1,3:8]), t))  %>%
  mutate(seed_production_manipulated = ifelse(seed_per_capita_threshold == seed_per_capita_original, "no", "yes"))

### LOW output


N_4yr_low_seed_production_per_capita <- list() # blank data frame to save loop output 


for (i in 1:t) { 
  N_4yr_low_seed_production_per_capita[[i]] =
    rot_4year_low_seed_production_per_cap(seed_C = fecundity18$C4_low,
                                          seed_S = fecundity18$S4_low,
                                          seed_O = fecundity18$O4_low,
                                          seed_A = fecundity18$A4_low)
}


N_4yr_low_seed_production_per_capita_df <- N_4yr_low_seed_production_per_capita  %>% 
  unlist(recursive = TRUE) %>% # make a long table
  data.frame() %>% #corn was listed before soybean in the customized f.
  dplyr::rename(seed_per_capita_threshold = ".") %>%
  dplyr::mutate(cohort = rep(c("1", "2", "3", "4", "5", "6"), t*4)) %>% #*4 for 4 phases
   mutate(cycle_no = as.character(rep(1:t, each = 24))) %>%
   mutate(Crop_ID = rep(c(rep("C4", 6), rep("S4", 6), rep("O4", 6), rep("A4", 6)), t)) %>%
  mutate(seed_per_capita_original = rep(c(fecundity18$C4_low[1,3:8],
                                          fecundity18$S4_low[1,3:8],
                                          fecundity18$O4_low[1,3:8],
                                          fecundity18$A4_low[1,3:8]), t))  %>%
  mutate(seed_production_manipulated = ifelse(seed_per_capita_threshold == seed_per_capita_original, "no", "yes"))
```


```{r 4yr-sim-a-all-output}
# total stratified seedbank density and lambdas in N_4yr_conv_lambda_df
# manipulated seed production in N_4yr_conv_seed_production_per_capita_df
# fixed mature plant density in N_4yr_conv_mature_plant_density_fixed_df 

N_4yr_conv_all_df <- N_4yr_conv_seed_production_per_capita_df %>% 
  left_join(N_4yr_conv_lambda_df, by = "cycle_no") %>%
  left_join(N_4yr_conv_mature_plant_density_fixed_df, by = c("Crop_ID", "cohort"))  


N_4yr_low_all_df <- N_4yr_low_seed_production_per_capita_df %>% 
  left_join(N_4yr_low_lambda_df, by = "cycle_no") %>%
  left_join(N_4yr_low_mature_plant_density_fixed_df, by = c("Crop_ID", "cohort")) 
```


```{r seed-allowance-combined, include=FALSE}

seed_allowance_sim_df <- rbind(N_2yr_conv_all_df,
                               N_2yr_low_all_df,
                               N_3yr_conv_all_df,
                               N_3yr_low_all_df,
                               N_4yr_conv_all_df,
                               N_4yr_low_all_df)
# check records
# table(seed_allowance_sim_df$Crop_ID, seed_allowance_sim_df$Corn_weed_management)


seed_allowance_sim_df <- seed_allowance_sim_df %>%
  mutate(seed_production_threshold = seed_per_capita_threshold*plant_counts_fixed) %>%
  mutate(seed_production_control_efficacy = 1 - seed_per_capita_threshold/seed_per_capita_original)

# View(seed_allowance_sim_df)

# See if any weird record, such as manipulated seed prod > original seed prod 

seed_allowance_sim_df %>% 
  filter(seed_per_capita_threshold > seed_per_capita_original)
```

```{r seed-production-threshold-sim-plot-manipulated, echo=FALSE, fig.cap= "Cohort-based seed production threshold on natural logarithm scale for waterhemp population stabilization over 100 rotational cycles (the 2-year rotation cycled over two years and ended at the soybean phase, the 3-year rotation cycled over three years and ended at the oat phase, and the 4-year rotation cycled over four years and ended at the alfalfa phase). All simulations started with a seed column of 10000 female seeds in the top 0 - 2 cm soil stratum and 0 female seed in the bottom 2 - 20 cm soil stratum. It was expected that no waterhemp cohorts in any crop environments but only the cohorts 1 through 3 in corn and soybean were manipulated to find the seed production thresholds. However, additional control efficacy was needed in some crop phases outside of the expected group to reduce seed production potentials. The dots colored blue are where control measures extended beyond waterhemp cohort 3 would be necessary. The relationships of aboveground mass and fecundity in Nguyen and Liebman (2022a) were used to estimate per-capita seed production in each cohort. The black horizontal line marks lambda = 1.", fig.height = 6, fig.width = 6}

# names(seed_allowance_sim_df)
seed_allowance_sim_df$Crop_ID <- factor(seed_allowance_sim_df$Crop_ID,
                                        levels = c("C2","S2",
                                                   "C3", "S3", "O3",
                                                   "C4", "S4", "O4", "A4"))

seed_allowance_sim_df %>% 
  na.omit() %>%
     ggplot() +
   aes(x = log(seed_production_threshold), 
       y = lambda_annualized,
       group = seed_production_manipulated) +
   facet_grid(Crop_ID ~ cohort) +
    geom_hline(yintercept = 1, color = "black") +
    scale_color_brewer(palette = "Set1") +
  geom_point(aes(y  = lambda_annualized, 
                 color = seed_production_manipulated,
                 shape = Corn_weed_management)) +
  scale_shape_manual(values = c(1, 4)) +
  ylab("Annualized population growth rate") +
  xlab(expression(paste("Seed production threshold by cohort (ln(seeds)",~m^-2,")"))) + 
  theme_bw() +
  guides(shape = guide_legend(title = "Corn \n weed management"),
         color = guide_legend(title = "Seed production \n manipulated")) +
  theme(legend.position = "top")
```
   



