---
title: 'Q1: how low is seed production allowance - rotation'
output: html_document 
---

## General pattern: In a corn monoculture, aiming at 945 seeds/plant (sd on natural logarithm scale under 0.86) shall keep the original population of 1000 seeds per meter squared stable.

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(readr)
library(magrittr) # for %<>%
library(nlme) # for gls
library(emmeans) # for joint_tests
library(patchwork)
library(pracma) #for nthroot
set.seed(722)
```

### first seed column
```{r}
starting_point <-  c(10000, 0, 0, 0, 0, 0, 0, 0) #10000 seeds top soil, 0 seed deep soil,0 plants from cohorts 1 through 6
```


### tillage matrices  

```{r tillage-sim}
fall_tillage <-readRDS("../2-Data/Clean/mean-post-harvest-tillage.RData")

spring_tillage <-readRDS("../2-Data/Clean/mean-pre-planting-tillage.RData")
```

### overwinter survival matrix  

```{r overwinter-sim}
overwinter <- readRDS("../2-Data/Clean/mean-winter-seed-survival-Sosnoskie.RData")
```

### emergence matrix  

```{r emerge-sim, echo=FALSE}
emergence <-readRDS("../2-Data/Clean/mean-emergence-prop.RData")
emergence$C2_conv
```


### seed survival rate and seedling to maturity success rate 

```{r survive-sim, echo=FALSE}
summer_seed_survival <- readRDS("../2-Data/Clean/mean-summer-seed-survival-Sosnoskie.RData")

#### plant survival: literature data - named as scenario2, because scenario1 is the point-estimates from the experiment 
#female_survival <- readRDS("../2-Data/Clean/female-survival-rate-cohort-equal.RData")

female_survival <- readRDS("../2-Data/Clean/mean-summer-seedling-survival-Hartzler.RData")

#### combine seed and plant survivals into one matrix with element-wise multiplication: this is not computing, but arranging using 1's as placeholders 
summer_survival <- purrr::map2(summer_seed_survival, female_survival, `*`)

summer_survival$C2_conv
```

### fecundity manipulation plan  

```{r fecund-sim}
# First step: use a cumulative fecundity of all cohorts as `
#new_seeds_total`=  mature_plant_total * fecundity_sim
# change fecundity_sim  to get lambda as close to 1 as possible 

# Focus next: manipulate cohorts 1 and 2 in corn and soybean fecundity only because 1) these cohorts are more likely exposed to control measures and 2) they have higher survival rates than cohorts 3+ 

# cohort 3+ fecundity: estimated fecundity as in population projection 

# Final goal: flatten the fecundity distribution curve in corn and soybean to see seed production "allowance" in cohorts 1 and 2

### find good pairs of mean and sd for rlnorm in Q1-fecund-prep

fecundity18 <- readRDS("../2-Data/Clean/mean-fecundity-18-cohort.RData")
```



#### 2-year rotation with corn under conventional weed management {-}

at t = 1: 10000 seeds, cohorts 1 through 3:  rlnorm(1, 3.09, 1.13)

```{r rot-2yr-conv}
# event sequence: seed dropped - chisel - overwinter - field cultivator - emerge - survive - new seed

# create a function 
# vec: starting seed column
# poh: post-harvest tillage
# ow: over winter seed survival
# prt: pre-planting-tillage
# em: emergence
# sv: seed survival rate and seedling to maturity success rate

rot_2year<- function(vec, poh_C, ow_C, prt_C, em_C, sv_C, seed_C,
                           poh_S, ow_S, prt_S, em_S, sv_S, seed_S){
  

  seed_C[1,3] <- rlnorm(1, 3.09, 1.13)
  seed_C[1,4] <- rlnorm(1, 3.09, 1.13)
  seed_C[1,5] <- rlnorm(1, 3.09, 1.13) # original data partitioned into 12 bins
#  seed_C[1,6] <- rlnorm(1, 3.66, 1.44) #original data partitioned into 6 bins


  seed_S[1,3] <- rlnorm(1, 3.09, 1.13)
  seed_S[1,4] <- rlnorm(1, 3.09, 1.13)
  seed_S[1,5] <- rlnorm(1, 3.09, 1.13)
#  seed_S[1,6] <- rlnorm(1, 3.66, 1.44)

 # corn phase dynamics   
   after_corn <- seed_C %*% sv_C %*% em_C %*% prt_C %*%  ow_C %*%  poh_C %*% vec 
# soybean phase dynamics

   after_soy <- seed_S %*% sv_S %*% em_S %*% prt_S %*%  ow_S %*%  poh_S %*% after_corn 

   after_soy
}
```

```{r 2yr-conv-sim1}

t <- 100
N_2yr_conv <- list() # blank data frame to save loop output 
## Calculate N0
N0_2yr_conv  <- rot_2year(vec = starting_point, 
#corn dynamics
   poh_C = fall_tillage$C2_conv,
   ow_C = overwinter$C2_conv,
   prt_C  = spring_tillage$C2_conv,
   em_C  = emergence$C2_conv,
   sv_C = summer_survival$C2_conv,
seed_C = fecundity18$C2_conv,

#soybean dynamics   
  poh_S = fall_tillage$S2_conv,
   ow_S = overwinter$S2_conv,
   prt_S  = spring_tillage$S2_conv,
   em_S  = emergence$S2_conv,
   sv_S = summer_survival$S2_conv,
seed_S = fecundity18$S2_conv)



 N_2yr_conv[[1]] <- N0_2yr_conv 


#loop over the "corn" function for C2-conv for "t" times 

for (i in 2:t) { 
  N_2yr_conv[[i]] = rot_2year(vec = N_2yr_conv[[i-1]],
                              poh_C = fall_tillage$C2_conv,
                              ow_C = overwinter$C2_conv,
                              prt_C  = spring_tillage$C2_conv,
                              em_C  = emergence$C2_conv,
                              sv_C = summer_survival$C2_conv,
                              seed_C = fecundity18$C2_conv,
                              
                              #soybean dynamics   
                              poh_S = fall_tillage$S2_conv,
                              ow_S = overwinter$S2_conv,
                              prt_S  = spring_tillage$S2_conv,
                              em_S  = emergence$S2_conv,
                              sv_S = summer_survival$S2_conv,
                              seed_S = fecundity18$S2_conv)
}

N_2yr_conv_df <-  N_2yr_conv %>% 
  unlist(recursive = FALSE) %>%
  data.frame() %>%
    dplyr::rename(counts = ".") %>%
  dplyr::mutate(category = rep(c("seed_top", "seed_bottom", "cohort_1", "cohort_2", "cohort_3", "cohort_4", "cohort_5", "cohort_6"),t)) %>%
    filter(category %in% c("seed_top", "seed_bottom")) %>%
 # dplyr::select(!rowname) %>%
 # tidyr::pivot_wider(names_from = category,
#                    values_from = no_,
#                     values_fn = list) %>% 
    unnest(cols = everything() ) %>%
    mutate(cycle_no = rep(1:t, each = 2)) %>%
  group_by(category) %>%
  mutate(lambda_cycle = counts/lag(counts),
         lambda_annualized = sqrt(lambda_cycle),
          Rotation = "2-year",
         Corn_weed_management = "conventional") %>%
    na.omit()
```
 

 
```{r 2yr-low-sim1}
N_2yr_low <- list() # blank dataframe to save loop output 
## Calculate N0
N0_2yr_low  <- rot_2year(vec = starting_point, 
                         #corn dynamics
                         poh_C = fall_tillage$C2_low,
                         ow_C = overwinter$C2_low,
                         prt_C  = spring_tillage$C2_low,
                         em_C  = emergence$C2_low,
                         sv_C = summer_survival$C2_low,
                         seed_C = fecundity18$C2_low,
                         
                         #soybean dynamics   
                         poh_S = fall_tillage$S2_low,
                         ow_S = overwinter$S2_low,
                         prt_S  = spring_tillage$S2_low,
                         em_S  = emergence$S2_low,
                         sv_S = summer_survival$S2_low,
                         seed_S = fecundity18$S2_low)



 N_2yr_low[[1]] <- N0_2yr_low


#loop over the "corn" function for C2-conv for "t" times 

for (i in 2:t) { 
  N_2yr_low[[i]] = rot_2year(vec = N_2yr_low[[i-1]],
                             poh_C = fall_tillage$C2_low,
                             ow_C = overwinter$C2_low,
                             prt_C  = spring_tillage$C2_low,
                             em_C  = emergence$C2_low,
                             sv_C = summer_survival$C2_low,
                             seed_C = fecundity18$C2_low,
                             
                             #soybean dynamics   
                             poh_S = fall_tillage$S2_low,
                             ow_S = overwinter$S2_low,
                             prt_S  = spring_tillage$S2_low,
                             em_S  = emergence$S2_low,
                             sv_S = summer_survival$S2_low,
                             seed_S = fecundity18$S2_low)
}

N_2yr_low_df <-  N_2yr_low %>% 
  unlist(recursive = FALSE) %>%
  data.frame() %>%
    dplyr::rename(counts = ".") %>%
  dplyr::mutate(category = rep(c("seed_top", "seed_bottom", "cohort_1", "cohort_2", "cohort_3", "cohort_4", "cohort_5", "cohort_6"),t)) %>%
    filter(category %in% c("seed_top", "seed_bottom")) %>%
    unnest(cols = everything() ) %>%
    mutate(cycle_no = rep(1:t, each = 2)) %>%
  group_by(category) %>%
  mutate(lambda_cycle = counts/lag(counts),
         lambda_annualized = sqrt(lambda_cycle),
          Rotation = "2-year",
         Corn_weed_management = "low") %>%
    na.omit()

#View(N_2yr_low_df)
```
  
#### 3-year rotation {-}

at t = 1: 10000 seeds, cohorts 1 through 3:  rlnorm(1, 3.09, 1.13)

```{r rot-3yr-function}
rot_3year <- function(vec, poh_C, ow_C, prt_C, em_C, sv_C,  seed_C, 
                           poh_S, ow_S, prt_S, em_S, sv_S, seed_S ,
                           poh_O, ow_O, prt_O, em_O, sv_O, seed_O){
  

  seed_C[1,3] <- rlnorm(1, 3.09, 1.13)
  seed_C[1,4] <- rlnorm(1, 3.09, 1.13)
  seed_C[1,5] <- rlnorm(1, 3.09, 1.13) 


  seed_S[1,3] <- rlnorm(1, 3.09, 1.13)
  seed_S[1,4] <- rlnorm(1, 3.09, 1.13)
  seed_S[1,5] <- rlnorm(1, 3.09, 1.13)

# corn phase dynamics  
   after_corn <- seed_C %*% sv_C %*% em_C %*% prt_C %*%  ow_C %*%  poh_C %*% vec 
 # soybean phase dynamics
   after_soy <- seed_S %*% sv_S %*% em_S %*% prt_S %*%  ow_S %*%  poh_S %*% after_corn
# oat phase dynamics
   after_oat <- seed_O %*% sv_O %*% em_O %*% prt_O %*%  ow_O %*%  poh_O %*% after_soy 
   
  after_oat
}
```
##### with corn under conventional weed management {-}
```{r 3yr-conv-sim1}

N_3yr_conv <- list() # blank dataframe to save loop output 
## Calculate N0
N0_3yr_conv  <- rot_3year(vec = starting_point, 
                          #corn dynamics
                          poh_C = fall_tillage$C3_conv,
                          ow_C = overwinter$C3_conv,
                          prt_C  = spring_tillage$C3_conv,
                          em_C  = emergence$C3_conv,
                          sv_C = summer_survival$C3_conv,
                          seed_C = fecundity18$C3_conv,
                          
                          #soybean dynamics   
                          poh_S = fall_tillage$S3_conv,
                          ow_S = overwinter$S3_conv,
                          prt_S  = spring_tillage$S3_conv,
                          em_S  = emergence$S3_conv,
                          sv_S = summer_survival$S3_conv,
                          seed_S = fecundity18$S3_conv,
                          
                          #oat dynamics   
                          poh_O = fall_tillage$O3_conv,
                          ow_O = overwinter$O3_conv,
                          prt_O  = spring_tillage$O3_conv,
                          em_O  = emergence$O3_conv,
                          sv_O = summer_survival$O3_conv,
                          seed_O = fecundity18$O3_conv)


 N_3yr_conv[[1]] <- N0_3yr_conv 


for (i in 2:t) { 
  N_3yr_conv[[i]] = rot_3year(vec = N_3yr_conv[[i-1]],
                              poh_C = fall_tillage$C3_conv,
                              ow_C = overwinter$C3_conv,
                              prt_C  = spring_tillage$C3_conv,
                              em_C  = emergence$C3_conv,
                              sv_C = summer_survival$C3_conv,
                              seed_C = fecundity18$C3_conv,
                              
                              #soybean dynamics   
                              poh_S = fall_tillage$S3_conv,
                              ow_S = overwinter$S3_conv,
                              prt_S  = spring_tillage$S3_conv,
                              em_S  = emergence$S3_conv,
                              sv_S = summer_survival$S3_conv,
                              seed_S = fecundity18$S3_conv,
                              
                              #oat dynamics   
                              poh_O = fall_tillage$O3_conv,
                              ow_O = overwinter$O3_conv,
                              prt_O  = spring_tillage$O3_conv,
                              em_O  = emergence$O3_conv,
                              sv_O = summer_survival$O3_conv,
                              seed_O = fecundity18$O3_conv)
}

N_3yr_conv_df <-  N_3yr_conv %>% 
  unlist(recursive = FALSE) %>%
  data.frame() %>%
    dplyr::rename(counts = ".") %>%
  dplyr::mutate(category = rep(c("seed_top", "seed_bottom", "cohort_1", "cohort_2", "cohort_3", "cohort_4", "cohort_5", "cohort_6"),t)) %>%
    filter(category %in% c("seed_top", "seed_bottom")) %>%
    unnest(cols = everything() ) %>%
    mutate(cycle_no = rep(1:t, each = 2)) %>%
  group_by(category) %>%
  mutate(lambda_cycle = counts/lag(counts),
         lambda_annualized = nthroot(lambda_cycle,3),
          Rotation = "3-year",
         Corn_weed_management = "conventional") %>%
    na.omit()
```
 
##### with corn under low herbicide weed management {-} 
```{r 3yr-low-sim1}

N_3yr_low <- list() # blank dataframe to save loop output 
## Calculate N0
N0_3yr_low  <- rot_3year(vec = starting_point, 
                         #corn dynamics
                         poh_C = fall_tillage$C3_low,
                         ow_C = overwinter$C3_low,
                         prt_C  = spring_tillage$C3_low,
                         em_C  = emergence$C3_low,
                         sv_C = summer_survival$C3_low,
                         seed_C = fecundity18$C3_low,
                         
                         #soybean dynamics   
                         poh_S = fall_tillage$S3_low,
                         ow_S = overwinter$S3_low,
                         prt_S  = spring_tillage$S3_low,
                         em_S  = emergence$S3_low,
                         sv_S = summer_survival$S3_low,
                         seed_S = fecundity18$S3_low,
                         
                         #oat dynamics   
                         poh_O = fall_tillage$O3_low,
                         ow_O = overwinter$O3_low,
                         prt_O  = spring_tillage$O3_low,
                         em_O  = emergence$O3_low,
                         sv_O = summer_survival$O3_low,
                         seed_O = fecundity18$O3_low)

 N_3yr_low[[1]] <- N0_3yr_low


for (i in 2:t) { 
  N_3yr_low[[i]] = rot_3year(vec = N_3yr_low[[i-1]],
                             poh_C = fall_tillage$C3_conv,
                             ow_C = overwinter$C3_low,
                             prt_C  = spring_tillage$C3_low,
                             em_C  = emergence$C3_low,
                             sv_C = summer_survival$C3_low,
                             seed_C = fecundity18$C3_low,
                             
                             #soybean dynamics   
                             poh_S = fall_tillage$S3_low,
                             ow_S = overwinter$S3_low,
                             prt_S  = spring_tillage$S3_low,
                             em_S  = emergence$S3_low,
                             sv_S = summer_survival$S3_low,
                             seed_S = fecundity18$S3_low,
                             
                             #oat dynamics   
                             poh_O = fall_tillage$O3_low,
                             ow_O = overwinter$O3_low,
                             prt_O  = spring_tillage$O3_low,
                             em_O  = emergence$O3_low,
                             sv_O = summer_survival$O3_low,
                             seed_O = fecundity18$O3_low)
}

N_3yr_low_df <-  N_3yr_low %>% 
  unlist(recursive = FALSE) %>%
  data.frame() %>%
    dplyr::rename(counts = ".") %>%
  dplyr::mutate(category = rep(c("seed_top", "seed_bottom", "cohort_1", "cohort_2", "cohort_3", "cohort_4", "cohort_5", "cohort_6"),t)) %>%
    filter(category %in% c("seed_top", "seed_bottom")) %>%
    unnest(cols = everything() ) %>%
    mutate(cycle_no = rep(1:t, each = 2)) %>%
  group_by(category) %>%
  mutate(lambda_cycle = counts/lag(counts),
         lambda_annualized = nthroot(lambda_cycle,3),
          Rotation = "3-year",
         Corn_weed_management = "low") %>%
    na.omit()
```


#### 4-year rotation {-}

at t = 1: 10000 seeds, cohorts 1 through 3:  rlnorm(1, 3.09, 1.13)

```{r rot-4yr-function}
rot_4year <- function(vec, poh_C, ow_C, prt_C, em_C, sv_C,  seed_C, 
                           poh_S, ow_S, prt_S, em_S, sv_S, seed_S ,
                           poh_O, ow_O, prt_O, em_O, sv_O, seed_O,
                       poh_A, ow_A, prt_A, em_A, sv_A, seed_A){
  
  seed_C[1,3] <- rlnorm(1, 3.09, 1.13)
  seed_C[1,4] <- rlnorm(1, 3.09, 1.13)
  seed_C[1,5] <- rlnorm(1, 3.09, 1.13) 


  seed_S[1,3] <- rlnorm(1, 3.09, 1.13)
  seed_S[1,4] <- rlnorm(1, 3.09, 1.13)
  seed_S[1,5] <- rlnorm(1, 3.09, 1.13)

# corn phase dynamics
   after_corn <- seed_C %*% sv_C %*% em_C %*% prt_C %*%  ow_C %*%  poh_C %*% vec
# soybean phase dynamics
   after_soy <- seed_S %*% sv_S %*% em_S %*% prt_S %*%  ow_S %*%  poh_S %*% after_corn
# oat phase dynamics
after_oat <- seed_O %*% sv_O %*% em_O %*% prt_O %*%  ow_O %*%  poh_O %*% after_soy 
# alfalfa phase dynamics
after_alfalfa <- seed_A %*% sv_A %*% em_A %*% prt_A %*%  ow_A %*%  poh_A %*% after_oat 
   
  after_alfalfa
}
```
##### with corn under conventional weed management {-}
```{r 4yr-conv-sim1}
N_4yr_conv <- list() # blank dataframe to save loop output 
## Calculate N0
N0_4yr_conv  <- rot_4year(vec = starting_point, 
                          #corn dynamics
                          poh_C = fall_tillage$C4_conv,
                          ow_C = overwinter$C4_conv,
                          prt_C  = spring_tillage$C4_conv,
                          em_C  = emergence$C4_conv,
                          sv_C = summer_survival$C4_conv,
                          seed_C = fecundity18$C4_conv,
                          
                          #soybean dynamics   
                          poh_S = fall_tillage$S4_conv,
                          ow_S = overwinter$S4_conv,
                          prt_S  = spring_tillage$S4_conv,
                          em_S  = emergence$S4_conv,
                          sv_S = summer_survival$S4_conv,
                          seed_S = fecundity18$S4_conv,
                          
                          #oat dynamics   
                          poh_O = fall_tillage$O4_conv,
                          ow_O = overwinter$O4_conv,
                          prt_O  = spring_tillage$O4_conv,
                          em_O  = emergence$O4_conv,
                          sv_O = summer_survival$O4_conv,
                          seed_O = fecundity18$O4_conv,
                          
                          #alfalfa dynamics   
                          poh_A = fall_tillage$A4_conv,
                          ow_A = overwinter$A4_conv,
                          prt_A  = spring_tillage$A4_conv,
                          em_A  = emergence$A4_conv,
                          sv_A = summer_survival$A4_conv,
                          seed_A = fecundity18$A4_conv)

# N0

N_4yr_conv[[1]] <- N0_4yr_conv 


for (i in 2:t) { 
  N_4yr_conv[[i]] = rot_4year(vec = N_4yr_conv[[i-1]],
                              poh_C = fall_tillage$C4_conv,
                              ow_C = overwinter$C4_conv,
                              prt_C  = spring_tillage$C4_conv,
                              em_C  = emergence$C4_conv,
                              sv_C = summer_survival$C4_conv,
                              seed_C = fecundity18$C4_conv,
                              
                              #soybean dynamics   
                              poh_S = fall_tillage$S4_conv,
                              ow_S = overwinter$S4_conv,
                              prt_S  = spring_tillage$S4_conv,
                              em_S  = emergence$S4_conv,
                              sv_S = summer_survival$S4_conv,
                              seed_S = fecundity18$S4_conv,
                              
                              #oat dynamics   
                              poh_O = fall_tillage$O4_conv,
                              ow_O = overwinter$O4_conv,
                              prt_O  = spring_tillage$O4_conv,
                              em_O  = emergence$O4_conv,
                              sv_O = summer_survival$O4_conv,
                              seed_O = fecundity18$O4_conv,
                              
                              #alfalfa dynamics   
                          poh_A = fall_tillage$A4_conv,
                          ow_A = overwinter$A4_conv,
                          prt_A  = spring_tillage$A4_conv,
                          em_A  = emergence$A4_conv,
                          sv_A = summer_survival$A4_conv,
                          seed_A = fecundity18$A4_conv)
}

N_4yr_conv_df <-  N_4yr_conv %>% 
  unlist(recursive = FALSE) %>%
  data.frame() %>%
  dplyr::rename(counts = ".") %>%
  dplyr::mutate(category = rep(c("seed_top", "seed_bottom", "cohort_1", "cohort_2", "cohort_4", "cohort_4", "cohort_5", "cohort_6"),t)) %>%
  filter(category %in% c("seed_top", "seed_bottom")) %>%
  unnest(cols = everything() ) %>%
  mutate(cycle_no = rep(1:t, each = 2)) %>%
  group_by(category) %>%
  mutate(lambda_cycle = counts/lag(counts),
         lambda_annualized = nthroot(lambda_cycle,4),
         Rotation = "4-year",
         Corn_weed_management = "conventional") %>%
  na.omit()
```
 
##### with corn under low herbicide weed management {-} 
```{r 4yr-low-sim1}
N_4yr_low <- list() # blank dataframe to save loop output 
## Calculate N0
N0_4yr_low  <- rot_4year(vec = starting_point, 
                          #corn dynamics
                          poh_C = fall_tillage$C4_low,
                          ow_C = overwinter$C4_low,
                          prt_C  = spring_tillage$C4_low,
                          em_C  = emergence$C4_low,
                          sv_C = summer_survival$C4_low,
                          seed_C = fecundity18$C4_low,
                          
                          #soybean dynamics   
                          poh_S = fall_tillage$S4_low,
                          ow_S = overwinter$S4_low,
                          prt_S  = spring_tillage$S4_low,
                          em_S  = emergence$S4_low,
                          sv_S = summer_survival$S4_low,
                          seed_S = fecundity18$S4_low,
                          
                          #oat dynamics   
                          poh_O = fall_tillage$O4_low,
                          ow_O = overwinter$O4_low,
                          prt_O  = spring_tillage$O4_low,
                          em_O  = emergence$O4_low,
                          sv_O = summer_survival$O4_low,
                          seed_O = fecundity18$O4_low,
                          
                          #alfalfa dynamics   
                          poh_A = fall_tillage$A4_low,
                          ow_A = overwinter$A4_low,
                          prt_A  = spring_tillage$A4_low,
                          em_A  = emergence$A4_low,
                          sv_A = summer_survival$A4_low,
                          seed_A = fecundity18$A4_low)

# N0

N_4yr_low[[1]] <- N0_4yr_low 


for (i in 2:t) { 
  N_4yr_low[[i]] = rot_4year(vec = N_4yr_low[[i-1]],
                              poh_C = fall_tillage$C4_low,
                              ow_C = overwinter$C4_low,
                              prt_C  = spring_tillage$C4_low,
                              em_C  = emergence$C4_low,
                              sv_C = summer_survival$C4_low,
                              seed_C = fecundity18$C4_low,
                              
                              #soybean dynamics   
                              poh_S = fall_tillage$S4_low,
                              ow_S = overwinter$S4_low,
                              prt_S  = spring_tillage$S4_low,
                              em_S  = emergence$S4_low,
                              sv_S = summer_survival$S4_low,
                              seed_S = fecundity18$S4_low,
                              
                              #oat dynamics   
                              poh_O = fall_tillage$O4_low,
                              ow_O = overwinter$O4_low,
                              prt_O  = spring_tillage$O4_low,
                              em_O  = emergence$O4_low,
                              sv_O = summer_survival$O4_low,
                              seed_O = fecundity18$O4_low,
                              
                              #alfalfa dynamics   
                              poh_A = fall_tillage$A4_low,
                              ow_A = overwinter$A4_low,
                              prt_A  = spring_tillage$A4_low,
                              em_A  = emergence$A4_low,
                              sv_A = summer_survival$A4_low,
                              seed_A = fecundity18$A4_low)
}

N_4yr_low_df <-  N_4yr_low %>% 
  unlist(recursive = FALSE) %>%
  data.frame() %>%
  dplyr::rename(counts = ".") %>%
  dplyr::mutate(category = rep(c("seed_top", "seed_bottom", "cohort_1", "cohort_2", "cohort_4", "cohort_4", "cohort_5", "cohort_6"),t)) %>%
  filter(category %in% c("seed_top", "seed_bottom")) %>%
  unnest(cols = everything() ) %>%
  mutate(cycle_no = rep(1:t, each = 2)) %>%
  group_by(category) %>%
  mutate(lambda_cycle = counts/lag(counts),
         lambda_annualized = nthroot(lambda_cycle,4),
         Rotation = "4-year",
         Corn_weed_management = "low") %>%
  na.omit()
```

```{r seed-allowance-sim-plot}
seed_allowance_sim_df <- rbind(N_2yr_conv_df,
                                N_2yr_low_df,
                               N_3yr_conv_df,
                                N_3yr_low_df,
                               N_4yr_conv_df,
                                N_4yr_low_df)

seed_allowance_sim_df %>%
   ggplot() +
   aes(x = cycle_no, y = lambda_annualized, group = category) +
  geom_line(aes(y = lambda_annualized, linetype =  category)) +
  geom_point(aes(y  = lambda_annualized, shape =  category)) +
  geom_hline(yintercept = 1, color = "red") +
  facet_wrap(Corn_weed_management ~ Rotation)
```