---
title: "Effects of crop rotation on common waterhemp population dynamics: prospective and retrospective analyses"
output:
    bookdown::word_document2:
      toc: false
      fig_caption: yes
      keep_md: true
bibliography: WH-pop-dynamics.bib
csl: apa-no-ampersand.csl 
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(purrr) # for map2()
library(pracma)
library(popbio)
library(tidyverse)
library(magrittr) # for %<>%
# convert this Rmd to R to source it in results.Rmd
library(flextable)
library(patchwork) #for wrap_plots
library(ggResidpanel)
library(broom)
library(nlme)
source("functions.R")
theme_set(theme_bw())
```

# Abstract {-}

```{r, child = "../5-Report/Abstract-model.Rmd"}
```


# Introduction {-}  

```{r, child = "../5-Report/Introduction-model.Rmd"}
```

# Materials and methods {-}

```{r, child = "../5-Report/M-and-M-model.Rmd"}
```

# Results and discussion {-}
All the multi-year variables failed the Levene's test for equal variance so those data were analyzed by year.  

## Waterhemp sub-annual demographic parameters from empirical measurement {-}

### Seedling emergence pattern and timing {-}

Within the same calendar year, cumulative whole-season and cohort-based seedling densities were lower in the warm-season crops than in the cool-season crops (Figure \@ref(fig:seedlings-density-and-cumu-prop-2019-2020)). Seedling emergence was delayed in the cool-season crops as compared to that in the warm-season crops (Tables \@ref(tab:Julian-19-20-emm) and \@ref(tab:Julian-19-20-ct-tab)). Waterhemp emergence was delayed by two weeks to a month in the alfalfa crop environment as compared to the corn and soybean environments. Waterhemp emergence was nine to sixteen days earlier in oat than in alfalfa.    

```{r, child= "~/Documents/Dissertation/AMATA-demography/4-Analysis/emergence-pattern-timing-Julian.Rmd"}
# diagnosis plots still look concerning
```

### 2019 soil seedbank and emergence proportion {-} 


```{r, child= "~/Documents/Dissertation/AMATA-demography/5-Report/seedbank-top-bottom.Rmd"}
# , child = "female-dens-mature.Rmd" removed because density was published
```


## Waterhemp population growth rates {-}

All the data sets for the empirically estimated sub-annual parameters were measured once. In 2019, individual plant size at maturity was measured for six cohorts in each quadrat of each eu and two stratum of soil seedbank were estimated between crop harvest and post-harvest tillage. In 2018, a different number of cohorts were followed in different crop environments because of limited labor availability and adverse weather conditions. 2019 data was used as scenario 1. Within the same rotation, a consistent pattern of top:bottom ratios of seed densities was observed between two corn weed management regimes. The seedbanks' top : bottom partitioning differed between rotations and between crop phases of the same rotation. 

### Scenario 1 {-}
Using 2019 fecundity rates that were estimated from individual plant size using eighteen equations from @nguyenImpactCroppingSystem2022, waterhemp population densities were projected to decrease in all treatments, except in the 4-year rotation with low herbicide corn weed management ($\lambda(4,low)$ = 1.4)  (Figure \@ref(fig:scenario1-phase-wise-lambda-plot)). Waterhemp populations were projected to decrease the most quickly in the 3-year and 4-year rotations under conventional herbicide corn weed management. 

The top and bottom soil strata were evenly populated at the completion of the corn phase in all rotations crossed with corn weed management treatments, but a smaller portion of the seedbank was in the top soil stratum than in the bottom soil stratum in the soybean phase of all the rotations crossed with corn weed management treatments, and in the alfalfa phase of the 4-year rotation regardless of the corn weed management regimes. The top soil stratum was less populated than the bottom soil stratum at the completion of O3 but more populated than the bottom soil stratum at the completion of O4. 


With the exception of the 4-year rotation in which corn was treated with low herbicide weed management, all the other treatments' decline of waterhemp population was attributed to the steady decrease of population size in the crop phases after corn. Even though the 2-year and 3-year rotations were of similar annualized population growth ($\lambda(2,conv) = 0.8$, $\lambda(2,low) = 0.9$, $\lambda(3,conv) = 0.7$, and $\lambda(3,low) = 0.9$), larger proportion of the seedbank was buried in the bottom soil stratum in the 3-year rotation, which can shrink the available seed pool for germination. 

### Scenario 2 {-}

Using 2018 fecundity rates, waterhemp population densities would increase rapidly in all rotations (Figure \@ref(fig:scenario2-phase-wise-lambda-plot)) , but most quickly in the 2-year rotation ($\lambda(2,conv) = 59.4$ and $\lambda(2,low) = 333.3$). In the more diverse rotations (3-year and 4-year), waterhemp population increased at lower annualized rates if the corn phase was treated with low herbicide weed management ($\lambda(3,conv) = 48.4$, $\lambda(3,low) = 28.5$, $\lambda(4,conv) = 14.7$ and $\lambda(4,low) = 10.5$). This trend suggests possible elevated herbicide resistance in the treatments in which corn was managed with conventional herbicide, as organisms developed resistance to the strongest selection factor against them.


The top and bottom soil strata were fairly evenly populated at the completion of the corn phase, but a larger portion of the seedbanks were in the top soil stratum at the completion of the soybean phase in the 2-year and 3-year rotations (S2 and S3). The top : bottom soil stratum seed density partition in the soybean phase of the 4-year rotation (S4) was on-and-off, in which the top stratum was projected to be more populated in one model cycle (four years) but less populated in another model cycle. The top soil stratum was less populated than the bottom soil stratum at the completion of O3, but more population than the bottom soil stratum at the completion of O4. The top soil stratum in the alfalfa phase was more populated than the bottom soil stratum. The alfalfa phase of the 4-year rotation (A4) offered an opportunity for decreasing $\lambda$, which was not possible in any other crop phases in the three examined rotations. However, the declines in and A4 ($\lambda(A4,conv)$ = 0.6 and $\lambda(A4,low)$ = 0.5) were not strong enough to deplete the replenishment from the corn ($\lambda(C4,conv)$ = 29.7, and $\lambda(C4,low)$ = 87.3) and soybean phases ($\lambda(S4,conv)$ = 773.7, and $\lambda(S4,low)$ = 92.1). 

## Elasticities of popilation growth rates to lower-level demographic parameters {-} **remove**

All the relative ranking of elasticity of $\lambda$ to the lower-level demographic parameters was consistent between corn weed management regimes (Figures \@ref(fig:scenario1-E-B-by-rot-rank) and \@ref(fig:scenario2-E-B-by-rot-rank)). However, the overall population change rate can be reflected by the relative importance of each $b_{ij}$ element on the elasticity of $\lambda$. 

Even though all the $B_h$ matrices, but $B_f$, were identical across scenarios, the ranking of each $b_{ij}$ element contribution to the elasticity of $\lambda$ might be slightly different across scenarios. For example, the the $\lambda$-decreasing case, the relative ranks of $t_{22_s}$ (the probability that a seed in the 2-20 cm stratum stay at that stratum) was higher that those of other $t_{ij_s}$, whereas $t_{22_s}$ in the population-growing case was the lowest among all $t_{ij_s}$. Similarly, $s_{s2}$ was more important than $s_{s1}$ in the elasticity of $\lambda$ in the population-shrinking versus population-growing case. Consistent patterns in the importance of $t_{ij_f}$ to elasticity of $\lambda$ was observed in the two scenarios.  

Because seed emergence and seedling survival rate contributed minimally to the elasticity of $\lambda$, this pattern suggests that even if weed control programs are of high, or even total efficacy, very few surviving plants prolifically producing new seeds, can sustain the population. The low relative importance of $g_k$ and $s_{p_k}$ can be attributed to the low emergence rate in weeds, the high weed control efficacy, while the high relative importance of of $f_k$ to elasticity of $\lambda$ can be explained by the high individual fecundity.  


In both scenarios, the patterns of contribution to elasticity of $\lambda$ in the 3-year rotation was slightly different from the 2-year and 4-year rotations. The contribution of plant survival rate to elasticity of $\lambda$ was non-existent in C2, S2, C4, S4, and O4 (absolute zeroes) but minimal in C3, S3, and O4 (very small positive values). The difference in contribution of plant survival rate on O3 and O4 might be attributed to the different companion crops to oat (red clover in O3 and alfalfa in O4). 

### Scenario 1 {-}  
The ranking of parameters or individual parameters to the elasticity of $\lambda_r$ differed by rotation and crop phase (Figure \@ref(fig:scenario1-E-B-by-rot-rank)). The seed preserving parameters were ranked higher in the corn phase of the 3-year rotation (C3) under both weed management regimes, the soybean phase of the 3-year rotation (S3) that followed corn under low herbicide weed management, the oat phases of the 3-year and 4-year rotations (O3 and O4) regardless of the preceding corn phase's weed management regime, and the alfalfa phase of the 4-year rotation (A4) regardless of the preceding corn phase's weed management regime. The seed producing parameters were ranked higher in the other crop environments (crop species in each rotation).  

The elasticity of $\lambda(r)$ to fecundity rate, as a group, were ranked first in the corn phase of the 2-year rotation (C2) (43.13 and 3.15), the soybean phase of the 2-year rotation (S2) (0.75 and 4.18), the soybean phase of the 3-year rotation (S3) that followed corn under low herbicide weed management program (0.91), the corn phase of the 4-year rotation (C4) (305.24 and 323.07), and the alfalfa in the 4-year rotation (A4) (36.78 and 14.49) under both weed management regimes.
The rate of pre-planting tillage induced seed movement to the 2-20 cm soil stratum was the first-ranked individual parameter to the elasticity of $\lambda(r)$'s in the corn phase of the 3-year rotation (C3) (1.1 and 1.1), the soybean in the 3-year rotation that followed corn under low herbicide weed management (0.89), the oat phase of the 3-year rotation (O3) (0.87 and 0.88), and the oat phase of the 4-year rotation (5.49 and 3.98). The overwinter survival rate of the seeds in the 0-2 cm soil stratum was the first-ranked individual parameter to the elasticity of $\lambda(r)$'s in the soybean phase of the 4-year rotation (S4) (3.32 and 2.41). 

As $\lambda(r)$s declined in this scenario, with some opportunity of increasing in the C4 phase, focusing on reducing plant fecundity in C4 would sufficiently stabilize $\lambda(4)$, provided that the weed control efficacy in the subsequent three crop phases remained adequate. 


### Scenario 2 {-}  

Unlike Scenario 1's dynamics, the differences in the absolute value of the elasticity of $\lambda_r$ to the first-ranked and second-ranked groups of parameters were multiple orders of magnitude. The seed producing parameters were ranked higher in all crop environments (crop species in each rotation).   

The elasticity of $\lambda(r)$'s to fecundity rates, as a group, were ranked first in all crop phases (Figure \@ref(fig:scenario2-E-B-by-rot-rank)). The elasticity of $\lambda$'s to cohort one fecundity were the highest in C2 (4.02 x $10^6$ and 3.44 x $10^6$), S3 (1.33 x $10^6$ and 1.61 x $10^6$), and O4 (682.28 and 1531.97) phases. The elasticity of $\lambda_r$'s elasticity to cohort two fecundity were the highest in S2 (1.82 x $10^{7}$ and 1.39 x $10^7$), C3 (2.22 x $10^6$ and 5.69 x $10^5$), O3 (76673.94 and 	
62845.84), C4 (5.84 x $10^7$ and 1.01 x $10^6$), S4 (5.74 x $10^6$ and  2.17 x $10^6$), and A4 (1850.3 and 156.54) phases.  

The second-ranked group of demographic parameters to $\lambda_r$ differed by rotation and crop phase. In the 2-year rotation, the second-ranked group of parameters in contribution to the elasticity of $\lambda(2)$ were both pre-planting tillage-induced seed movement and overwinter seed survival. In the 3-year rotation, no clear pattern of the second-ranked group of parameters in contribution to elasticity of $\lambda(3)$ was observed: overwinter survival rate of the top stratum seeds in C3, non-germinating seeds in the top stratum in S3, and overwinter survival rate of the bottom stratum seeds in O3. In the 4-year rotation, the second-ranked group of parameters in contribution to elasticity of $\lambda(4)$ was overwinter seed survival in the warm-season crop phase and non-germinating seeds in the top stratum in the cool-season crop phases.



```{r, child = "prospective-dynamics-rotation-wise-projection-scenario1.Rmd"}
```


```{r, child = "prospective-dynamics-rotation-wise-projection-scenario2.Rmd"}
```

## Simulation {-}

The more diverse rotations (3-year and 4-year) can tolerate higher seed production than the 2-year rotation can (Figure \@ref(fig:seed-allowance-sim-plot)). This pattern is consistent with the population projection presented previously.  

Using the population dynamics demonstrated in Scenario 1, an initial population density of 10000 seeds/m$^2$ would remain stable if the first three cohorts in C2 and S2 produced 42 seeds/m$^2$ in total (or 0.04 gram/$^2$ of dried aboveground mass), which would be equivalent to almost total control.  

The 4-year rotation can tolerate bigger plants in the corn and soybean phases than the 3-year rotation, but the different was more substantial under low herbicide corn weed management. When corn was managed with conventional herbicide, 4821 seeds/m$^2$ in C4 versus 1101 seeds/m$^2$ in C3 and 3017 seeds/m$^2$ in S4 versus 2531 seeds/m$^2$ in S3 was tolerated. Within the seed production allowance under conventional corn weed management, 2187 seeds/m$^2$ could be tolerated in C4's and S4's first three waterhemp cohorts but 627 seeds/m$^2$ could be tolerated in C3's  and S3's first three waterhemp cohorts. However, when low herbicide regime was used for corn weed management, 2394 seeds/m$^2$ in C4 versus 889 seeds/m$^2$ in C3 and 2672 seeds/m$^2$ in S4 versus 785 seeds/m$^2$ in S3 was tolerated. Within the seed production allowance under low herbicide corn weed management, 2187 seeds/m$^2$ could be tolerated in C4's and S4's first three waterhemp cohorts but 627 seeds/m$^2$ could be tolerated in C3's and S3's first three waterhemp cohorts.



```{r, child = "Q1-seed-production-allowance-rot.Rmd"}
```

## Conclusion {-}

Delayed but steady emergence (Tables \@ref(tab:Julian-19-20-emm) and \@ref(tab:Julian-19-20-ct-tab)), Figure \@ref(fig:seedlings-density-and-cumu-prop-2019-2020)) in the cool-season crop environments decreased population fecundity in the oat and alfalfa phases, and thus, accelerated seedbank size reduction in scenario 1 and slowed seedbank size increment in scenario 2. 

Opportunities to deplete and risks to replenish the soil seedbank are intertwined as illustrated in the population trajectories in Scenario 1's 4-year rotation with corn managed with low herbicide and in Scenario 2's all trajectories. The soil seedbank density's response to the oat crop environment was unstable and dependent upon oat's establishment performance. The oat in our experiment did not establish well and was hail-damaged soon before harvest in 2018 (Scenario 2's input), but established well and was not damaged in 2019 (Scenario 1's input) and this performance was reflected in the projected phase-wise $\lambda$ (Figures \@ref(fig:scenario1-phase-wise-lambda-plot) and \@ref(fig:scenario2-phase-wise-lambda-plot)). The seedbank's more stable and predictable response to alfalfa crop environment effect could be attributed to alfalfa's ability to form dense canopy.  


The hypothesis that "extending a conventional 2-year rotation of corn and soybean with cool-season crops can accelerate soil seedbank depletion" was supported. Unlike giant foxtail whose bottle neck point in the life cycle was overwinter seed survival rate [@davisCroppingSystemEffects2002], waterhemp's bottle neck point in its life cycle was seedling emergence rate and seedling to maturity success rate. However, if the early emerged plants were unaffected by weed control programs, either through herbicide resistance or escape from cultivation, and successfully complete their life cycle, a small number of large, prolific female plants would sufficiently replenish the soil seedbank.  

Since the elasticity of $\lambda$ was the least affected by seedling emergence in both the population-increasing and population-declining scenarios, it would be helpful to focus on disrupting seed production by reducing individual plant size by delaying emergence or applying multiple stress factors to successfully established plants, and by limiting seed deposit to the soil. Under the current 4-year rotation's crop sequence, oat (O4) and alfalfa (A4) provided valuable opportunities to reduce plant size [@nguyenWeedCommunityComposition2022; @nguyenImpactCroppingSystem2022]. 

A reduction in the mass of applied herbicide active ingredients [@nguyenWeedCommunityComposition2022] was not coincident with population declining in the Scenario 1 ($\lambda < 1$) but did coincide with population increases in scenario 2 ($\lambda > 1$). In consideration of Scenario 2, population-increasing, it would be useful to examine how many years of continuous overwinter crops would be necessary and which cool-season crop species would be most efficient in decreasing waterhemp $\lambda$ after steady seedbank replenishment. Such an investigation is needed because even if the increased abundance of waterhemp and other weed species has not been observed at the experiment site [@nguyenWeedCommunityComposition2022], an abundant seedbank of a highly competitive weed species harbors risks of weed outbreaks.

The weed control pressure for the corn and soybean phases could be lessened with more extended rotations as bigger waterhemp plants in cohorts 1 through 3 can be tolerated in the 3-year and 4-year rotation than in the 2-year rotation. This higher tolerance can be attributed to the weed control effects provided by extending a cropping system of corn and soybean only with cool-season crops. As waterhemp fecundity was manipulated for cohorts 1 through 3 in corn and soybean environments only, it is implied that if effort was made to severely suppress waterhemp in corn and soybean phases, no additional weed management may be needed in the oat and alfalfa phases, so even the big waterhemp plants in oat and alfalfa phases would not need attention. Practitioners who chose the 4-year rotation for the economic and ecological benefits it offers [@davisIncreasingCroppingSystem2012; @huntReducingFreshwaterToxicity2017; @huntCroppingSystemDiversity2019; @liebmanWeedSeedbankDiversity2021] might be less concerned with the risk of waterhemp seedbank size increment as compared to the 2-year rotation.  

The cohort-based female emergence rates at the experiment site were not realistically estimated due to small sample size of the soil seedbank and the plant community [@nguyenReducingHerbicideUse2016]. Similarly, the cohort-based female survival rates at the experiment site were not realistically estimated due to small sample size in a high-efficacy weed management program (Table \@ref(tab:female-cohort-survive-point-est-tab). Future experiments should focus on assessing cohort-based female vital rates, especially in cool-season crop environments. 

## Appendix {-}

### Seed densities at the top and bottom soil strata {-}  

```{r, echo=TRUE}
## ANOVA of Crop ID x Corn weed management effects on AMATA seedbank density at the 0-2 cm soil stratum
AMATA_female_top_lm <- lm(log(AMATA_total_viable_density + 1) ~ Block + 
                     Crop_ID * Corn_weed_management, 
                   data = top_stratum_female)


## ANOVA table of female seedbank density at the 0-2 cm soil stratum
AMATA_female_top_emm_log <- emmeans(AMATA_female_top_lm, 
                                    c("Crop_ID" , "Corn_weed_management"))

joint_tests(AMATA_female_top_emm_log )
```

```{r, echo=TRUE}
## ANOVA of Crop ID x Corn weed management effects on AMATA seedbank density at the 2-20 cm soil stratum
AMATA_female_bottom_lm <- lm(log(AMATA_total_viable_density + 1) ~ Block + 
                     Crop_ID * Corn_weed_management, 
                   data = bottom_stratum_female)
 

## ANOVA table of female seedbank density at the 2-20 cm soil stratum
AMATA_female_bottom_emm_log <- emmeans(AMATA_female_bottom_lm, c("Crop_ID" , "Corn_weed_management"))

joint_tests(AMATA_female_bottom_emm_log)
```

```{r, echo=FALSE, fig.cap = "Diagnosis plots for the effects of crop identity and crop weed managennt on the seedbank densities at the top (A) and bottom (B) soil strata"}

list(AMATA_female_top_lm, AMATA_female_bottom_lm) %>%
   lapply(resid_panel, nrow = 4) %>% 
  wrap_plots() +
  plot_annotation(tag_levels = "A")
```


### Seedbank density from 2014 through 2019 {-}  


```{r, child= "~/Documents/Dissertation/AMATA-demography/4-Analysis/seedbank-fates-2014through2019.Rmd"}

```


### Emergence pattern and timing in different crop environments {-}


```{r, echo = TRUE}
## Did crop identity and corn weed management affect waterhemp's emergence pattern in 2019?
emerge_cohort_19_gls <- gls(log(eu_female_cohort_Density_begin+1)  ~ Block + 
                   Crop_ID*Corn_weed_management + Cohort +
                   Crop_ID:Cohort + Corn_weed_management:Cohort,
                 correlation=corCompSymm(form=~1 | bt),
                 weights=varIdent(form= ~1 | Cohort),
data = density_emerge_19)

 joint_tests(emerge_cohort_19_gls)
```

```{r, echo = TRUE}
## Did crop identity and corn weed management affect waterhemp's emergence pattern in 2020?
# 
emerge_cohort_20_gls <- gls(log(cohort_female_Seedling_density +1)  ~ Block + 
                   Crop_ID*Corn_weed_management + Cohort +
                   Crop_ID:Cohort + Corn_weed_management:Cohort,
                 correlation=corCompSymm(form=~1 | bt),
                 weights=varIdent(form= ~1 | Cohort),
data=cohort_emerge_20_first_six )

 joint_tests(emerge_cohort_20_gls)
```


```{r, echo = TRUE}
## Did crop identity and corn weed management affect waterhemp's first emergence timing in 2019?
Julian_cohort1_19_lm <- lm(Julian_day  ~ Block + 
                         Crop * Corn_weed_management,
data = cohort1_2019) 

joint_tests(Julian_cohort1_19_lm)
```

```{r, echo = TRUE}
## Did crop identity and corn weed management affect waterhemp's first emergence timing in 2020?
Julian_cohort1_20_lm <- lm(Julian_day ~ Block + 
                         Crop*Corn_weed_management,
                         data=cohort1_2020)

joint_tests(Julian_cohort1_20_lm )
```


```{r, echo=FALSE, fig.cap = "Diagnosis plots for the effects of crop identity and corn weed management on the seedbank densities at the top (A) and bottom (B) soil strata"}
list(Julian_cohort1_19_lm, Julian_cohort1_20_lm) %>%
   lapply(resid_panel, nrow = 4) %>% 
  wrap_plots() +
  plot_annotation(tag_levels = "A")
```


```{r emerge-pattern-19, echo=FALSE, fig.cap= "Diagnosis plot for the effects of crop identity, corn weed management, and cohort on seedling densities in 2019"}

plot(emerge_cohort_19_gls, resid(., type = "p") ~ fitted(.)) 
```


```{r emerge-pattern-20, echo=FALSE, fig.cap= "Diagnosis plot for the effects of crop identity, corn weed management, and cohort on seedling densities in 2020"}
plot(emerge_cohort_20_gls, resid(., type = "p") ~ fitted(.))

```



### 2019 female survival rate by cohort {-}

```{r, child= "~/Documents/Dissertation/AMATA-demography/4-Analysis/female-survival-point-est-2019.Rmd"}
# , child = "female-dens-mature.Rmd" removed because density was published
```

### Matrix assembly {-}

```{r, child= "../5-Report/Appendix-matrices-list.Rmd"}

```

### Simulation {-}  

```{r, child= "Q1-fecund-prep.Rmd"}

```

# Acknowledgement {-}

# References {-}