---
output:
    bookdown::word_document2:
      toc: false
      fig_caption: yes
    reference_docx: style_template.docx
bibliography: prospective-pop-dynamics.bib
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(purrr) # for map2()
library(pracma)
library(popbio)
library(tidyverse)
library(magrittr) # for %<>%
# convert this Rmd to R to source it in results.Rmd
library(flextable)
library(patchwork) #for wrap_plots
library(ggResidpanel)
source("functions.R")
theme_set(theme_classic())
```

# Introduction {-}  

```{r, child = "../5-Report/Introduction-model.Rmd"}
```

# Materials and methods {-}

```{r, child = "../5-Report/M-and-M-model.Rmd"}
```

# Results and discussion {-}

## Waterhemp sub-annual demographic parameters from empirical measurement {-}

### Seedling emergence timing {-}


```{r, child= "~/Documents/Dissertation/AMATA-demography/4-Analysis/emergence-pattern-timing-Julian.Rmd"}
# diagnosis plots still look concerning
```

### 2019 soil seedbank and emergence proportion {-} 


```{r, child= "~/Documents/Dissertation/AMATA-demography/5-Report/seedbank-top-bottom.Rmd"}
# , child = "female-dens-mature.Rmd" removed because density was published
```


## Waterhemp population growth rates {-}

### Scenario 1 {-}
Using 2019 fecundity rates, that were estimated from individual plant size using eighteen equations from @nguyenImpactCroppingSystem2022, waterhemp population densities were decreasing in all rotations, the fastest in the 2-year rotation under low herbicide corn weed management and the 3-year rotation under both corn weed management programs (Figures \@ref(fig:scenario1-phase-wise-lambda-plot)). $\lambda$'s decreased right after the first phase of the 2-year (C2) and 3-year (C3) rotations, but increased in the first phase of the 4-year rotation (C4) by three ($\lambda(C4, low)$ = 3.24) to fourteen ($\lambda(C4, low)$ = 14.34) folds. The steady decrease of population density in the soybean, oat, and alfalfa phases ($\lambda$'s = 0.5) were sufficient to keep the rotation-wise populations decreasing within a four year cycle even if $\lambda$ may increase again when the cropping system circles back to corn. The majority of the seedbank in the 4-year rotation is at the 2-20 cm stratum, which limits the seed pool density available for emergence.   

### Scenario 2 {-}

Using 2018 fecundity rates, waterhemp population densities were increasing rapidly in all rotations, the fastest in the 2-year rotation (Figures \@ref(fig:scenario2-phase-wise-lambda-plot)). The population increase in the 3-year rotation under low herbicide corn weed management was similar to that of the 4-year rotation. Even though $\lambda(r)$ were all increasing, the partition of seedbanks differed between rotations. Within the same rotation, the partition of seedbank densities in the two strata of the soil was consistent between two corn weed management regimes. In the 2-year rotation, a large portion of the seedbanks were at the 0-2 cm soil stratum; in the 3-year rotation, the top and bottom soil strata were fairly evenly populated; and in the 4-year rotation, the majority of the seedbanks were at the bottom stratum.   

The oat phase of the 3-year rotation (O3) alfalfa phase of the 4-year rotation (A4) offered an opportunity for decreasing $\lambda$, which was not possible in any other crop phases in the three examined rotations. However, the decline in $\lambda(r)$'s in O3 ($\lambda(O3, low)$ = 0.68) and A4 ($\lambda(A4, low)$ = 0.60) were not strong enough to deplete the replenishment from the corn ($\lambda(C3, low)$ = 2553 and $\lambda(C4, low)$ = 35676.54) and soybean ($\lambda(S3, low)$ = 14567.90 and $\lambda(S4, low)$ = 30.20) phases. 

## Elasticities of popilation growth rates to lower-level demographic parameters {-}

All the elasticity of $\lambda$ to the lower-level demographic parameters mentioned in the brackets below are bold in Tables \@ref(tab:scenario1-E-B-by-rot-tab) and \@ref(tab:scenario2-E-B-by-rot-tab). The tables exclude all-zero columns. Zeros in the tables were from rounding or absolute values being zeros. For examples, the elasticity of $\lambda$'s to seed emergence were negligibly small (at $10^{-9}$ to $10^{-14}$ range), and thus, were excluded from the tables. This pattern suggests that even if weed control programs are of high, or even total efficacy, the very few survived plants, if successfully and prolifically produce new seeds, can help sustain the population.  

### Scenario 1 {-}  
The ranking of parameters or individual parameters to the elasticity of $\lambda_r$ differed by rotation and crop phase. The seed preserving parameters were ranked higher in the corn phase of the 3-year rotation (C3) under both weed management regimes, the soybean phase of the 3-year rotation (S3) that followed corn under low herbicide weed management, the oat phases of the 3-year and 4-year rotations (O3 and O4) regardless of the preceding corn phase's weed management regime, and the alfalfa phase of the 4-year rotation (A4) regardless of the preceding corn phase's weed management regime. The seed producing parameters were ranked higher in the other crop environments (crop species in each rotation).  

The elasticity of $\lambda(r)$ to fecundity rate, as a group, were ranked first in the corn phase of the 2-year rotation (C2) (43.13 and 3.15), the soybean phase of the 2-year rotation (S2) (0.75 and 4.18), the soybean phase of the 3-year rotation (S3) that followed corn under low herbicide weed management program (0.91), the corn phase of the 4-year rotation (C4) (305.24 and 323.07), and the alfalfa in the 4-year rotation (A4) (36.78 and 14.49) under both weed management regimes.
The rate of pre-planting tillage induced seed movement to the 2-20 cm soil stratum was the first-ranked individual parameter to the elasticity of $\lambda(r)$'s in the corn phase of the 3-year rotation (C3) (1.1 and 1.1), the soybean in the 3-year rotation that followed corn under low herbicide weed management (0.89), the oat phase of the 3-year rotation (O3) (0.87 and 0.88), and the oat phase of the 4-year rotation (5.49 and 3.98). The overwinter survival rate of the seeds at the 0-2 cm soil stratum was the first-ranked individual parameter to the elasticity of $\lambda(r)$'s in the soybean phase of the 4-year rotation (S4) (3.32 and 2.41). 

As $\lambda(r)$s were all declining in this scenario, with some opportunity of increasing in the C4 phase, focusing on reducing plant fecundity in C4 would sufficiently stabilize $\lambda(4)$, provided that the weed control efficacy in the subsequent three crop phases remain adequate. 


### Scenario 2 {-}  

Unlike scenario one's dynamics, the difference in the absolute value of the elasticity of $\lambda(r)$ to the first-ranked and second-ranked groups of parameters were multiple orders of magnitude. The seed producing parameters were ranked higher in all crop environments (crop species in each rotation).   

The elasticity of $\lambda(r)$'s to fecundity rates, as a group, were ranked first in all crop phases; but the specific influence of each cohort's fecundity differed by crop phase. The elasticity of $\lambda$'s to cohort one fecundity were the highest in C2 (4.1 x $10^9$ and 2.8 x $10^9$), S3 (4.4 x $10^9$ and 9.7 x $10^{10}$), and O4 (8.3 x $10^6$ and 6.5 x $10^6$) phases. The elasticity of $\lambda_r$'s elasticity to cohort two fecundity were the highest in S2 (1.4 x $10^{10}$ and 8.4 x $10^9$), C3 (1 x $10^{10}$ and 1 x $10^9$), O3 (3 x $10^8$ and 2.5 x $10^7$), C4 (1.6 x $10^9$ and 1 x $10^{10}$), S4 (8.4 x $10^7$ and  8.8 x $10^7$), and A4 (1.1 x $10^7$ and  3 x $10^5$) phases. The second-ranked group of demographic parameters to $\lambda_r$ differed by rotation and crop phase. In the 2-year rotation, the second-ranked group of parameters in contribution to the elasticity of $\lambda(2)$ were both pre-planting tillage-induced seed movement and overwinter seed survival. In the 3-year rotation, the second-ranked group of parameters in contribution to elasticity of $\lambda(3)$ was pre-planting tillage-induced seed movement in the warm-season crop phases; and overwinter seed survival in the cool-season crop phase. In the 4-year rotation, the second-ranked group of parameters in contribution to elasticity of $\lambda(4)$ was overwinter seed survival in the cool-season crop phase and pre-planting tillage-induced seed movement in the warm-season crop phases; a reverse trend as compared to the 3-year rotation's dynamics.

```{r, child = "prospective-dynamics-rotation-wise-projection-scenario1.Rmd"}
```


```{r, child = "prospective-dynamics-rotation-wise-projection-scenario2.Rmd"}
```

## Conclusion {-}

The hypothesis that "extending a conventional 2-year rotation of corn and soybean with cool-season crops can accelerate soil seedbank depletion" was supported. The findings here illustrated that "many little hammers" concept could provide effective control of common waterhemp [@liebmanManyLittleHammers1997]. Unlike giant foxtail whose bottle neck point in the life cycle was overwinter survival rate [@davisCroppingSystemEffects2002], waterhemp's bottle neck point in its life cycle was seedling emergence rate. However, if the early emerged plants were unaffected by weed control programs, either through herbicide resistance or lack of exposure to cultivation, and successfully complete their life cycle, very few massive, prolific female plants would sufficiently replenish the soil seedbank.  

Since the elasticity of $\lambda$ was the least contributed by seed emergence in both the $\lambda$-increasing and $\lambda$-decreasing scenarios, it would be helpful to focus on disrupting seed production by reducing individual plant size with delaying emergence or applying multiple mortality factors to successfully establishing plants, and by limiting seed deposit to the soil. Under the current 4-year rotation's crop sequence, oat (O4) and alfalfa (A4) provided valuable opportunities to reduce plant size [@nguyenWeedCommunityComposition2022; @nguyenImpactCroppingSystem2022]. 

Reduction of herbicide mass [@nguyenWeedCommunityComposition2022] was not coincided with population growing in scenario one ($\lambda < 1$) but coincided with population growing in scenario two ($\lambda > 1$). In consideration of scenario two, $\lambda$-increasing, it would be useful to examine how many years of continuous overwinter cover crops is necessary and which cool-season crop species would be efficient in decreasing waterhemp $\lambda$ after steady seedbank replenishment. This investigation is needed because even if the increased in abundance of waterhemp and other present weed species, has not been observed at the experiment site [@nguyenWeedCommunityComposition2022], an affluent seedbank of a highly competitive weed species habors risks of weed outbreaks. 

The cohort-based female survival rates at the experiment site were not realistically estimated due to small sample size in a high efficacy weed management program. Future experiments should focus on assessing cohort-based female survival rates in cool-season crop environments. 

## Appendix {-}

### Seed densities at the top and bottom soil strata {-}  

```{r, echo=TRUE}
## ANOVA of Crop ID x Corn weed management effects on AMATA seedbank density at the 0-2 cm soil stratum
AMATA_female_top_lm <- lm(log(AMATA_total_viable_density + 1) ~ Block + 
                     Crop_ID * Corn_weed_management, 
                   data = top_stratum_female)


## ANOVA table of female seedbank density at the 0-2 cm soil stratum
AMATA_female_top_emm_log <- emmeans(AMATA_female_top_lm, 
                                    c("Crop_ID" , "Corn_weed_management"))

joint_tests(AMATA_female_top_emm_log )
```

```{r, echo=TRUE}
## ANOVA of Crop ID x Corn weed management effects on AMATA seedbank density at the 2-20 cm soil stratum
AMATA_female_bottom_lm <- lm(log(AMATA_total_viable_density + 1) ~ Block + 
                     Crop_ID * Corn_weed_management, 
                   data = bottom_stratum_female)
 

## ANOVA table of female seedbank density at the 2-20 cm soil stratum
AMATA_female_bottom_emm_log <- emmeans(AMATA_female_bottom_lm, c("Crop_ID" , "Corn_weed_management"))

joint_tests(AMATA_female_bottom_emm_log)
```

```{r, echo=FALSE, fig.cap = "Diagnosis plots for the effects of crop identity and crop weed managennt on the seedbank densities at the top (A) and bottom (B) soil strata"}

list(AMATA_female_top_lm, AMATA_female_bottom_lm) %>%
   lapply(resid_panel, nrow = 4) %>% 
  wrap_plots() +
  plot_annotation(tag_levels = "A")
```

### Emergence timing in different crop environments {-}

```{r, echo = TRUE}
## Did crop identity and corn weed management affect waterhemp's first emergence timing in 2019?
Julian_cohort1_19_lm <- lm(Julian_day  ~ Block + 
                         Crop_ID*Corn_weed_management,
data = cohort1_2019_eu ) 

joint_tests(Julian_cohort1_19_lm)
```

```{r, echo = TRUE}
## Did crop identity and corn weed management affect waterhemp's first emergence timing in 2020?
Julian_cohort1_20_lm <- lm(Julian_day ~ Block + 
                         Crop_ID*Corn_weed_management,
                         data=cohort1_2020)

joint_tests(Julian_cohort1_20_lm )
```

```{r, echo=FALSE, fig.cap = "Diagnosis plots for the effects of crop identity and corn weed management on the seedbank densities at the top (A) and bottom (B) soil strata"}
list(Julian_cohort1_19_lm, Julian_cohort1_20_lm) %>%
   lapply(resid_panel, nrow = 4) %>% 
  wrap_plots() +
  plot_annotation(tag_levels = "A")

```

### 2019 female survival rate by cohort {-}

```{r, child= "~/Documents/Dissertation/AMATA-demography/4-Analysis/female-survival-point-est-2019.Rmd"}
# , child = "female-dens-mature.Rmd" removed because density was published
```

# Reference {-}