---
output:
    bookdown::word_document2:
      toc: false
      fig_caption: yes
    reference_docx: style_template.docx
---

The periodic matrices here are listed by chronological order, from spring tillage to overwinter every year. Each list contains eighteen matrices, corresponding to nine crop identities crossed with two corn weed management regimes. Crop identities in each list are ordered alphabetically.   

#### Pre-planting tillage {-}

Field cultivator before planting all the crops, except alfalfa sole crop (A4).  

```{r, echo=FALSE}
## Pre-planting tillage 
spring_tillage <-  readRDS("../2-Data/Clean/pre-planting-tillage.RData")
spring_tillage 
```

#### Recruitment rate {-}

Numbers were calculated from the stand count densities in 2020 with regards to 2019's seedbank densities. The 2019 seedbank densities were stratified into top (0-2cm) and bottom (2-20cm) sections. The recruitment proportion was calculated under the assumption that seeds only germinate from the top stratum.  
```{r, echo=FALSE}
## Emergence 
emergence_scenario1 <- readRDS("../2-Data/Clean/emergence-prop-scenario1.RData")
emergence_scenario1 
```

#### In-season survival {-}  

The first two elements of the diagonal was adopted from a study that tracked waterhemp seed survival over time at different depths [sosnoskieGlyphosateResistanceDoes2013], and the last six elements calculated from 2019 census survey at our experiment site for cohort survivals. To account for the emergence-delaying effects offered by the cool-season crops, shorter in-season periods were assigned to populations grown in the presence of oat (5.5 months) and alfalfa (4.5 months); populations grown in the presence of corn and soybean were assigned 6 months of in-season. 

```{r, echo=FALSE}
## Seed survival and plant survival 
#### seed survival 

summer_seed_survival_scenario1  <- readRDS("../2-Data/Clean/summer-seed-survival-scenario1.RData")

#### plant survival
female_survival <- readRDS("../2-Data/Clean/female-survival-rate-cohort-equal.RData")
#### combine seed and plant survivals into one matrix with element-wise multiplication 
summer_survival_scenario1 <- purrr::map2(summer_seed_survival_scenario1 , female_survival, `*`)

summer_survival_scenario1
```
#### Fecundity {-}  

Individual fecundity [@nguyenImpactCroppingSystem2022] was summarized to cohort-averaged fecundity by partitioning plant size and fecundity into six size-based group.  
```{r, echo=FALSE}
## Fecundity scenario 2
fecundity_2018 <- readRDS("../2-Data/Clean/fecundity-mean-18-cohort.RData")
fecundity_2018 
```

#### Post-harvest tillage {-}  

At our experiment site, chisel tillage followed corn (C2, C3, and C4); moldboard followed oat in the 3-year rotation (O3) and alfalfa (A4); and no till followed oat in the 4-year rotation (O4) and soybean (S2, S3, and S4). The top left 2x2 section of each matrix was resized from a relevant 18x18 matrix in the Seed Chaser program [@spokasSeedChaserVerticalSoil2007].  

```{r, echo = FALSE}
## Post-harvest tillage
post_harvest_tillage <- readRDS("../2-Data/Clean/post-harvest-tillage.RData")
post_harvest_tillage
```

#### Off-season survival {-}  

The first two elements of the diagonal was also adopted from a study that tracked waterhemp seed survival over time at different depths [sosnoskieGlyphosateResistanceDoes2013]. To account for the emergence-delaying effects offered by the cool-season crops, longer off-season periods were assigned to populations grown in the presence of oat (6.5 months) and alfalfa (7.5 months); populations grown in the presence of corn and soybean were assigned 6 months of off-season.


```{r, echo=FALSE}
## Overwinter survival 
overwinter_scenario1 <- readRDS("../2-Data/Clean/winter-seed-survival-scenario1.RData")

overwinter_scenario1
```
